---
title: "NSF Equations"
author: "Andy"
date: "2025-01-21"
output: pdf_document
---

# Prey Displacement Rate

## Part 1: Predator Encounter Rate

For part I, the encounter rate equation is:

$$ Z = 2rN*sqrt(v^2 + s^2)$$

where $s$ is prey speed (i.e., displacement rate), $v$ is predator speed, $N$ is predator density, and $r$ is the radius within which the predator can perceive prey.

Scott simulates this equation for varying levels of $s$, and four constant levels of $v$ (0, 0.5, 1, and 2). Predator density and detection radius are kept constant such that $2rN = 1$.

$v = 0$ would simply mean a sit-and-wait predator.


## Part 2: Cost vs. Gain for Displacement Rate

Scott mentions incorporating a Type II functional response, such that higher body velocity has diminishing returns due to time spent handling food.

$$ f(R)= \frac{sR}{1 + shR}$$

where $s$ is prey displacement rate, $h$ is handling time, and $R$ is resource density.


Scott describes the fitness curve (without predation, so the growth curve) using the equation from Part 2, and predation using the equation from Part 1.

```{r}
# Generating the baseline curves
trait.vec <- seq(0, 1, by = 0.01)
speed.vec <- 1 - trait.vec


# These formulas are simplified by setting some of the values as constants
# r is always assumed to be 1, while k, s, and v are manipulated values
pred.curve <- function(k, v, s){
  k*sqrt(v^2 + s^2)
}

# R is always assumed to be 1, while s and h are manipulated values
growth.curve <- function(s, R, h){
  (R * s)/(1 + s*R*h)
}


# Build growth curves, for h = 0, 0.5, 1, 2
R_val <- 1

growthcurve.df <- as.data.frame(mat.or.vec(length(speed.vec), 5))
colnames(growthcurve.df) <- c("s", "h_0", "h_0.5", "h_1", "h_2")
growthcurve.df$s <- trait.vec
growthcurve.df$h_0 <- growth.curve(s = growthcurve.df$s, R = R_val, h = 0)
growthcurve.df$h_0.5 <- growth.curve(s = growthcurve.df$s, R = R_val, h = 0.5)
growthcurve.df$h_1 <- growth.curve(s = growthcurve.df$s, R = R_val, h = 1)
growthcurve.df$h_2 <- growth.curve(s = growthcurve.df$s, R = R_val, h = 2)


# Build cost curves, for v = 0, 0.5, 1, 2
k_val <- 1

mortcurve.df <- as.data.frame(mat.or.vec(length(speed.vec), 5))
colnames(mortcurve.df) <- c("s", "v_0", "v_0.5", "v_1", "v_2")
mortcurve.df$s <- trait.vec
mortcurve.df$v_0 <- pred.curve(k = k_val, v = 0, s = mortcurve.df$s)
mortcurve.df$v_0.5 <- pred.curve(k = k_val, v = 0.5, s = mortcurve.df$s)
mortcurve.df$v_1 <- pred.curve(k = k_val, v = 1, s = mortcurve.df$s)
mortcurve.df$v_2 <- pred.curve(k = k_val, v = 2, s = mortcurve.df$s)
```

```{r}
# Combine the two to generate IREF figures

# This is the baseline scenario
baseline.df <- as.data.frame( mat.or.vec( length( speed.vec ), 6) )
colnames(baseline.df) <- c("trait", "speed", "growth_curve", "mort_curve", "fitness", "NCE_CE")
baseline.df$trait <- trait.vec
baseline.df$speed <- speed.vec
h_val <- 8
R_val <- 1
cost_constant_val <- 9
v_val <- 0.2
k_val <- 0.45
pred_constant_val <- 0.98058

baseline.df$growth_curve <- cost_constant_val * growth.curve( s = baseline.df$speed, R = R_val, h = h_val ) # Growth curve
baseline.df$mort_curve <- pred_constant_val * pred.curve( k = k_val, v = v_val, s = baseline.df$speed ) # Mortality curve
baseline.df$fitness <- baseline.df$growth_curve - baseline.df$mort_curve # fitness
baseline.df$NCE_CE <- (1 - baseline.df$growth_curve)/baseline.df$mort_curve # NCE/CE

baseline_optim <- baseline.df$trait[which(baseline.df$fitness == max(baseline.df$fitness))]
baseline_optim_fitness <- baseline.df$fitness[which(baseline.df$fitness == max(baseline.df$fitness))]
baseline_optim_CE <- baseline.df$mort_curve[which(baseline.df$fitness == max(baseline.df$fitness))]
baseline_optim_NCE <- 1 - baseline_optim_CE - baseline_optim_fitness



# Grant figures *appear* to be these params for predator density
# Predator density (k_val) is doubled
preddens.df <- as.data.frame( mat.or.vec( length( speed.vec ), 6) )
colnames(preddens.df) <- c("trait", "speed", "growth_curve", "mort_curve", "fitness", "NCE_CE")
preddens.df$trait <- trait.vec
preddens.df$speed <- speed.vec
h_val <- 8
R_val <- 1
cost_constant_val <- 9
v_val <- 0.2 
k_val <- 0.9 
pred_constant_val <- 0.98058

preddens.df$growth_curve <- cost_constant_val * growth.curve( s = preddens.df$speed, R = R_val, h = h_val ) # Growth curve
preddens.df$mort_curve <- pred_constant_val * pred.curve( k = k_val, v = v_val, s = preddens.df$speed ) # Mortality curve
preddens.df$fitness <- preddens.df$growth_curve - preddens.df$mort_curve # fitness
preddens.df$NCE_CE <- (1 - preddens.df$growth_curve)/preddens.df$mort_curve # NCE/CE

preddens_optim <- preddens.df$trait[which(preddens.df$fitness == max(preddens.df$fitness))]
preddens_optim_fitness <- preddens.df$fitness[which(preddens.df$fitness == max(preddens.df$fitness))]
preddens_optim_CE <- preddens.df$mort_curve[which(preddens.df$fitness == max(preddens.df$fitness))]
preddens_optim_NCE <- 1 - preddens_optim_CE - preddens_optim_fitness


# These *appear* to be the params for foraging mode
# v is five times higher, but k is also (more than) doubled? why?
predmode.df <- as.data.frame( mat.or.vec( length( speed.vec ), 6) )
colnames(predmode.df) <- c("trait", "speed", "growth_curve", "mort_curve", "fitness", "NCE_CE")
predmode.df$trait <- trait.vec
predmode.df$speed <- speed.vec
h_val <- 8
R_val <- 1
cost_constant_val <- 9
v_val <- 1 # this value changes between the baseline and this one, not sure why
k_val <- 1 # This changes from 0.45, which I think is where the change in predator density is coming from
pred_constant_val <- 0.7071068 # this constant changes too, also not sure why


predmode.df$growth_curve <- cost_constant_val * growth.curve( s = predmode.df$speed, R = R_val, h = h_val ) # Growth curve
predmode.df$mort_curve <- pred_constant_val * pred.curve( k = k_val, v = v_val, s = predmode.df$speed ) # Mortality curve
predmode.df$fitness <- predmode.df$growth_curve - predmode.df$mort_curve # fitness
predmode.df$NCE_CE <- (1 - predmode.df$growth_curve)/predmode.df$mort_curve # NCE/CE

predmode_optim <- predmode.df$trait[which(predmode.df$fitness == max(predmode.df$fitness))]
predmode_optim_fitness <- predmode.df$fitness[which(predmode.df$fitness == max(predmode.df$fitness))]
predmode_optim_CE <- predmode.df$mort_curve[which(predmode.df$fitness == max(predmode.df$fitness))]
predmode_optim_NCE <- 1 - predmode_optim_CE - predmode_optim_fitness


# And these *appear* to be the params for resource intake rate
# h is reduced to 1/8 of its original value
# k remains double the baseline, not sure why
prey_h.df <- as.data.frame( mat.or.vec( length( speed.vec ), 6) )
colnames(prey_h.df) <- c("trait", "speed", "growth_curve", "mort_curve", "fitness", "NCE_CE")
prey_h.df$trait <- trait.vec
prey_h.df$speed <- speed.vec
h_val <- 1
R_val <- 1
cost_constant_val <- 2
v_val <- 0.2 
k_val <- 0.9 
pred_constant_val <- 0.98058 

prey_h.df$growth_curve <- cost_constant_val * growth.curve( s = prey_h.df$speed, R = R_val, h = h_val ) # Growth curve
prey_h.df$mort_curve <- pred_constant_val * pred.curve( k = k_val, v = v_val, s = prey_h.df$speed ) # Mortality curve
prey_h.df$fitness <- prey_h.df$growth_curve - prey_h.df$mort_curve # fitness
prey_h.df$NCE_CE <- (1 - prey_h.df$growth_curve)/prey_h.df$mort_curve # NCE/CE

prey_h_optim <- prey_h.df$trait[which(prey_h.df$fitness == max(prey_h.df$fitness))]
prey_h_optim_fitness <- prey_h.df$fitness[which(prey_h.df$fitness == max(prey_h.df$fitness))]
prey_h_optim_CE <- prey_h.df$mort_curve[which(prey_h.df$fitness == max(prey_h.df$fitness))]
prey_h_optim_NCE <- 1 - prey_h_optim_CE - prey_h_optim_fitness
```

```{r}
library(ggplot2)
library(cowplot)

# Fig. 1
fig1_a.plot <- ggplot( data = baseline.df, aes( x = trait ) ) +
  geom_smooth( aes( y = growth_curve ), se = FALSE, color = "blue" ) +
  geom_smooth( aes( y = mort_curve ), se = FALSE, color = "red", linetype = "dashed" ) +
  theme_classic( base_size = 10, base_line_size = 1 ) + 
  ylab("Growth and Mortality") +
  xlab("Displacement Rate") +
  scale_x_continuous(expand = c(0, 0), breaks = c(0, 0.5, 1), labels = c(1000, 500, 1) ) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1), breaks = c(0, 0.5, 1) ) +
  annotate( "text", x = 0.25, y = 0.85, label = "Growth", size = 8/.pt  ) +
  annotate( "text", x = 0.5, y = 0.35, label = "Mortality", size = 8/.pt  ) +
  theme( panel.border = element_rect( color = "black", fill = NA, linewidth = 0.75 ), aspect.ratio = 1 )

fig1_b.plot <- ggplot( data = baseline.df, aes( x = trait ) ) +
  geom_smooth( aes( y = growth_curve, color = "Growth"  ), se = FALSE) +
  geom_smooth( aes( y = mort_curve, color = "Mortality"), se = FALSE, linetype = "dashed" ) +
  geom_smooth( aes( y = fitness, color = "Fitness"), se = FALSE ) +
  theme_classic( base_size = 10, base_line_size = 1 ) + 
  ylab("Growth and Mortality") +
  xlab("Displacement Rate") +
  geom_vline( xintercept = baseline_optim,  size = 0.5, linetype = "dashed", colour = "green" ) +
  annotate( "text", x = 0.3, y = 0.7, label = "Fitness", size = 8/.pt  ) +
  annotate( "text", x = 0.35, y = 0.1, label = "Optimal Trait", size = 8/.pt  ) +
  scale_x_continuous(expand = c(0, 0), breaks = c(0, 0.5, 1), labels = c(1000, 500, 1) ) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1), breaks = c(0, 0.5, 1) ) +
  scale_colour_manual(name = "Curve", values = c("green", "blue", "red")) +
  theme( panel.border = element_rect( color = "black", fill = NA, linewidth = 0.75 ), aspect.ratio = 1,
         axis.title.y = element_blank(), legend.position = "bottom" )

fig1_c.plot <- ggplot( data = baseline.df, aes( x = trait ) ) +
  geom_smooth( aes( y = growth_curve ), color = "blue", se = FALSE ) +
  geom_smooth( aes( y = mort_curve ), color = "red", se = FALSE, linetype = "dashed" ) +
  theme_classic( base_size = 10, base_line_size = 1 ) + 
  ylab("Growth and Mortality") +
  xlab("Displacement Rate") +
  geom_vline( xintercept = baseline_optim,  size = 0.5, linetype = "dashed", colour = "green" ) +
  annotate( "text", x = baseline_optim + 0.15, y = 0.95, label = "NCE = 0.12", size = 7/.pt  ) +
  annotate( "text", x = baseline_optim + 0.15, y = 0.55, label = "0.66", size = 7/.pt  ) +
  annotate( "text", x = baseline_optim + 0.15, y = 0.075, label = "CE = 0.22", size = 7/.pt  ) +
  scale_x_continuous(expand = c(0, 0), breaks = c(0, 0.5, 1), labels = c(1000, 500, 1) ) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1), breaks = c(0, 0.5, 1) ) +
  theme( panel.border = element_rect( color = "black", fill = NA, linewidth = 0.75 ), aspect.ratio = 1,
         axis.title.y = element_blank() )


get_legend_35 <- function(plot) {
  # return all legend candidates
  legends <- get_plot_component(plot, "guide-box", return_all = TRUE)
  # find non-zero legends
  nonzero <- vapply(legends, \(x) !inherits(x, "zeroGrob"), TRUE)
  idx <- which(nonzero)
  # return first non-zero legend if exists, and otherwise first element (which will be a zeroGrob) 
  if (length(idx) > 0) {
    return(legends[[idx[1]]])
  } else {
    return(legends[[1]])
  }
}

fig1.legend <- get_legend_35(fig1_b.plot)
fig1_b.plot <- fig1_b.plot + theme(legend.position = "none")

fig1.plot <- plot_grid( fig1_a.plot, fig1_b.plot, fig1_c.plot, ncol = 3, labels = "auto", label_y = 0.75, align = c("hv") )
fig1.plot <- plot_grid( fig1.plot, fig1.legend, rel_heights = c(1, 0.1), ncol = 1)

# ggsave("fig1.png", plot = fig1.plot, dpi = 300)

```

```{r}
# Fig. 2
baseline.plot <- ggplot( data = baseline.df, aes( x = trait ) ) +
  geom_smooth( aes( y = growth_curve ), se = FALSE, color = "blue" ) +
  geom_smooth( aes( y = mort_curve ), se = FALSE, color = "red", linetype = "dashed" ) +
  theme_classic( base_size = 10, base_line_size = 1 ) + 
  ylab("Growth and Mortality") +
  xlab("Displacement Rate") +
  geom_vline( xintercept = baseline_optim,  size = 0.5, linetype = "dashed", colour = "green" ) +
  annotate( "text", x = baseline_optim + 0.15, y = 0.95, label = "NCE = 0.12", size = 7/.pt  ) +
  annotate( "text", x = baseline_optim + 0.15, y = 0.55, label = "0.66", size = 7/.pt  ) +
  annotate( "text", x = baseline_optim + 0.15, y = 0.075, label = "CE = 0.22", size = 7/.pt  ) +
  scale_x_continuous(expand = c(0, 0), breaks = c(0, 0.5, 1), labels = c(1000, 500, 1) ) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1), breaks = c(0, 0.5, 1) ) +
  theme( panel.border = element_rect( color = "black", fill = NA, linewidth = 0.75 ), aspect.ratio = 1,
         axis.title.x = element_blank(),
         axis.text.x = element_blank() )

preddens.plot <- ggplot( data = preddens.df, aes( x = trait ) ) +
  geom_smooth( aes( y = growth_curve ), se = FALSE, color = "blue" ) +
  geom_smooth( aes( y = mort_curve ), se = FALSE, color = "red", linetype = "dashed" ) +
  theme_classic( base_size = 10, base_line_size = 1 ) + 
  ylab("Growth and Mortality") +
  xlab("Displacement Rate") +
  geom_vline( xintercept = preddens_optim,  size = 0.5, linetype = "dashed", colour = "green" ) +
  annotate( "text", x = preddens_optim + 0.15, y = 0.95, label = "NCE = 0.20", size = 7/.pt  ) +
  annotate( "text", x = preddens_optim + 0.15, y = 0.45, label = "0.48", size = 7/.pt  ) +
  annotate( "text", x = preddens_optim + 0.15, y = 0.15, label = "CE = 0.32", size = 7/.pt  ) +
  scale_x_continuous(expand = c(0, 0), breaks = c(0, 0.5, 1), labels = c(1000, 500, 1) ) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1), breaks = c(0, 0.5, 1) ) +
  theme( panel.border = element_rect( color = "black", fill = NA, linewidth = 0.75 ), aspect.ratio = 1,
         axis.title.x = element_blank(),
         axis.text.x = element_blank(),
         axis.title.y = element_blank(),
         axis.text.y = element_blank() )

predmode.plot <- ggplot( data = predmode.df, aes( x = trait ) ) +
  geom_smooth( aes( y = growth_curve ), se = FALSE, color = "blue" ) +
  geom_smooth( aes( y = mort_curve ), se = FALSE, color = "red", linetype = "dashed" ) +
  theme_classic( base_size = 10, base_line_size = 1 ) + 
  ylab("Growth and Mortality") +
  xlab("Displacement Rate") +
  geom_vline( xintercept = predmode_optim,  size = 0.5, linetype = "dashed", colour = "green" ) +
  annotate( "text", x = predmode_optim + 0.15, y = 0.95, label = "NCE = 0.09", size = 7/.pt  ) +
  annotate( "text", x = predmode_optim + 0.15, y = 0.8, label = "0.11", size = 7/.pt  ) +
  annotate( "text", x = predmode_optim + 0.15, y = 0.5, label = "CE = 0.80", size = 7/.pt  ) +
  scale_x_continuous(expand = c(0, 0), breaks = c(0, 0.5, 1), labels = c(1000, 500, 1) ) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1), breaks = c(0, 0.5, 1) ) +
  theme( panel.border = element_rect( color = "black", fill = NA, linewidth = 0.75 ), aspect.ratio = 1 )

prey_h.plot <- ggplot( data = prey_h.df, aes( x = trait ) ) +
  geom_smooth( aes( y = growth_curve, colour = "Growth" ), se = FALSE ) +
  geom_smooth( aes( y = mort_curve, colour = "Mortality" ), se = FALSE, linetype = "dashed" ) +
  theme_classic( base_size = 10, base_line_size = 1 ) + 
  ylab("Growth and Mortality") +
  xlab("Displacement Rate") +
  geom_vline( xintercept = prey_h_optim,  size = 0.5, linetype = "dashed", colour = "green" ) +
  annotate( "text", x = prey_h_optim + 0.15, y = 0.8, label = "NCE = 0.29", size = 7/.pt  ) +
  annotate( "text", x = prey_h_optim + 0.15, y = 0.5, label = "0.19", size = 7/.pt  ) +
  annotate( "text", x = prey_h_optim + 0.15, y = 0.25, label = "CE = 0.52", size = 7/.pt  ) +
  scale_x_continuous(expand = c(0, 0), breaks = c(0, 0.5, 1), labels = c(1000, 500, 1) ) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1), breaks = c(0, 0.5, 1) ) +
  scale_colour_manual(name = "Curve", values = c("blue", "red")) +
  theme( panel.border = element_rect( color = "black", fill = NA, linewidth = 0.75 ), aspect.ratio = 1,
         axis.title.y = element_blank(), 
         axis.text.y = element_blank(),
         legend.position = "bottom" )



iref.legend <- get_legend_35(prey_h.plot)
prey_h.plot <- prey_h.plot + theme( legend.position = "none" )

iref.fig <- plot_grid( baseline.plot, preddens.plot, predmode.plot, prey_h.plot, ncol = 2, labels = "auto", align = c("hv") )
iref.fig <- plot_grid( iref.fig, iref.legend, ncol = 1, rel_heights = c(1, 0.1) )

# ggsave("fig2.png", plot = iref.fig, dpi = 300)
```
# Predator Density Effects

For this last section, I'd like to make a figure illustrating the very simple, intuitive point that below a certain predator density, the optimal strategy is  actually to not respond at all.

```{r, echo = FALSE}

threshold.df <- as.data.frame( mat.or.vec( length( speed.vec ), 6) )
colnames(threshold.df) <- c("trait", "speed", "growth_curve", "mort_curve", "fitness", "NCE_CE")
threshold.df$trait <- trait.vec
threshold.df$speed <- speed.vec
h_val <- 8
R_val <- 1
cost_constant_val <- 9
v_val <- 1
k_val <- 0.45
pred_constant_val <- 1.41421
p_vec <- seq(0, 1, by = 0.05)

thresh_plot.df <- as.data.frame( mat.or.vec( length(p_vec), 3 ) )
colnames(thresh_plot.df) <- c("preddens", "optim", "relstr")


for (i in 1:length(p_vec)){
  
threshold.df$growth_curve <- cost_constant_val * growth.curve( s = threshold.df$speed, R = R_val, h = h_val ) # Growth curve
threshold.df$mort_curve <- pred_constant_val * p_vec[i] * pred.curve( k = k_val, v = v_val, s = threshold.df$speed ) # Mortality curve
threshold.df$fitness <- threshold.df$growth_curve - threshold.df$mort_curve # fitness
threshold.df$NCE_CE <- (1 - threshold.df$growth_curve)/threshold.df$mort_curve # NCE/CE

thresh_plot.df$preddens[i] <- p_vec[i]
thresh_plot.df$optim[i] <- threshold.df$trait[which(threshold.df$fitness == max(threshold.df$fitness))]
thresh_plot.df$relstr[i] <- (1 - threshold.df$growth_curve[which(threshold.df$fitness == max(threshold.df$fitness))])/((1 - threshold.df$growth_curve[which(threshold.df$fitness == max(threshold.df$fitness))]) + threshold.df$mort_curve[which(threshold.df$fitness == max(threshold.df$fitness))])
}

thresh.plot <- ggplot( data = thresh_plot.df, aes( x = preddens, y = optim ) ) +
  geom_point( ) +
  theme_classic( base_size = 10, base_line_size = 1 ) + 
  ylab("Optimal Trait Value") +
  xlab("Predator Density") +
  scale_x_continuous(expand = c(0, 0) ) + 
  theme( panel.border = element_rect( color = "black", fill = NA, linewidth = 0.75 ), aspect.ratio = 1)

# ggsave("fig3.png", plot = thresh.plot, dpi = 300)

```

# Presence of Refuge Habitat

```{r, echo = FALSE}
## We'll start with the extreme case: what if there's a refuge with *no food* but *no predators*?

# First, set up dataframe and important variables
refuge.df <- as.data.frame( mat.or.vec( 25, 5 ) )
colnames(refuge.df) <- c("proptime", "growth_curve", "mort_curve", "fitness", "NCE_CE")
refuge.df$proptime <- seq(0, 1, by = 1/24) # Prey can spend between 0 and 24 hours in the refuge, incremented by hours

# For now, assume one constant speed by prey in both habitats
# These constants come from the baseline scenario, i.e. Fig. 2 in the grant proposal
# With the addition of some constants to describe conditions in the refuge, noted below
h_val <- 8 # prey handling time
R_val <- 1 # resource density outside of refuge
R_refuge_val <- 0 # resource density inside refuge
cost_constant_val <- 9 # a constant Scott applied to the growth curve to make the numbers cleaner
s_val <- 0.5 # prey velocity
v_val <- 0.2 # predator velocity
k_val <- 0.45 # predator density outside refuge
k_refuge_val <- 0 # predator density in refuge
pred_constant_val <- 0.98058 # a constant Scott applied to the mortality curve to make the numbers cleaner


for ( i in 1:nrow( refuge.df ) ){ # For each possible value of proptime, calculate the total risk and mortality in a 24hr period
refuge.df$growth_curve[i] <- (1 - refuge.df$proptime[i]) * ( cost_constant_val * growth.curve( s = s_val, R = R_val, h = h_val ) ) # Growth curve is simply how much food you can get in the time you spend outside of the refuge

refuge.df$mort_curve[i] <- exp( ( ( refuge.df$proptime[i] ) * log( pred_constant_val * pred.curve( k = k_val, v = v_val, s =
                                                                                            s_val ) ) ) ) * ( pred_constant_val * pred.curve( k = k_val, v = v_val, s =
                                                                                            s_val ) )  # Mortality curve, since risk in the refuge is zero this only considers risk during time spent outside of the refuge
}

refuge.df$fitness <- 1 - ( 1 - refuge.df$growth_curve ) - refuge.df$mort_curve
refuge.df$NCE_CE <- refuge.df$growth_curve / refuge.df$mort_curve

refuge_optim <- refuge.df$proptime[which(refuge.df$fitness == max(refuge.df$fitness))]
refuge_optim_fitness <- refuge.df$fitness[which(refuge.df$fitness == max(refuge.df$fitness))]
refuge_optim_CE <- refuge.df$mort_curve[which(refuge.df$fitness == max(refuge.df$fitness))]
refuge_optim_NCE <- 1 - refuge_optim_CE - refuge_optim_fitness

ggplot( data = refuge.df, aes( x = proptime ) ) +
  geom_smooth( aes( y = growth_curve, colour = "Growth" ), se = FALSE ) +
  geom_smooth( aes( y = mort_curve, colour = "Mortality" ), se = FALSE, linetype = "dashed" ) +
  theme_classic( base_size = 10, base_line_size = 1 ) + 
  ylab("Growth and Mortality") +
  xlab("Proportion of Time Spent in Refuge") +
  scale_x_continuous(expand = c(0, 0), breaks = c(0, 0.5, 1) ) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1), breaks = c(0, 0.5, 1) ) +
  scale_colour_manual(name = "Curve", values = c("blue", "red")) +
  theme( panel.border = element_rect( color = "black", fill = NA, linewidth = 0.75 ), aspect.ratio = 1,
         axis.title.y = element_blank(), 
         axis.text.y = element_blank(),
         legend.position = "bottom" )
```

