---
title: "NSF Equations"
author: "Andy"
date: "2025-01-21"
output: pdf_document
---

# Prey Displacement Rate

## Part 1: Predator Encounter Rate

For part I, the encounter rate equation is:

$$ Z = 2rN*sqrt(v^2 + s^2)$$

where $s$ is prey speed (i.e., displacement rate), $v$ is predator speed, $N$ is predator density, and $r$ is the radius within which the predator can perceive prey.

Scott simulates this equation for varying levels of $s$, and four constant levels of $v$ (0, 0.5, 1, and 2). Predator density and detection radius are kept constant such that $2rN = 1$.

$v = 0$ would simply mean a sit-and-wait predator.


## Part 2: Cost vs. Gain for Displacement Rate

Scott mentions incorporating a Type II functional response, such that higher body velocity has diminishing returns due to time spent handling food.

$$ f(R)= \frac{sR}{1 + shR}$$

where $s$ is prey displacement rate, $h$ is handling time, and $R$ is resource density.


Scott describes the fitness curve (without predation, so the growth curve) using the equation from Part 2, and predation using the equation from Part 1.

```{r}
# Generating the baseline curves
trait.vec <- seq(0, 1, by = 0.01)
speed.vec <- 1 - trait.vec


# These formulas are simplified by setting some of the values as constants
# r is always assumed to be 1, while k, s, and v are manipulated values
pred.curve <- function(k, v, s){
  k*sqrt(v^2 + s^2)
}

# R is always assumed to be 1, while s and h are manipulated values
growth.curve <- function(s, R, h){
  (R * s)/(1 + s*R*h)
}


# Build growth curves, for h = 0, 0.5, 1, 2
R_val <- 1

growthcurve.df <- as.data.frame(mat.or.vec(length(speed.vec), 5))
colnames(growthcurve.df) <- c("s", "h_0", "h_0.5", "h_1", "h_2")
growthcurve.df$s <- trait.vec
growthcurve.df$h_0 <- growth.curve(s = growthcurve.df$s, R = R_val, h = 0)
growthcurve.df$h_0.5 <- growth.curve(s = growthcurve.df$s, R = R_val, h = 0.5)
growthcurve.df$h_1 <- growth.curve(s = growthcurve.df$s, R = R_val, h = 1)
growthcurve.df$h_2 <- growth.curve(s = growthcurve.df$s, R = R_val, h = 2)


# Build cost curves, for v = 0, 0.5, 1, 2
k_val <- 1

mortcurve.df <- as.data.frame(mat.or.vec(length(speed.vec), 5))
colnames(mortcurve.df) <- c("s", "v_0", "v_0.5", "v_1", "v_2")
mortcurve.df$s <- trait.vec
mortcurve.df$v_0 <- pred.curve(k = k_val, v = 0, s = mortcurve.df$s)
mortcurve.df$v_0.5 <- pred.curve(k = k_val, v = 0.5, s = mortcurve.df$s)
mortcurve.df$v_1 <- pred.curve(k = k_val, v = 1, s = mortcurve.df$s)
mortcurve.df$v_2 <- pred.curve(k = k_val, v = 2, s = mortcurve.df$s)
```

```{r}
# Combine the two to generate IREF figures

# This is the baseline scenario
baseline.df <- as.data.frame( mat.or.vec( length( speed.vec ), 6) )
colnames(baseline.df) <- c("trait", "speed", "growth_curve", "mort_curve", "fitness", "NCE_CE")
baseline.df$trait <- trait.vec
baseline.df$speed <- speed.vec
h_val <- 8
R_val <- 1
cost_constant_val <- 9
v_val <- 0.2
k_val <- 0.45
pred_constant_val <- 0.98058

baseline.df$growth_curve <- cost_constant_val * growth.curve( s = baseline.df$speed, R = R_val, h = h_val ) # Growth curve
baseline.df$mort_curve <- pred_constant_val * pred.curve( k = k_val, v = v_val, s = baseline.df$speed ) # Mortality curve
baseline.df$fitness <- baseline.df$growth_curve - baseline.df$mort_curve # fitness
baseline.df$NCE_CE <- (1 - baseline.df$growth_curve)/baseline.df$mort_curve # NCE/CE

baseline_optim <- baseline.df$trait[which(baseline.df$fitness == max(baseline.df$fitness))]
baseline_optim_fitness <- baseline.df$fitness[which(baseline.df$fitness == max(baseline.df$fitness))]
baseline_optim_CE <- baseline.df$mort_curve[which(baseline.df$fitness == max(baseline.df$fitness))]
baseline_optim_NCE <- 1 - baseline_optim_CE - baseline_optim_fitness



# Grant figures *appear* to be these params for predator density
# Predator density (k_val) is doubled
preddens.df <- as.data.frame( mat.or.vec( length( speed.vec ), 6) )
colnames(preddens.df) <- c("trait", "speed", "growth_curve", "mort_curve", "fitness", "NCE_CE")
preddens.df$trait <- trait.vec
preddens.df$speed <- speed.vec
h_val <- 8
R_val <- 1
cost_constant_val <- 9
v_val <- 0.2 
k_val <- 0.9 
pred_constant_val <- 0.98058

preddens.df$growth_curve <- cost_constant_val * growth.curve( s = preddens.df$speed, R = R_val, h = h_val ) # Growth curve
preddens.df$mort_curve <- pred_constant_val * pred.curve( k = k_val, v = v_val, s = preddens.df$speed ) # Mortality curve
preddens.df$fitness <- preddens.df$growth_curve - preddens.df$mort_curve # fitness
preddens.df$NCE_CE <- (1 - preddens.df$growth_curve)/preddens.df$mort_curve # NCE/CE

preddens_optim <- preddens.df$trait[which(preddens.df$fitness == max(preddens.df$fitness))]
preddens_optim_fitness <- preddens.df$fitness[which(preddens.df$fitness == max(preddens.df$fitness))]
preddens_optim_CE <- preddens.df$mort_curve[which(preddens.df$fitness == max(preddens.df$fitness))]
preddens_optim_NCE <- 1 - preddens_optim_CE - preddens_optim_fitness
# Maximum mortality is at 0.8999994


# These *appear* to be the params for foraging mode
# v is five times higher, but k is also (more than) doubled? why?
predmode.df <- as.data.frame( mat.or.vec( length( speed.vec ), 6) )
colnames(predmode.df) <- c("trait", "speed", "growth_curve", "mort_curve", "fitness", "NCE_CE")
predmode.df$trait <- trait.vec
predmode.df$speed <- speed.vec
h_val <- 8
R_val <- 1
cost_constant_val <- 9
v_val <- 1 # this value changes between the baseline and this one, not sure why
k_val <- 1 # This changes from 0.45, which I think is where the change in predator density is coming from
pred_constant_val <- 0.7071068 * 0.8999994 # this changes to anchor the mortality curve in the same spot


predmode.df$growth_curve <- cost_constant_val * growth.curve( s = predmode.df$speed, R = R_val, h = h_val ) # Growth curve
predmode.df$mort_curve <- pred_constant_val * pred.curve( k = k_val, v = v_val, s = predmode.df$speed ) # Mortality curve
predmode.df$fitness <- predmode.df$growth_curve - predmode.df$mort_curve # fitness
predmode.df$NCE_CE <- (1 - predmode.df$growth_curve)/predmode.df$mort_curve # NCE/CE

predmode_optim <- predmode.df$trait[which(predmode.df$fitness == max(predmode.df$fitness))]
predmode_optim_fitness <- predmode.df$fitness[which(predmode.df$fitness == max(predmode.df$fitness))]
predmode_optim_CE <- predmode.df$mort_curve[which(predmode.df$fitness == max(predmode.df$fitness))]
predmode_optim_NCE <- 1 - predmode_optim_CE - predmode_optim_fitness


# And these *appear* to be the params for resource intake rate
# h is reduced to 1/8 of its original value
# k remains double the baseline, not sure why
prey_h.df <- as.data.frame( mat.or.vec( length( speed.vec ), 6) )
colnames(prey_h.df) <- c("trait", "speed", "growth_curve", "mort_curve", "fitness", "NCE_CE")
prey_h.df$trait <- trait.vec
prey_h.df$speed <- speed.vec
h_val <- 1
R_val <- 1
cost_constant_val <- 2
v_val <- 0.2 
k_val <- 0.9 
pred_constant_val <- 0.98058 

prey_h.df$growth_curve <- cost_constant_val * growth.curve( s = prey_h.df$speed, R = R_val, h = h_val ) # Growth curve
prey_h.df$mort_curve <- pred_constant_val * pred.curve( k = k_val, v = v_val, s = prey_h.df$speed ) # Mortality curve
prey_h.df$fitness <- prey_h.df$growth_curve - prey_h.df$mort_curve # fitness
prey_h.df$NCE_CE <- (1 - prey_h.df$growth_curve)/prey_h.df$mort_curve # NCE/CE

prey_h_optim <- prey_h.df$trait[which(prey_h.df$fitness == max(prey_h.df$fitness))]
prey_h_optim_fitness <- prey_h.df$fitness[which(prey_h.df$fitness == max(prey_h.df$fitness))]
prey_h_optim_CE <- prey_h.df$mort_curve[which(prey_h.df$fitness == max(prey_h.df$fitness))]
prey_h_optim_NCE <- 1 - prey_h_optim_CE - prey_h_optim_fitness
```

```{r}
library(ggplot2)
library(cowplot)
library(ggbrace)

# Fig. 1
fig1_a.plot <- ggplot( data = baseline.df, aes( x = trait ) ) +
  geom_smooth( aes( y = growth_curve ), se = FALSE, color = "blue" ) +
  geom_smooth( aes( y = mort_curve ), se = FALSE, color = "red", linetype = "dashed" ) +
  theme_classic( base_size = 10, base_line_size = 1 ) + 
  ylab("Growth and Mortality") +
  xlab("Displacement Rate") +
  scale_x_continuous(expand = c(0, 0), breaks = c(0, 0.5, 1), labels = c(1000, 500, 1) ) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1), breaks = c(0, 0.5, 1) ) +
  annotate( "text", x = 0.25, y = 0.85, label = "Growth", size = 8/.pt  ) +
  annotate( "text", x = 0.5, y = 0.35, label = "Mortality", size = 8/.pt  ) +
  theme( panel.border = element_rect( color = "black", fill = NA, linewidth = 0.75 ), aspect.ratio = 1 )

fig1_b.plot <- ggplot( data = baseline.df, aes( x = trait ) ) +
  geom_smooth( aes( y = growth_curve, color = "Growth"  ), se = FALSE) +
  geom_smooth( aes( y = mort_curve, color = "Mortality"), se = FALSE, linetype = "dashed" ) +
  geom_smooth( aes( y = fitness, color = "Fitness"), se = FALSE ) +
  theme_classic( base_size = 10, base_line_size = 1 ) + 
  ylab("Growth and Mortality") +
  xlab("Displacement Rate") +
  geom_vline( xintercept = baseline_optim,  size = 0.5, linetype = "dashed", colour = "green" ) +
  annotate( "text", x = 0.3, y = 0.7, label = "Fitness", size = 8/.pt  ) +
  annotate( "text", x = 0.35, y = 0.1, label = "Optimal Trait", size = 8/.pt  ) +
  scale_x_continuous(expand = c(0, 0), breaks = c(0, 0.5, 1), labels = c(1000, 500, 1) ) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1), breaks = c(0, 0.5, 1) ) +
  scale_colour_manual(name = "Curve", values = c("green", "blue", "red")) +
  theme( panel.border = element_rect( color = "black", fill = NA, linewidth = 0.75 ), aspect.ratio = 1,
         axis.title.y = element_blank(), legend.position = "bottom" )

fig1_c.plot <- ggplot( data = baseline.df, aes( x = trait ) ) +
  geom_smooth( aes( y = growth_curve ), color = "blue", se = FALSE ) +
  geom_smooth( aes( y = mort_curve ), color = "red", se = FALSE, linetype = "dashed" ) +
  theme_classic( base_size = 10, base_line_size = 1 ) + 
  ylab("Growth and Mortality") +
  xlab("Displacement Rate") +
  annotate(geom = "rect", xmin = baseline_optim - 0.025, xmax = baseline_optim + 0.025, ymin = 1 - baseline_optim_NCE, ymax = 0.998, color = "gray2", fill = "lightblue", alpha = 0.5) +
  annotate( "text", x = baseline_optim + 0.05, y = 1 - baseline_optim_NCE/2, label = "NCE = 0.12", hjust = "left", size = 7/.pt  ) +
  annotate(geom = "rect", xmin = baseline_optim - 0.025, xmax = baseline_optim + 0.025, ymin = baseline_optim_CE, ymax = baseline_optim_CE + baseline_optim_fitness, color = "gray2", fill = "lightgreen", alpha = 0.5) +
  annotate( "text", x = baseline_optim + 0.05, y = baseline_optim_CE + baseline_optim_fitness/2, label = "0.66", hjust = "left", size = 7/.pt  ) +
  annotate(geom = "rect", xmin = baseline_optim - 0.025, xmax = baseline_optim + 0.025, ymin = 0.001, ymax = baseline_optim_CE, color = "gray2", fill = "red", alpha = 0.5) +
  annotate( "text", x = baseline_optim + 0.05, y = baseline_optim_CE/2, label = "CE = 0.22", hjust = "left", size = 7/.pt  ) +
  scale_x_continuous(expand = c(0, 0), breaks = c(0, 0.5, 1), labels = c(1000, 500, 1) ) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1), breaks = c(0, 0.5, 1) ) +
  theme( panel.border = element_rect( color = "black", fill = NA, linewidth = 0.75 ), aspect.ratio = 1,
         axis.title.y = element_blank() )



get_legend_35 <- function(plot) {
  # return all legend candidates
  legends <- get_plot_component(plot, "guide-box", return_all = TRUE)
  # find non-zero legends
  nonzero <- vapply(legends, \(x) !inherits(x, "zeroGrob"), TRUE)
  idx <- which(nonzero)
  # return first non-zero legend if exists, and otherwise first element (which will be a zeroGrob) 
  if (length(idx) > 0) {
    return(legends[[idx[1]]])
  } else {
    return(legends[[1]])
  }
}

fig1.legend <- get_legend_35(fig1_b.plot)
fig1_b.plot <- fig1_b.plot + theme(legend.position = "none")

fig1.plot <- plot_grid( fig1_a.plot, fig1_b.plot, fig1_c.plot, ncol = 3, labels = "auto", label_y = 0.75, align = c("hv") )
fig1.plot <- plot_grid( fig1.plot, fig1.legend, rel_heights = c(1, 0.1), ncol = 1)

# ggsave("fig1.png", plot = fig1.plot, dpi = 300)

```

```{r}
# Fig. 2 - predator density
preddens.df <- cbind(preddens.df, baseline.df$mort_curve)
colnames(preddens.df) <- c("trait", "speed", "growth_curve", "mort_curve", "fitness", "NCE_CE", "mort_baseline")

preddens_1.plot <- ggplot( data = preddens.df, aes( x = trait ) ) +
  geom_line( aes( y = growth_curve ), color = "blue" ) +
  geom_line( aes( y = mort_curve ), color = "red", linetype = "solid" ) +
  geom_line( aes( y = mort_baseline ), color = "red", linetype = "dashed" ) +
  geom_ribbon( aes( ymin = mort_curve, ymax = mort_baseline ), fill = "red", alpha = 0.25 )  +
  geom_segment(aes( x = 0.5, y = 0.3, xend = 0.5, yend = 0.45), arrow = arrow(length = unit(0.5, "cm") ), color = "red", size = 1 ) +
  geom_vline( xintercept = baseline_optim,  size = 0.5, linetype = "dashed", colour = "green" ) +
  theme_classic( base_size = 10, base_line_size = 1 ) + 
  ylab("Growth and Mortality") +
  xlab("Displacement Rate") +
  scale_x_continuous(expand = c(0, 0), breaks = c(0, 0.5, 1), labels = c(1000, 500, 1) ) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1), breaks = c(0, 0.5, 1) ) +
  theme( panel.border = element_rect( color = "black", fill = NA, linewidth = 0.75 ), aspect.ratio = 1,
         axis.title.x = element_blank(),
         axis.text.x = element_blank() )

preddens_2.plot <- ggplot( data = preddens.df, aes( x = trait ) ) +
  geom_line( aes( y = growth_curve, colour = "Growth" ), ) +
  geom_line( aes( y = mort_curve, colour = "Mortality" ), linetype = "solid" ) +
  annotate( "rect", xmin = baseline_optim, xmax = preddens_optim, ymin = 0, ymax = 1, fill = "green", alpha = 0.25 ) +
  theme_classic( base_size = 10, base_line_size = 1 ) + 
  ylab("Growth and Mortality") +
  xlab("Displacement Rate") +
  geom_vline( xintercept = baseline_optim,  size = 0.5, linetype = "dashed", colour = "green" ) +
  geom_segment(aes( x = baseline_optim + 0.01, y = 0.25, xend = preddens_optim - 0.01, yend = 0.25), arrow = arrow(length = unit(0.5, "cm") ), color = "green", size = 1 ) +
  geom_vline( xintercept = preddens_optim,  size = 0.5, linetype = "solid", colour = "green" ) +
  annotate( "text", x = preddens_optim + 0.15, y = 0.95, label = "NCE = 0.20", size = 7/.pt  ) +
  annotate( "text", x = preddens_optim + 0.15, y = 0.45, label = "0.48", size = 7/.pt  ) +
  annotate( "text", x = preddens_optim + 0.15, y = 0.15, label = "CE = 0.32", size = 7/.pt  ) +
  scale_x_continuous(expand = c(0, 0), breaks = c(0, 0.5, 1), labels = c(1000, 500, 1) ) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1), breaks = c(0, 0.5, 1) ) +
  scale_colour_manual(name = "Curve", values = c("blue", "red")) +
  theme( panel.border = element_rect( color = "black", fill = NA, linewidth = 0.75 ), aspect.ratio = 1 ) +
  theme( legend.position = "bottom" )

preddens.legend <- get_legend_35(preddens_2.plot)
preddens_2.plot <- preddens_2.plot + theme( legend.position = "none" )

preddens.fig <- plot_grid( preddens_1.plot, preddens_2.plot, ncol = 1, labels = "auto", hjust = -12.5, align = c("hv") )
preddens.fig <- plot_grid( preddens.fig, preddens.legend, ncol = 1, rel_heights = c(1, 0.1) )

# ggsave("fig2.png", plot = preddens.fig, dpi = 300)
```

```{r}
# Fig. 4 - predator hunting mode
predmode.df <- cbind(predmode.df, preddens.df$mort_curve)
colnames(predmode.df) <- c("trait", "speed", "growth_curve", "mort_curve", "fitness", "NCE_CE", "mort_baseline")

predmode_1.plot <- ggplot( data = predmode.df, aes( x = trait ) ) +
  geom_line( aes( y = growth_curve ), color = "blue" ) +
  geom_line( aes( y = mort_curve ), color = "red", linetype = "solid" ) +
  geom_line( aes( y = mort_baseline ), color = "red", linetype = "dashed" ) +
  geom_ribbon( aes( ymin = mort_curve, ymax = mort_baseline ), fill = "red", alpha = 0.25 )  +
  geom_vline( xintercept = preddens_optim,  size = 0.5, linetype = "dashed", colour = "green" ) +
  geom_segment(aes( x = 0.5, y = 0.5, xend = 0.5, yend = 0.7), arrow = arrow(length = unit(0.5, "cm") ), color = "red", size = 1 ) +
  theme_classic( base_size = 10, base_line_size = 1 ) + 
  ylab("Growth and Mortality") +
  xlab("Displacement Rate") +
  scale_x_continuous(expand = c(0, 0), breaks = c(0, 0.5, 1), labels = c(1000, 500, 1) ) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1), breaks = c(0, 0.5, 1) ) +
  theme( panel.border = element_rect( color = "black", fill = NA, linewidth = 0.75 ), aspect.ratio = 1,
         axis.title.x = element_blank(),
         axis.text.x = element_blank() )

predmode_2.plot <- ggplot( data = predmode.df, aes( x = trait ) ) +
  geom_line( aes( y = growth_curve, colour = "Growth" ), ) +
  geom_line( aes( y = mort_curve, colour = "Mortality" ), linetype = "solid" ) +
  annotate( "rect", xmin = preddens_optim, xmax = predmode_optim, ymin = 0, ymax = 1, fill = "green", alpha = 0.25 ) +
  theme_classic( base_size = 10, base_line_size = 1 ) + 
  ylab("Growth and Mortality") +
  xlab("Displacement Rate") +
  geom_vline( xintercept = preddens_optim,  size = 0.5, linetype = "dashed", colour = "green" ) +
  geom_vline( xintercept = predmode_optim,  size = 0.5, linetype = "solid", colour = "green" ) +
  annotate( "text", x = predmode_optim, y = 0.975, label = "NCE = 0.09", size = 7/.pt  ) +
  annotate( "text", x = predmode_optim - 0.125, y = 0.85, label = "0.11", size = 7/.pt  ) +
  annotate( "text", x = predmode_optim - 0.125, y = 0.5, label = "CE = 0.80", size = 7/.pt  ) +
  geom_segment(aes( x = preddens_optim - 0.01, y = 0.25, xend = predmode_optim + 0.01, yend = 0.25), arrow = arrow(length = unit(0.5, "cm") ), color = "green", size = 1 ) +
  scale_x_continuous(expand = c(0, 0), breaks = c(0, 0.5, 1), labels = c(1000, 500, 1) ) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1), breaks = c(0, 0.5, 1) ) +
  scale_colour_manual(name = "Curve", values = c("blue", "red")) +
  theme( panel.border = element_rect( color = "black", fill = NA, linewidth = 0.75 ), aspect.ratio = 1 ) +
  theme( legend.position = "bottom" )

predmode.legend <- get_legend_35(predmode_2.plot)
predmode_2.plot <- predmode_2.plot + theme( legend.position = "none" )

predmode.fig <- plot_grid( predmode_1.plot, predmode_2.plot, ncol = 1, labels = "auto", hjust = -12.5, align = c("hv") )
predmode.fig <- plot_grid( predmode.fig, predmode.legend, ncol = 1, rel_heights = c(1, 0.1) )

# ggsave("fig4.png", plot = predmode.fig, dpi = 300)
```

```{r}
prey_h.df <- cbind(prey_h.df, baseline.df$growth_curve)
colnames(prey_h.df) <- c("trait", "speed", "growth_curve", "mort_curve", "fitness", "NCE_CE", "growth_baseline")

# Fig. 4 - Prey handling time
prey_h_1.plot <- ggplot( data = prey_h.df, aes( x = trait ) ) +
  geom_line( aes( y = growth_curve ), color = "blue" ) +
  geom_line( aes( y = mort_curve ), color = "red", linetype = "solid" ) +
  geom_line( aes( y = growth_baseline ), color = "blue", linetype = "dashed" ) +
  geom_vline( xintercept = preddens_optim,  size = 0.5, linetype = "dashed", colour = "green" ) +
  geom_ribbon( aes( ymin = growth_baseline, ymax = growth_curve ), fill = "blue", alpha = 0.25 )  +
  geom_segment(aes( x = 0.5, y = 0.85, xend = 0.5, yend = 0.7), arrow = arrow(length = unit(0.5, "cm") ), color = "blue", size = 1 ) +
  theme_classic( base_size = 10, base_line_size = 1 ) + 
  ylab("Growth and Mortality") +
  xlab("Displacement Rate") +
  scale_x_continuous(expand = c(0, 0), breaks = c(0, 0.5, 1), labels = c(1000, 500, 1) ) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1), breaks = c(0, 0.5, 1) ) +
  theme( panel.border = element_rect( color = "black", fill = NA, linewidth = 0.75 ), aspect.ratio = 1,
         axis.title.x = element_blank(),
         axis.text.x = element_blank() )

prey_h_2.plot <- ggplot( data = prey_h.df, aes( x = trait ) ) +
  geom_line( aes( y = growth_curve, colour = "Growth" ), ) +
  geom_line( aes( y = mort_curve, colour = "Mortality" ), linetype = "solid" ) +
  annotate( "rect", xmin = preddens_optim, xmax = prey_h_optim, ymin = 0, ymax = 1, fill = "green", alpha = 0.25 ) +
  theme_classic( base_size = 10, base_line_size = 1 ) + 
  ylab("Growth and Mortality") +
  xlab("Displacement Rate") +
  geom_vline( xintercept = preddens_optim,  size = 0.5, linetype = "dashed", colour = "green" ) +
  geom_vline( xintercept = prey_h_optim,  size = 0.5, linetype = "solid", colour = "green" ) +
  geom_segment(aes( x = preddens_optim - 0.05, y = 0.25, xend = prey_h_optim + 0.01, yend = 0.25), arrow = arrow(length = unit(0.5, "cm") ), color = "green", size = 1 ) +
  annotate( "text", x = prey_h_optim - 0.15, y = 0.925, label = "NCE = 0.29", size = 7/.pt  ) +
  annotate( "text", x = prey_h_optim - 0.15, y = 0.75, label = "0.19", size = 7/.pt  ) +
  annotate( "text", x = prey_h_optim - 0.15, y = 0.35, label = "CE = 0.52", size = 7/.pt  ) +
  scale_x_continuous(expand = c(0, 0), breaks = c(0, 0.5, 1), labels = c(1000, 500, 1) ) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1), breaks = c(0, 0.5, 1) ) +
  scale_colour_manual(name = "Curve", values = c("blue", "red")) +
  theme( panel.border = element_rect( color = "black", fill = NA, linewidth = 0.75 ), aspect.ratio = 1 ) +
  theme( legend.position = "bottom" )

prey_h.legend <- get_legend_35(prey_h_2.plot)
prey_h_2.plot <- prey_h_2.plot + theme( legend.position = "none" )

prey_h.fig <- plot_grid( prey_h_1.plot, prey_h_2.plot, ncol = 1, labels = "auto", hjust = -12.5, align = c("hv") )
prey_h.fig <- plot_grid( prey_h.fig, prey_h.legend, ncol = 1, rel_heights = c(1, 0.1) )

ggsave("fig5.png", plot = prey_h.fig, dpi = 300)
```

# Predator Density Effects

For this last section, I'd like to make a figure illustrating the very simple, intuitive point that below a certain predator density, the optimal strategy is  actually to not respond at all.

```{r, echo = FALSE}

threshold.df <- as.data.frame( mat.or.vec( length( speed.vec ), 6) )
colnames(threshold.df) <- c("trait", "speed", "growth_curve", "mort_curve", "fitness", "NCE_CE")
threshold.df$trait <- trait.vec
threshold.df$speed <- speed.vec
h_val <- 8
R_val <- 1
cost_constant_val <- 9
v_val <- 1 
k_val <- 0.15 
pred_constant_val <- 0.98058

# set p to approx 0.2 or lower

threshold.df$growth_curve <- cost_constant_val * growth.curve( s = threshold.df$speed, R = R_val, h = h_val ) # Growth curve
threshold.df$mort_curve <- pred_constant_val * pred.curve( k = k_val, v = v_val, s = threshold.df$speed ) # Mortality curve
threshold.df$fitness <- threshold.df$growth_curve - threshold.df$mort_curve # fitness
threshold.df$NCE_CE <- (1 - threshold.df$growth_curve)/threshold.df$mort_curve # NCE/CE

thresh_optim <- threshold.df$trait[which(threshold.df$fitness == max(threshold.df$fitness))]
thresh_optim_fitness <- threshold.df$fitness[which(threshold.df$fitness == max(threshold.df$fitness))]
thresh_optim_CE <- threshold.df$mort_curve[which(threshold.df$fitness == max(threshold.df$fitness))]
thresh_optim_NCE <- 1 - thresh_optim_CE - thresh_optim_fitness

thresh.plot <- ggplot( data = threshold.df, aes( x = trait ) ) +
  geom_line( aes( y = growth_curve, colour = "Growth" ), ) +
  geom_line( aes( y = mort_curve, colour = "Mortality" ), linetype = "dashed" ) +
  geom_line( aes( y = fitness, colour = "Fitness" ) ) +
  theme_classic( base_size = 10, base_line_size = 1 ) + 
  ylab("Growth and Mortality") +
  xlab("Displacement Rate") +
  geom_vline( xintercept = thresh_optim,  size = 0.5, linetype = "dashed", colour = "green" ) +
  annotate( "text", x = 0.1, y = 0.95, label = "NCE = 0", size = 7/.pt  ) +
  annotate( "text", x = 0.1, y = 0.65, label = "0.79", size = 7/.pt  ) +
  annotate( "text", x = 0.1, y = 0.15, label = "CE = 0.21", size = 7/.pt  ) +
  scale_x_continuous(expand = c(0, 0), breaks = c(0, 0.5, 1), labels = c(1000, 500, 1) ) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1), breaks = c(0, 0.5, 1) ) +
  scale_colour_manual(name = "Curve", values = c("green", "blue", "red")) +
  theme( panel.border = element_rect( color = "black", fill = NA, linewidth = 0.75 ), aspect.ratio = 1 ) +
  theme( legend.position = "bottom" )

ggsave("fig3.png", plot = thresh.plot, dpi = 300)

```