---
title: "NSF Equations"
author: "Andy"
date: "2025-01-21"
output: pdf_document
---

# Prey Displacement Rate

## Part 1: Predator Encounter Rate

For part I, the encounter rate equation is:

$$ Z = 2rN*sqrt(v^2 + s^2)$$

where $s$ is prey speed (i.e., displacement rate), $v$ is predator speed, $N$ is predator density, and $r$ is the radius within which the predator can perceive prey.

Scott simulates this equation for varying levels of $s$, and four constant levels of $v$ (0, 0.5, 1, and 2). Predator density and detection radius are kept constant such that $2rN = 1$.

$v = 0$ would simply mean a sit-and-wait predator.


## Part 2: Cost vs. Gain for Displacement Rate

Scott mentions incorporating a Type II functional response, such that higher body velocity has diminishing returns due to time spent handling food.

$$ = \frac{sR}{1 + shR}$$

where $s$ is prey displacement rate, $h$ is handling time, and $R$ is resource density.


## Combining the Two

Scott describes the fitness curve (without predation, so the growth curve) using the equation from Part 2, and predation using the equation from Part 1.

```{r}
trait.vec <- seq(0, 1, by = 0.05)
speed.vec <- 1 - trait.vec


# These formulas are simplified by setting some of the values as constants
# r is always assumed to be 1, while N, s, and v are manipulated values
pred.curve <- function(N, v, s){
  N*sqrt(v^2 + s^2)
}

# R is always assumed to be 1, while s and h are manipulated values
growth.curve <- function(k, s, h){
  (k * s)/(1 + s*h)
}

# For temperature-dependence, we'll use a Gaussian curve here
# b is the center of the peak, a is the height, and c is the width
gauss.curve <- function( a, T0, b, c ){ 
 a * exp( -0.5 * ( abs( T0 - b ) / c )^2 )
}

temp.gradient <- seq(5, 30, by = 1)
plot( x = temp.gradient, y = gauss.curve( a = 1, T0 = temp.gradient, b = 25, c = 7.5) )
```

```{r}
# Example 1: Cool-adapted prey and warm-adapted predator
preytemp.curve <- gauss.curve( a = 1, T0 = temp.gradient, b = 20.5, c = 4)
predtemp.curve <- gauss.curve( a = 1, T0 = temp.gradient, b = 23.5, c = 4)

tpc.df <- cbind(temp.gradient, preytemp.curve, predtemp.curve)

# Temperature affects pred velocity, prey velocity, prey handling time
# Make a plot of the two curves side-by-side
tpc.plot <- ggplot( data = tpc.df, aes( x = temp.gradient ) ) +
  geom_smooth( aes( y = predtemp.curve ), se = FALSE, color = "black", linetype = "dashed" ) +
  geom_smooth( aes( y = preytemp.curve ), se = FALSE, color = "black" ) +
  theme_classic( base_size = 10, base_line_size = 1 ) + 
  ylab("Rate") +
  xlab("Temperature (°C)") +
  xlim( 15, 25 ) +
  geom_vline( xintercept = 20, color = "blue", size = 0.5 ) +
  geom_vline( xintercept = 22.5, color = "red", size = 0.5 ) +
  annotate( "text", x = 19.25, y = 0.3, label = "T1", color = "blue", size = 10/.pt  ) +
  annotate( "text", x = 21.75, y = 0.3, label = "T2", color = "red", size = 10/.pt  ) +
  theme( panel.border = element_rect( color = "black", fill = NA, linewidth = 0.75 ), aspect.ratio = 1 )
```

```{r}
# Simulate twice. Use T1 = 20.25, T2 = 22.5.
# Sim1
prey.modifier <- gauss.curve( a = 1, T0 = 20, b = 20.5, c = 4)
pred.modifier <- gauss.curve( a = 1, T0 = 20, b = 23.5, c = 4)

example.df$pred <- pred.curve(N = 1.96116 * 0.45, v = pred.modifier * 0.2, s = prey.modifier * example.df$speed.vec)
example.df$growth <- growth.curve(k = 9, s = prey.modifier * example.df$speed.vec, h = prey.modifier * 8)

# Rescale axis
example.df$growth <- example.df$growth / max( example.df$growth )
example.df$pred <- ( example.df$growth / max( example.df$growth ) ) * example.df$pred


# Find max value for fitness
example.df$fitness <- example.df$growth - example.df$pred

ex1.plot <- ggplot( example.df, aes( x = speed.vec ) ) +
  geom_smooth( aes( y = pred ), se = FALSE, color = "red", linetype = "dashed" ) +
  geom_smooth( aes( y = growth ), se = FALSE ) +
  geom_vline( xintercept = 0.3, color = "green", linetype = "dashed", size = 0.75) +
  theme_classic( base_size = 10, base_line_size = 1 ) + 
  ylab("Fitness, growth, and mortality") +
  xlab("Trait") +
  annotate( "text", x = 0.9, y = 0.1, label = "T1", color = "blue", size = 10/.pt  ) +
  annotate( "text", x = 0.2, y = 0.9, label = "NCE = 0.21", size = 8/.pt ) +
  annotate( "text", x = 0.2, y = 0.5, label = "0.56", size = 8/.pt  ) +
  annotate( "text", x = 0.2, y = 0.05, label = "CE = 0.23", size = 8/.pt  ) +
  scale_x_reverse(expand = c(0, 0), limits = c(1, 0) ) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1)) +
  theme( panel.border = element_rect( color = "black", fill = NA, linewidth = 0.75 ), aspect.ratio = 1 )

```         
         
```{r}       
# Sim2
prey.modifier <- gauss.curve( a = 1, T0 = 22.5, b = 20.5, c = 4)
pred.modifier <- gauss.curve( a = 1, T0 = 22.5, b = 23.5, c = 4)

example.df$pred <- pred.curve(N = 1.96116 * 0.45, v = pred.modifier * 0.2, s = prey.modifier * example.df$speed.vec)
example.df$growth <- growth.curve(k = 9, s = prey.modifier * example.df$speed.vec, h = prey.modifier * 8)

# Rescale axis
example.df$growth <- example.df$growth / max( example.df$growth )
example.df$pred <- ( example.df$growth / max( example.df$growth ) ) * example.df$pred

# Find max value for fitness
example.df$fitness <- example.df$growth - example.df$pred

ex2.plot <- ggplot( example.df, aes( x = speed.vec ) ) +
  geom_smooth( aes( y = pred ), se = FALSE, color = "red", linetype = "dashed" ) +
  geom_smooth( aes( y = growth ), se = FALSE ) +
  geom_vline( xintercept = 0.35, color = "green", linetype = "dashed", size = 0.75) +
  theme_classic( base_size = 10, base_line_size = 1 ) + 
  ylab("Fitness, growth, and mortality") +
  xlab("Trait") +
  annotate( "text", x = 0.9, y = 0.1, label = "T2", color = "red", size = 10/.pt) +
  annotate( "text", x = 0.2, y = 0.9, label = "NCE = 0.20", size = 8/.pt) +
  annotate( "text", x = 0.2, y = 0.5, label = "0.54", size = 8/.pt) +
  annotate( "text", x = 0.2, y = 0.05, label = "CE = 0.26", size = 8/.pt) +
  scale_x_reverse(expand = c(0, 0), limits = c(1, 0) ) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1)) +
  theme( panel.border = element_rect( color = "black", fill = NA, linewidth = 0.75 ), aspect.ratio = 1 )

library(cowplot)
tpc.fig <- plot_grid( tpc.plot, ex1.plot, ex2.plot, ncol = 1 )

ggsave("tpcfig.png", plot = tpc.fig, dpi = 300)
```

