---
title: "NSF Equations"
author: "Andy"
date: "2025-01-21"
output: pdf_document
---

# Prey Displacement Rate

## Part 1: Predator Encounter Rate

For part I, the encounter rate equation is:

$$ Z = 2rN*sqrt(v^2 + s^2)$$

where $s$ is prey speed (i.e., displacement rate), $v$ is predator speed, $N$ is predator density, and $r$ is the radius within which the predator can perceive prey.

Scott simulates this equation for varying levels of $s$, and four constant levels of $v$ (0, 0.5, 1, and 2). Predator density and detection radius are kept constant such that $2rN = 1$.

$v = 0$ would simply mean a sit-and-wait predator.


## Part 2: Cost vs. Gain for Displacement Rate

Scott mentions incorporating a Type II functional response, such that higher body velocity has diminishing returns due to time spent handling food.

$$ f(R)= \frac{sR}{1 + shR}$$

where $s$ is prey displacement rate, $h$ is handling time, and $R$ is resource density.


Scott describes the fitness curve (without predation, so the growth curve) using the equation from Part 2, and predation using the equation from Part 1.

Questions for Scott: what is the cost constant? pred constant? P factor? How are those incorporated into the base equations?

What range of parameters did he use for the figures in the grant? What are the baselines used for Fig. 2?

Where is the analysis for resource intake rate?

```{r}
# Generating the baseline curves
trait.vec <- seq(0, 1, by = 0.01)
speed.vec <- 1 - trait.vec


# These formulas are simplified by setting some of the values as constants
# r is always assumed to be 1, while k, s, and v are manipulated values
pred.curve <- function(k, v, s){
  k*sqrt(v^2 + s^2)
}

# R is always assumed to be 1, while s and h are manipulated values
growth.curve <- function(s, R, h){
  (R * s)/(1 + s*R*h)
}


# Build growth curves, for h = 0, 0.5, 1, 2
R_val <- 1

growthcurve.df <- as.data.frame(mat.or.vec(length(speed.vec), 5))
colnames(growthcurve.df) <- c("s", "h_0", "h_0.5", "h_1", "h_2")
growthcurve.df$s <- trait.vec
growthcurve.df$h_0 <- growth.curve(s = growthcurve.df$s, R = R_val, h = 0)
growthcurve.df$h_0.5 <- growth.curve(s = growthcurve.df$s, R = R_val, h = 0.5)
growthcurve.df$h_1 <- growth.curve(s = growthcurve.df$s, R = R_val, h = 1)
growthcurve.df$h_2 <- growth.curve(s = growthcurve.df$s, R = R_val, h = 2)


# Build cost curves, for v = 0, 0.5, 1, 2
k_val <- 1

mortcurve.df <- as.data.frame(mat.or.vec(length(speed.vec), 5))
colnames(mortcurve.df) <- c("s", "v_0", "v_0.5", "v_1", "v_2")
mortcurve.df$s <- trait.vec
mortcurve.df$v_0 <- pred.curve(k = k_val, v = 0, s = mortcurve.df$s)
mortcurve.df$v_0.5 <- pred.curve(k = k_val, v = 0.5, s = mortcurve.df$s)
mortcurve.df$v_1 <- pred.curve(k = k_val, v = 1, s = mortcurve.df$s)
mortcurve.df$v_2 <- pred.curve(k = k_val, v = 2, s = mortcurve.df$s)
```

```{r}
# Combine the two to generate IREF figures

# This is the baseline scenario
baseline.df <- as.data.frame( mat.or.vec( length( speed.vec ), 6) )
colnames(baseline.df) <- c("trait", "speed", "growth_curve", "mort_curve", "fitness", "NCE_CE")
baseline.df$trait <- trait.vec
baseline.df$speed <- speed.vec
h_val <- 8
R_val <- 1
cost_constant_val <- 9
v_val <- 0.2
k_val <- 0.45
pred_constant_val <- 0.98058

baseline.df$growth_curve <- cost_constant_val * growth.curve( s = baseline.df$speed, R = R_val, h = h_val ) # Growth curve
baseline.df$mort_curve <- pred_constant_val * pred.curve( k = k_val, v = v_val, s = baseline.df$speed ) # Mortality curve
baseline.df$fitness <- baseline.df$growth_curve - baseline.df$mort_curve # fitness
baseline.df$NCE_CE <- (1 - baseline.df$growth_curve)/baseline.df$mort_curve # NCE/CE

baseline_optim <- baseline.df$trait[which(baseline.df$fitness == max(baseline.df$fitness))]
baseline_optim_fitness <- baseline.df$fitness[which(baseline.df$fitness == max(baseline.df$fitness))]
baseline_optim_CE <- baseline.df$mort_curve[which(baseline.df$fitness == max(baseline.df$fitness))]
baseline_optim_NCE <- 1 - baseline_optim_CE - baseline_optim_fitness



# Grant figures *appear* to be these params for predator density
# Predator density (k_val) is doubled
preddens.df <- as.data.frame( mat.or.vec( length( speed.vec ), 6) )
colnames(preddens.df) <- c("trait", "speed", "growth_curve", "mort_curve", "fitness", "NCE_CE")
preddens.df$trait <- trait.vec
preddens.df$speed <- speed.vec
h_val <- 8
R_val <- 1
cost_constant_val <- 9
v_val <- 1 
k_val <- 0.9 
pred_constant_val <- 0.7071068

preddens.df$growth_curve <- cost_constant_val * growth.curve( s = preddens.df$speed, R = R_val, h = h_val ) # Growth curve
preddens.df$mort_curve <- pred_constant_val * pred.curve( k = k_val, v = v_val, s = preddens.df$speed ) # Mortality curve
preddens.df$fitness <- preddens.df$growth_curve - preddens.df$mort_curve # fitness
preddens.df$NCE_CE <- (1 - preddens.df$growth_curve)/preddens.df$mort_curve # NCE/CE

preddens_optim <- preddens.df$trait[which(preddens.df$fitness == max(preddens.df$fitness))]
preddens_optim_fitness <- preddens.df$fitness[which(preddens.df$fitness == max(preddens.df$fitness))]
preddens_optim_CE <- preddens.df$mort_curve[which(preddens.df$fitness == max(preddens.df$fitness))]
preddens_optim_NCE <- 1 - preddens_optim_CE - preddens_optim_fitness


# These *appear* to be the params for foraging mode
# v is five times higher, but k is also (more than) doubled? why?
predmode.df <- as.data.frame( mat.or.vec( length( speed.vec ), 6) )
colnames(predmode.df) <- c("trait", "speed", "growth_curve", "mort_curve", "fitness", "NCE_CE")
predmode.df$trait <- trait.vec
predmode.df$speed <- speed.vec
h_val <- 8
R_val <- 1
cost_constant_val <- 9
v_val <- 1 # this value changes between the baseline and this one, not sure why
k_val <- 1 # This changes from 0.45, which I think is where the change in predator density is coming from
pred_constant_val <- 0.7071068 # this constant changes too, also not sure why


predmode.df$growth_curve <- cost_constant_val * growth.curve( s = predmode.df$speed, R = R_val, h = h_val ) # Growth curve
predmode.df$mort_curve <- pred_constant_val * pred.curve( k = k_val, v = v_val, s = predmode.df$speed ) # Mortality curve
predmode.df$fitness <- predmode.df$growth_curve - predmode.df$mort_curve # fitness
predmode.df$NCE_CE <- (1 - predmode.df$growth_curve)/predmode.df$mort_curve # NCE/CE

predmode_optim <- predmode.df$trait[which(predmode.df$fitness == max(predmode.df$fitness))]
predmode_optim_fitness <- predmode.df$fitness[which(predmode.df$fitness == max(predmode.df$fitness))]
predmode_optim_CE <- predmode.df$mort_curve[which(predmode.df$fitness == max(predmode.df$fitness))]
predmode_optim_NCE <- 1 - predmode_optim_CE - predmode_optim_fitness


# And these *appear* to be the params for resource intake rate
# h is reduced to 1/8 of its original value
# k remains double the baseline, not sure why
prey_h.df <- as.data.frame( mat.or.vec( length( speed.vec ), 6) )
colnames(prey_h.df) <- c("trait", "speed", "growth_curve", "mort_curve", "fitness", "NCE_CE")
prey_h.df$trait <- trait.vec
prey_h.df$speed <- speed.vec
h_val <- 1
R_val <- 1
cost_constant_val <- 2
v_val <- 0.2 
k_val <- 0.9 
pred_constant_val <- 0.98058 

prey_h.df$growth_curve <- cost_constant_val * growth.curve( s = prey_h.df$speed, R = R_val, h = h_val ) # Growth curve
prey_h.df$mort_curve <- pred_constant_val * pred.curve( k = k_val, v = v_val, s = prey_h.df$speed ) # Mortality curve
prey_h.df$fitness <- prey_h.df$growth_curve - prey_h.df$mort_curve # fitness
prey_h.df$NCE_CE <- (1 - prey_h.df$growth_curve)/prey_h.df$mort_curve # NCE/CE

prey_h_optim <- prey_h.df$trait[which(prey_h.df$fitness == max(prey_h.df$fitness))]
prey_h_optim_fitness <- prey_h.df$fitness[which(prey_h.df$fitness == max(prey_h.df$fitness))]
prey_h_optim_CE <- prey_h.df$mort_curve[which(prey_h.df$fitness == max(prey_h.df$fitness))]
prey_h_optim_NCE <- 1 - prey_h_optim_CE - prey_h_optim_fitness
```

```{r}
library(ggplot2)
library(cowplot)

# Fig. 1
fig1_a.plot <- ggplot( data = baseline.df, aes( x = trait ) ) +
  geom_smooth( aes( y = growth_curve ), se = FALSE, color = "blue" ) +
  geom_smooth( aes( y = mort_curve ), se = FALSE, color = "red", linetype = "dashed" ) +
  theme_classic( base_size = 10, base_line_size = 1 ) + 
  ylab("Growth and Mortality") +
  xlab("Displacement Rate") +
  scale_x_continuous(expand = c(0, 0), breaks = c(0, 0.5, 1), labels = c(1000, 500, 1) ) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1), breaks = c(0, 0.5, 1) ) +
  annotate( "text", x = 0.25, y = 0.85, label = "Growth", size = 8/.pt  ) +
  annotate( "text", x = 0.5, y = 0.35, label = "Mortality", size = 8/.pt  ) +
  theme( panel.border = element_rect( color = "black", fill = NA, linewidth = 0.75 ), aspect.ratio = 1 )

fig1_b.plot <- ggplot( data = baseline.df, aes( x = trait ) ) +
  geom_smooth( aes( y = growth_curve, color = "Growth"  ), se = FALSE) +
  geom_smooth( aes( y = mort_curve, color = "Mortality"), se = FALSE, linetype = "dashed" ) +
  geom_smooth( aes( y = fitness, color = "Fitness"), se = FALSE ) +
  theme_classic( base_size = 10, base_line_size = 1 ) + 
  ylab("Growth and Mortality") +
  xlab("Displacement Rate") +
  geom_vline( xintercept = baseline_optim,  size = 0.5, linetype = "dashed", colour = "green" ) +
  annotate( "text", x = 0.3, y = 0.7, label = "Fitness", size = 8/.pt  ) +
  annotate( "text", x = 0.35, y = 0.1, label = "Optimal Trait", size = 8/.pt  ) +
  scale_x_continuous(expand = c(0, 0), breaks = c(0, 0.5, 1), labels = c(1000, 500, 1) ) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1), breaks = c(0, 0.5, 1) ) +
  scale_colour_manual(name = "Curve", values = c("green", "blue", "red")) +
  theme( panel.border = element_rect( color = "black", fill = NA, linewidth = 0.75 ), aspect.ratio = 1,
         axis.title.y = element_blank(), legend.position = "bottom" )

fig1_c.plot <- ggplot( data = baseline.df, aes( x = trait ) ) +
  geom_smooth( aes( y = growth_curve ), color = "blue", se = FALSE ) +
  geom_smooth( aes( y = mort_curve ), color = "red", se = FALSE, linetype = "dashed" ) +
  theme_classic( base_size = 10, base_line_size = 1 ) + 
  ylab("Growth and Mortality") +
  xlab("Displacement Rate") +
  geom_vline( xintercept = baseline_optim,  size = 0.5, linetype = "dashed", colour = "green" ) +
  annotate( "text", x = baseline_optim + 0.15, y = 0.95, label = "NCE = 0.12", size = 7/.pt  ) +
  annotate( "text", x = baseline_optim + 0.15, y = 0.55, label = "0.66", size = 7/.pt  ) +
  annotate( "text", x = baseline_optim + 0.15, y = 0.075, label = "CE = 0.22", size = 7/.pt  ) +
  scale_x_continuous(expand = c(0, 0), breaks = c(0, 0.5, 1), labels = c(1000, 500, 1) ) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1), breaks = c(0, 0.5, 1) ) +
  theme( panel.border = element_rect( color = "black", fill = NA, linewidth = 0.75 ), aspect.ratio = 1,
         axis.title.y = element_blank() )


get_legend_35 <- function(plot) {
  # return all legend candidates
  legends <- get_plot_component(plot, "guide-box", return_all = TRUE)
  # find non-zero legends
  nonzero <- vapply(legends, \(x) !inherits(x, "zeroGrob"), TRUE)
  idx <- which(nonzero)
  # return first non-zero legend if exists, and otherwise first element (which will be a zeroGrob) 
  if (length(idx) > 0) {
    return(legends[[idx[1]]])
  } else {
    return(legends[[1]])
  }
}

fig1.legend <- get_legend_35(fig1_b.plot)
fig1_b.plot <- fig1_b.plot + theme(legend.position = "none")

fig1.plot <- plot_grid( fig1_a.plot, fig1_b.plot, fig1_c.plot, ncol = 3, labels = "auto", label_y = 0.75, align = c("hv") )
fig1.plot <- plot_grid( fig1.plot, fig1.legend, rel_heights = c(1, 0.1), ncol = 1)

ggsave("fig1.png", plot = fig1.plot, dpi = 300)

```

```{r}
# Fig. 2
baseline.plot <- ggplot( data = baseline.df, aes( x = trait ) ) +
  geom_smooth( aes( y = growth_curve ), se = FALSE, color = "blue" ) +
  geom_smooth( aes( y = mort_curve ), se = FALSE, color = "red", linetype = "dashed" ) +
  theme_classic( base_size = 10, base_line_size = 1 ) + 
  ylab("Growth and Mortality") +
  xlab("Displacement Rate") +
  geom_vline( xintercept = baseline_optim,  size = 0.5, linetype = "dashed" ) +
  annotate( "text", x = baseline_optim + 0.15, y = 0.95, label = "NCE = 0.12", size = 7/.pt  ) +
  annotate( "text", x = baseline_optim + 0.15, y = 0.55, label = "0.66", size = 7/.pt  ) +
  annotate( "text", x = baseline_optim + 0.15, y = 0.075, label = "CE = 0.22", size = 7/.pt  ) +
  scale_x_continuous(expand = c(0, 0), breaks = c(0, 0.5, 1), labels = c(1000, 500, 1) ) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1), breaks = c(0, 0.5, 1) ) +
  theme( panel.border = element_rect( color = "black", fill = NA, linewidth = 0.75 ), aspect.ratio = 1,
         axis.title.x = element_blank(),
         axis.text.x = element_blank() )

preddens.plot <- ggplot( data = preddens.df, aes( x = trait ) ) +
  geom_smooth( aes( y = growth_curve ), se = FALSE, color = "blue" ) +
  geom_smooth( aes( y = mort_curve ), se = FALSE, color = "red", linetype = "dashed" ) +
  theme_classic( base_size = 10, base_line_size = 1 ) + 
  ylab("Growth and Mortality") +
  xlab("Displacement Rate") +
  geom_vline( xintercept = preddens_optim,  size = 0.5, linetype = "dashed" ) +
  annotate( "text", x = preddens_optim + 0.15, y = 0.95, label = "NCE = 0.08", size = 7/.pt  ) +
  annotate( "text", x = preddens_optim + 0.15, y = 0.8, label = "0.20", size = 7/.pt  ) +
  annotate( "text", x = preddens_optim + 0.15, y = 0.6, label = "CE = 0.72", size = 7/.pt  ) +
  scale_x_continuous(expand = c(0, 0), breaks = c(0, 0.5, 1), labels = c(1000, 500, 1) ) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1), breaks = c(0, 0.5, 1) ) +
  theme( panel.border = element_rect( color = "black", fill = NA, linewidth = 0.75 ), aspect.ratio = 1,
         axis.title.x = element_blank(),
         axis.text.x = element_blank(),
         axis.title.y = element_blank(),
         axis.text.y = element_blank() )

predmode.plot <- ggplot( data = predmode.df, aes( x = trait ) ) +
  geom_smooth( aes( y = growth_curve ), se = FALSE, color = "blue" ) +
  geom_smooth( aes( y = mort_curve ), se = FALSE, color = "red", linetype = "dashed" ) +
  theme_classic( base_size = 10, base_line_size = 1 ) + 
  ylab("Growth and Mortality") +
  xlab("Displacement Rate") +
  geom_vline( xintercept = predmode_optim,  size = 0.5, linetype = "dashed" ) +
  annotate( "text", x = predmode_optim + 0.15, y = 0.95, label = "NCE = 0.09", size = 7/.pt  ) +
  annotate( "text", x = predmode_optim + 0.15, y = 0.8, label = "0.11", size = 7/.pt  ) +
  annotate( "text", x = predmode_optim + 0.15, y = 0.5, label = "CE = 0.80", size = 7/.pt  ) +
  scale_x_continuous(expand = c(0, 0), breaks = c(0, 0.5, 1), labels = c(1000, 500, 1) ) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1), breaks = c(0, 0.5, 1) ) +
  theme( panel.border = element_rect( color = "black", fill = NA, linewidth = 0.75 ), aspect.ratio = 1 )

prey_h.plot <- ggplot( data = prey_h.df, aes( x = trait ) ) +
  geom_smooth( aes( y = growth_curve, colour = "Growth" ), se = FALSE ) +
  geom_smooth( aes( y = mort_curve, colour = "Mortality" ), se = FALSE, linetype = "dashed" ) +
  theme_classic( base_size = 10, base_line_size = 1 ) + 
  ylab("Growth and Mortality") +
  xlab("Displacement Rate") +
  geom_vline( xintercept = prey_h_optim,  size = 0.5, linetype = "dashed" ) +
  annotate( "text", x = prey_h_optim + 0.15, y = 0.8, label = "NCE = 0.29", size = 7/.pt  ) +
  annotate( "text", x = prey_h_optim + 0.15, y = 0.5, label = "0.19", size = 7/.pt  ) +
  annotate( "text", x = prey_h_optim + 0.15, y = 0.25, label = "CE = 0.52", size = 7/.pt  ) +
  scale_x_continuous(expand = c(0, 0), breaks = c(0, 0.5, 1), labels = c(1000, 500, 1) ) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1), breaks = c(0, 0.5, 1) ) +
  scale_colour_manual(name = "Curve", values = c("blue", "red")) +
  theme( panel.border = element_rect( color = "black", fill = NA, linewidth = 0.75 ), aspect.ratio = 1,
         axis.title.y = element_blank(), 
         axis.text.y = element_blank(),
         legend.position = "bottom" )



iref.legend <- get_legend_35(prey_h.plot)
prey_h.plot <- prey_h.plot + theme( legend.position = "none" )

iref.fig <- plot_grid( baseline.plot, preddens.plot, predmode.plot, prey_h.plot, ncol = 2, labels = "auto", align = c("hv") )
iref.fig <- plot_grid( iref.fig, iref.legend, ncol = 1, rel_heights = c(1, 0.1) )

ggsave("fig2.png", plot = iref.fig, dpi = 300)
```


# Temperature-Dependent Examples

```{r, eval = FALSE}
# Example 1: Cool-adapted prey and warm-adapted predator
preytemp.curve <- gauss.curve( a = 1, T0 = temp.gradient, b = 20.5, c = 4)
predtemp.curve <- gauss.curve( a = 1, T0 = temp.gradient, b = 23.5, c = 4)

tpc.df <- cbind(temp.gradient, preytemp.curve, predtemp.curve)

# Temperature affects pred velocity, prey velocity, prey handling time
# Make a plot of the two curves side-by-side
tpc.plot <- ggplot( data = tpc.df, aes( x = temp.gradient ) ) +
  geom_smooth( aes( y = predtemp.curve ), se = FALSE, color = "black", linetype = "dashed" ) +
  geom_smooth( aes( y = preytemp.curve ), se = FALSE, color = "black" ) +
  theme_classic( base_size = 10, base_line_size = 1 ) + 
  ylab("Rate") +
  xlab("Temperature (°C)") +
  xlim( 15, 25 ) +
  geom_vline( xintercept = 20, color = "blue", size = 0.5 ) +
  geom_vline( xintercept = 22.5, color = "red", size = 0.5 ) +
  annotate( "text", x = 19.25, y = 0.3, label = "T1", color = "blue", size = 10/.pt  ) +
  annotate( "text", x = 21.75, y = 0.3, label = "T2", color = "red", size = 10/.pt  ) +
  theme( panel.border = element_rect( color = "black", fill = NA, linewidth = 0.75 ), aspect.ratio = 1 )
```

```{r, eval = FALSE}
# Simulate twice. Use T1 = 20.25, T2 = 22.5.
# Sim1
prey.modifier <- gauss.curve( a = 1, T0 = 20, b = 20.5, c = 4)
pred.modifier <- gauss.curve( a = 1, T0 = 20, b = 23.5, c = 4)

example.df$pred <- pred.curve(N = 1.96116 * 0.45, v = pred.modifier * 0.2, s = prey.modifier * example.df$speed.vec)
example.df$growth <- growth.curve(k = 9, s = prey.modifier * example.df$speed.vec, h = prey.modifier * 8)

# Rescale axis
example.df$growth <- example.df$growth / max( example.df$growth )
example.df$pred <- ( example.df$growth / max( example.df$growth ) ) * example.df$pred


# Find max value for fitness
example.df$fitness <- example.df$growth - example.df$pred

ex1.plot <- ggplot( example.df, aes( x = speed.vec ) ) +
  geom_smooth( aes( y = pred ), se = FALSE, color = "red", linetype = "dashed" ) +
  geom_smooth( aes( y = growth ), se = FALSE ) +
  geom_vline( xintercept = 0.3, color = "green", linetype = "dashed", size = 0.75) +
  theme_classic( base_size = 10, base_line_size = 1 ) + 
  ylab("Fitness, growth, and mortality") +
  xlab("Trait") +
  annotate( "text", x = 0.9, y = 0.1, label = "T1", color = "blue", size = 10/.pt  ) +
  annotate( "text", x = 0.2, y = 0.9, label = "NCE = 0.21", size = 8/.pt ) +
  annotate( "text", x = 0.2, y = 0.5, label = "0.56", size = 8/.pt  ) +
  annotate( "text", x = 0.2, y = 0.05, label = "CE = 0.23", size = 8/.pt  ) +
  scale_x_reverse(expand = c(0, 0), limits = c(1, 0) ) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1)) +
  theme( panel.border = element_rect( color = "black", fill = NA, linewidth = 0.75 ), aspect.ratio = 1 )

```         
         
```{r, eval = FALSE}
# Sim2
prey.modifier <- gauss.curve( a = 1, T0 = 22.5, b = 20.5, c = 4)
pred.modifier <- gauss.curve( a = 1, T0 = 22.5, b = 23.5, c = 4)

example.df$pred <- pred.curve(N = 1.96116 * 0.45, v = pred.modifier * 0.2, s = prey.modifier * example.df$speed.vec)
example.df$growth <- growth.curve(k = 9, s = prey.modifier * example.df$speed.vec, h = prey.modifier * 8)

# Rescale axis
example.df$growth <- example.df$growth / max( example.df$growth )
example.df$pred <- ( example.df$growth / max( example.df$growth ) ) * example.df$pred

# Find max value for fitness
example.df$fitness <- example.df$growth - example.df$pred

ex2.plot <- ggplot( example.df, aes( x = speed.vec ) ) +
  geom_smooth( aes( y = pred ), se = FALSE, color = "red", linetype = "dashed" ) +
  geom_smooth( aes( y = growth ), se = FALSE ) +
  geom_vline( xintercept = 0.35, color = "green", linetype = "dashed", size = 0.75) +
  theme_classic( base_size = 10, base_line_size = 1 ) + 
  ylab("Fitness, growth, and mortality") +
  xlab("Trait") +
  annotate( "text", x = 0.9, y = 0.1, label = "T2", color = "red", size = 10/.pt) +
  annotate( "text", x = 0.2, y = 0.9, label = "NCE = 0.20", size = 8/.pt) +
  annotate( "text", x = 0.2, y = 0.5, label = "0.54", size = 8/.pt) +
  annotate( "text", x = 0.2, y = 0.05, label = "CE = 0.26", size = 8/.pt) +
  scale_x_reverse(expand = c(0, 0), limits = c(1, 0) ) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1)) +
  theme( panel.border = element_rect( color = "black", fill = NA, linewidth = 0.75 ), aspect.ratio = 1 )

library(cowplot)
tpc.fig <- plot_grid( tpc.plot, ex1.plot, ex2.plot, ncol = 1 )

ggsave("tpcfig.png", plot = tpc.fig, dpi = 300)
```

