---
title: "Grant Proposal Fig. 4"
author: "Andy"
date: "2025-06-30"
output: pdf_document
---

This document contains an attempt to recreate Fig. 4 in the grant proposal, with the goal of determining why the percentage of the total predator effect composed of the NCE generally levels off given increasing risk.

```{r, warning = FALSE}
library(ggplot2)
library(dplyr)

# Setting the range of prey displacement rate
trait.vec <- seq(0, 1, by = 0.01)
speed.vec <- 1 - trait.vec

# These formulas are simplified by setting some of the values as constants
# r is always assumed to be 1, while k, s, and v are manipulated values
pred.curve <- function(k, v, s){
  k*sqrt(v^2 + s^2)
}

# R is always assumed to be 1, while s and h are manipulated values
growth.curve <- function(s, R, h){
  (R * s)/(1 + s*R*h)
}

k_vec <- c(0, 0.1, 0.13, 0.2, 0.3, 0.4, 0.6, 1) # Range of values used by Scott for Fig. 4 in the grant proposal
h_val <- 8
R_val <- 1
cost_constant_val <- 9
v_val <- 0.2
pred_constant_val <- 0.98058

# Generate growth and mortality curves for each value of predator density (k_vec)
fig4.df <- as.data.frame( mat.or.vec( length( trait.vec ) * length(k_vec), 7) )
colnames(fig4.df) <- c("trait", "speed", "growth_curve", "mort_curve", "fitness", "pct_NCE", "pred_dens")

row <- 0

for (i in 1:length(k_vec)){
  for (j in 1:length(trait.vec)){
    row <- row + 1
    fig4.df[row, 1] <- trait.vec[j]
    fig4.df[row, 2] <- speed.vec[j]
    fig4.df[row, 3] <- cost_constant_val * growth.curve( s = fig4.df[row, 2], R = R_val, h = h_val )
    fig4.df[row, 4] <- pred_constant_val * pred.curve( k = k_vec[i], v = v_val, s = fig4.df[row, 2] )
    fig4.df[row, 5] <- fig4.df[row, 3] - fig4.df[row, 4]
    fig4.df[row, 6] <- (1 - fig4.df[row, 3])/((1 - fig4.df[row, 3]) + fig4.df[row, 4])
    fig4.df[row, 7] <- k_vec[i]
  }
}


# Summarize data by finding optimum fitness for each value of predator density
optim.df <- fig4.df %>%
  group_by( pred_dens ) %>%
  slice( which.max(fitness) )

fig4.df$pred_dens <- as.factor(fig4.df$pred_dens)
optim.df[1, 6] <- 0 # Removing an NA created by dividing by zero where pred_dens = 0

# Plot IREF curves by predator density
ggplot( data = fig4.df, aes( x = trait ) ) +
  geom_line( aes( y = growth_curve ), color = "blue" ) +
  geom_line( aes( y = mort_curve ), color = "red" ) +
  geom_line( aes( y = fitness ), color = "green" ) +
  theme_classic( base_size = 10, base_line_size = 1 ) + 
  ylab("Growth and Mortality") +
  xlab("Displacement Rate") +
  scale_x_continuous(expand = c(0, 0), breaks = c(0, 0.5, 1), labels = c(1000, 500, 1) ) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1), breaks = c(0, 0.5, 1) ) +
  facet_wrap(~pred_dens) +
  theme( panel.border = element_rect( color = "black", fill = NA, linewidth = 0.75 ), aspect.ratio = 1 )

# Plot % contribution of NCE to total predator effect by predator density
ggplot( data = optim.df, aes( x = pred_dens ) ) +
  geom_point( aes( y = pct_NCE ) ) +
  theme_classic( base_size = 10, base_line_size = 1 ) + 
  ylab("% NCE contribution to") +
  xlab("Predator Density") +
  theme( panel.border = element_rect( color = "black", fill = NA, linewidth = 0.75 ), aspect.ratio = 1)
```

The reason that the NCE's contribution to the total predator effect tends to level off with increasing risk seems to occur for two main reasons. 

The optimal RITR seems to occur roughly around the inflection point of the growth curve, except when predator density (and therefore risk) is low (k > 0.2) and a weak response is expected by the prey.

Put another way, as risk increases, there is eventually a point at which the benefit of the RITR (in terms of reduction of mortality due to predation) becomes too costly. Quantitatively, this is probably where the slope (or derivative) of the growth curve exceeds -1 (or, the increase in cost for a 1 unit decrease in displacement rate exceeds greatly exceeds the decrease in mortality afforded by it).

This finding is probably general in cases where the growth curve is concave-down, while predation risk declines in a linear or saturating way.

Incidentally, the inflection point is at around a displacement rate of 375 (calculated below using the R package 'inflection').

```{r, warning=FALSE}
library(inflection)

x <- speed.vec
y <- cost_constant_val * growth.curve( s = speed.vec, R = R_val, h = h_val )
cc=check_curve(x,y);cc
ede(x, y, cc$index)
```

Note that this is the trait value, so to get the displacement rate you must subtract it from 1, then multiply by 1000.