---
title: "IREF - Theory Case Studies"
author: "Andy Davidson"
date: "2025-09-29"
output: pdf_document
---

```{r, include = FALSE}
library(ggplot2)
library(cowplot)
library(ggbrace)
library(knitr)
knitr::opts_chunk$set(tidy.opts = list(width.cutoff = 60), tidy = TRUE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

# Prey Displacement Rate

## Part 1: Predator Encounter Rate

For part I, the encounter rate equation is:

$$ Z = 2rN*sqrt(v^2 + s^2)$$

where $s$ is prey speed (i.e., displacement rate), $v$ is predator speed, $N$ is predator density, and $r$ is the radius within which the predator can perceive prey.

Scott simulates this equation for varying levels of $s$, and four constant levels of $v$ (0, 0.5, 1, and 2). Predator density and detection radius are kept constant such that $2rN = 1$.

$v = 0$ would simply mean a sit-and-wait predator.


## Part 2: Cost vs. Gain for Displacement Rate

Scott mentions incorporating a Type II functional response, such that higher body velocity has diminishing returns due to time spent handling food.

$$ f(R)= \frac{sR}{1 + shR}$$

where $s$ is prey displacement rate, $h$ is handling time, and $R$ is resource density.


Scott describes the fitness curve (without predation, so the growth curve) using the equation from Part 2, and predation using the equation from Part 1.

```{r Displacement Trait Curve Functions}
## Important formulae for calculating displacement rate's growth and mortality curves

# Generating the baseline curves
trait.vec <- seq(0, 1, by = 0.01)
speed.vec <- 1 - trait.vec


# These formulas are simplified by setting some of the values as constants
# r is always assumed to be 1, while k, s, and v are manipulated values
pred.curve <- function(k, v, s){
  k*sqrt(v^2 + s^2)
}

# R is always assumed to be 1, while s and h are manipulated values
growth.curve <- function(s, R, h){
  (R * s)/(1 + s*R*h)
}
```


```{r Displacement Case Studies}
### Combine the two to generate IREF figures, with displacement rate as trait

## 1) This is the baseline scenario
baseline.df <- as.data.frame( mat.or.vec( length( speed.vec ), 7) )
colnames(baseline.df) <- c("trait", "speed", "growth_curve", "mort_curve", "fitness", "NCE_CE", "mu_g")
baseline.df$trait <- trait.vec
baseline.df$speed <- speed.vec
h_val <- 8
R_val <- 1
cost_constant_val <- 9
v_val <- 0.2
k_val <- 0.45
pred_constant_val <- 0.98058

baseline.df$growth_curve <- cost_constant_val * growth.curve( s = baseline.df$speed, R = R_val, h = h_val ) # Growth curve
baseline.df$mort_curve <- pred_constant_val * pred.curve( k = k_val, v = v_val, s = baseline.df$speed ) # Mortality curve
baseline.df$fitness <- baseline.df$growth_curve - baseline.df$mort_curve # fitness
baseline.df$NCE_CE <- (1 - baseline.df$growth_curve)/baseline.df$mort_curve # NCE/CE
baseline.df$mu_g <- baseline.df$mort_curve/baseline.df$growth_curve # mu/g



## 2) Change predator density by increasing k
preddens.df <- as.data.frame( mat.or.vec( length( speed.vec ), 7) )
colnames(preddens.df) <- c("trait", "speed", "growth_curve", "mort_curve", "fitness", "NCE_CE", "mu_g")
preddens.df$trait <- trait.vec
preddens.df$speed <- speed.vec
k_val <- 0.9 

preddens.df$growth_curve <- cost_constant_val * growth.curve( s = preddens.df$speed, R = R_val, h = h_val ) # Growth curve
preddens.df$mort_curve <- pred_constant_val * pred.curve( k = k_val, v = v_val, s = preddens.df$speed ) # Mortality curve
preddens.df$fitness <- preddens.df$growth_curve - preddens.df$mort_curve # fitness
preddens.df$NCE_CE <- (1 - preddens.df$growth_curve)/preddens.df$mort_curve # NCE/CE
preddens.df$mu_g <- preddens.df$mort_curve/preddens.df$growth_curve # mu/g



## 3) Change predator foraging mode from a sit-and-wait to active predator by increasing v
predmode.df <- as.data.frame( mat.or.vec( length( speed.vec ), 7) )
colnames(predmode.df) <- c("trait", "speed", "growth_curve", "mort_curve", "fitness", "NCE_CE", "mu_g")
predmode.df$trait <- trait.vec
predmode.df$speed <- speed.vec
v_val <- 1
pred_constant_val <- 0.7071063


predmode.df$growth_curve <- cost_constant_val * growth.curve( s = predmode.df$speed, R = R_val, h = h_val ) # Growth curve
predmode.df$mort_curve <- pred_constant_val * pred.curve( k = k_val, v = v_val, s = predmode.df$speed ) # Mortality curve
predmode.df$fitness <- predmode.df$growth_curve - predmode.df$mort_curve # fitness
predmode.df$NCE_CE <- (1 - predmode.df$growth_curve)/predmode.df$mort_curve # NCE/CE
predmode.df$mu_g <- predmode.df$mort_curve/predmode.df$growth_curve # mu/g



## 4) Calculate summary stats for each data set, inc. optimum, fitness, CE and NCE strength, mu/g, etc
displacement_summary.df <- as.data.frame( mat.or.vec( 3, 8) )
colnames(displacement_summary.df) <- c("case", "v", "k", "optim", "fitness", "CE", "NCE", "optim_mu_g")

displacement_summary.df$case <- c("base", "preddens", "predmode")
displacement_summary.df$v <- c("0.2", "0.2", "1")
displacement_summary.df$k <- c("0.45", "0.9", "0.9")

#For base scenario:
displacement_summary.df[1, 4] <- 
  baseline.df$trait[which(baseline.df$fitness == max(baseline.df$fitness))] # Optimum under 1 - NCE - CE
displacement_summary.df[1, 5] <- 
  baseline.df$fitness[which(baseline.df$fitness == max(baseline.df$fitness))] # Fitness at optimum
displacement_summary.df[1, 6] <- 
  baseline.df$mort_curve[which(baseline.df$fitness == max(baseline.df$fitness))] # CE strength at optimum
displacement_summary.df[1, 7] <- 
  1 - displacement_summary.df[1, 6] - displacement_summary.df[1, 5] # NCE strength at optimum
displacement_summary.df[1, 8] <- 
  baseline.df$trait[which(baseline.df$mu_g == min(baseline.df$mu_g))] # Optimum using mu/g

#For predator density scenario:
displacement_summary.df[2, 4] <- 
  preddens.df$trait[which(preddens.df$fitness == max(preddens.df$fitness))] # Optimum under 1 - NCE - CE
displacement_summary.df[2, 5] <- 
  preddens.df$fitness[which(preddens.df$fitness == max(preddens.df$fitness))] # Fitness at optimum
displacement_summary.df[2, 6] <- 
  preddens.df$mort_curve[which(preddens.df$fitness == max(preddens.df$fitness))] # CE strength at optimum
displacement_summary.df[2, 7] <- 
  1 - displacement_summary.df[2, 6] - displacement_summary.df[2, 5] # NCE strength at optimum
displacement_summary.df[2, 8] <- 
  preddens.df$trait[which(preddens.df$mu_g == min(preddens.df$mu_g))] # Optimum using mu/g

#For predator mode scenario:
displacement_summary.df[3, 4] <- 
  predmode.df$trait[which(predmode.df$fitness == max(predmode.df$fitness))] # Optimum under 1 - NCE - CE
displacement_summary.df[3, 5] <- 
  predmode.df$fitness[which(predmode.df$fitness == max(predmode.df$fitness))] # Fitness at optimum
displacement_summary.df[3, 6] <- 
  predmode.df$mort_curve[which(predmode.df$fitness == max(predmode.df$fitness))] # CE strength at optimum
displacement_summary.df[3, 7] <-  
  1 - displacement_summary.df[3, 6] - displacement_summary.df[3, 5] # NCE strength at optimum
displacement_summary.df[3, 8] <- 
  predmode.df$trait[which(predmode.df$mu_g == min(predmode.df$mu_g))] # Optimum using mu/g



## 5) Calculate change in NCE % of total predator effect, and optimal RITR, with increasing k and v:
# For predator density:
k_vec <- seq( 0, 2.5, by = 0.05 )
preddens_pctNCE.df <- as.data.frame(mat.or.vec( length(k_vec), 4 ) ) 
colnames(preddens_pctNCE.df) <- c("k", "NCE_TPE", "OptimalRITR", "Fitness")

for (i in 1:length( k_vec ) ){
preddens_pctNCE.df[i, 1] <- k_vec[i]

pd_growth_curve <- cost_constant_val * growth.curve( s = speed.vec, R = 1, h = 8 ) # Growth curve
pd_mort_curve <- pred_constant_val * pred.curve( k = k_vec[i], v = 0.2, s = speed.vec ) # Mortality curve
pd_fitness <- pd_growth_curve - pd_mort_curve # fitness
pd_NCE_TPE <- (1 - pd_growth_curve)/(pd_mort_curve + (1 - pd_growth_curve))# NCE/TPE

preddens_pctNCE.df[i, 2] <- pd_NCE_TPE[ which.max(pd_fitness) ]

# Some other important metrics
preddens_pctNCE.df[i, 3] <- speed.vec[ which.max(pd_fitness) ]
preddens_pctNCE.df[i, 4] <- max(pd_fitness)
}


# Repeat for mu/g
k_vec <- seq( 0, 2.5, by = 0.05 )
preddens_pctNCE_mug.df <- as.data.frame(mat.or.vec( length(k_vec), 4 ) ) 
colnames(preddens_pctNCE_mug.df) <- c("k", "NCE_TPE", "OptimalRITR", "Fitness")

for (i in 1:length( k_vec ) ){
preddens_pctNCE_mug.df[i, 1] <- k_vec[i]

pd_growth_curve <- cost_constant_val * growth.curve( s = speed.vec, R = 1, h = 8 ) # Growth curve
pd_mort_curve <- pred_constant_val * pred.curve( k = k_vec[i], v = 0.2, s = speed.vec ) # Mortality curve
pd_fitness <- pd_mort_curve / pd_growth_curve # fitness
pd_NCE_TPE <- (1 - pd_growth_curve)/(pd_mort_curve + (1 - pd_growth_curve))# NCE/TPE

preddens_pctNCE_mug.df[i, 2] <- pd_NCE_TPE[ which.min(pd_fitness) ]

# Some other important metrics
preddens_pctNCE_mug.df[i, 3] <- speed.vec[ which.min(pd_fitness) ]
preddens_pctNCE_mug.df[i, 4] <- min(pd_fitness)
}
# This is completely flat



# For predator foraging mode:
v_vec <- seq( 0, 2.5, by = 0.05 )
predmode_pctNCE.df <- as.data.frame(mat.or.vec( length(v_vec), 4 ) )
pred_constant_val <- 0.7071068 * 0.8999994 # this changes to anchor the mortality curve in the same spot
colnames(predmode_pctNCE.df) <- c("v", "NCE_TPE", "OptimalRITR", "Fitness")

for (i in 1:length( v_vec ) ){
predmode_pctNCE.df[i, 1] <- v_vec[i]

pm_growth_curve <- cost_constant_val * growth.curve( s = speed.vec, R = 1, h = 8 ) # Growth curve
pm_mort_curve <- pred_constant_val * pred.curve( k = 1, v = v_vec[i], s = speed.vec ) # Mortality curve
pm_fitness <- pm_growth_curve - pm_mort_curve # fitness
pm_NCE_TPE <- (1 - pm_growth_curve)/(pm_mort_curve + (1 - pm_growth_curve))# NCE/TPE

predmode_pctNCE.df[i, 2] <- pm_NCE_TPE[ which.max(pm_fitness) ]

# Some other important metrics
predmode_pctNCE.df[i, 3] <- speed.vec[ which.max(pm_fitness) ]
predmode_pctNCE.df[i, 4] <- max(pm_fitness)
}

# Repeat for mu/g
v_vec <- seq( 0, 2.5, by = 0.05 )
predmode_pctNCE_mug.df <- as.data.frame(mat.or.vec( length(v_vec), 4 ) )
pred_constant_val <- 0.7071068 * 0.8999994 # this changes to anchor the mortality curve in the same spot
colnames(predmode_pctNCE_mug.df) <- c("v", "NCE_TPE", "OptimalRITR", "Fitness")

for (i in 1:length( v_vec ) ){
predmode_pctNCE_mug.df[i, 1] <- v_vec[i]

pm_growth_curve <- cost_constant_val * growth.curve( s = speed.vec, R = 1, h = 8 ) # Growth curve
pm_mort_curve <- pred_constant_val * pred.curve( k = 1, v = v_vec[i], s = speed.vec ) # Mortality curve
pm_fitness <-  pm_mort_curve / pm_growth_curve # fitness
pm_NCE_TPE <- (1 - pm_growth_curve)/(pm_mort_curve + (1 - pm_growth_curve))# NCE/TPE

predmode_pctNCE_mug.df[i, 2] <- pm_NCE_TPE[ which.min(pm_fitness) ]

# Some other important metrics
predmode_pctNCE_mug.df[i, 3] <- speed.vec[ which.min(pm_fitness) ]
predmode_pctNCE_mug.df[i, 4] <- min(pm_fitness)
}

```


```{r Displacement Data Export, eval = FALSE}
# Export the six data sets
write.csv(baseline.df, "1-1_displacement_baseline.csv.csv", row.names = FALSE)
write.csv(preddens.df, "1-2-1_displacement_preddens.csv", row.names = FALSE)
write.csv(preddens_pctNCE.df, "1-2-2_preddens_TPE_Opt.csv", row.names = FALSE)
write.csv(predmode.df, "1-3-1_displacement_predmode.csv", row.names = FALSE)
write.csv(predmode_pctNCE.df, "1-3-2_predmode_TPE_Opt.csv", row.names = FALSE)
write.csv(displacement_summary.df, "1-4_displacement_summary.csv", row.names = FALSE)
write.csv(preddens_pctNCE_mug.df, "1-5-1_preddens_mug.csv", row.names = FALSE)
write.csv(predmode_pctNCE_mug.df, "1-5-2_predmode_mug.csv", row.names = FALSE)
```


```{r Refuge Use Case Studies, pt. 1}
### First scenario - resource poor refuge
proptime_vec <- seq(1, 0, by = -0.01)

## 1) Baseline scenario
refuge.df <- as.data.frame( mat.or.vec( length(proptime_vec), 7 ) ) 
colnames(refuge.df) <- c("proptime", "resource_dens", "growth_curve", "mort_curve", "fitness", "NCE_CE", "mu_g")

# Growth follows a Type II FR, where instead of speed we're manipulating resource density
refuge.df$proptime <- proptime_vec
refuge.df$resource_dens <- seq(0, 1, by = 0.01) # This needs to be the same length as proptime_vec
refuge.df$growth_curve <- growth.curve( s = 1, R = refuge.df$resource_dens, h = 10 )
refuge.df$growth_curve <- refuge.df$growth_curve * ( 1 / max( refuge.df$growth_curve ) ) # Anchor at 1

# Mortality curve, which linearly declines from val_A to val_B
mort_val_A <- 0.4 # Mortality outside the refuge during an hour
mort_val_B <- 0 # Mortality inside the refuge during an hour
refuge.df$mort_curve <- ( 1 - refuge.df$proptime ) * mort_val_A + ( refuge.df$proptime ) * mort_val_B

# Calculate fitness and relative strength of NCEs and CEs
refuge.df$fitness <- 1 - ( 1 - refuge.df$growth_curve ) - refuge.df$mort_curve
refuge.df$NCE_CE <- (1 - refuge.df$growth_curve) / refuge.df$mort_curve
refuge.df$mu_g <- refuge.df$mort_curve / refuge.df$growth_curve



## 2) Decrease handling time, i.e., increase resource processing rate
refuge2.df <- as.data.frame( mat.or.vec( length(proptime_vec), 7 ) ) 
colnames(refuge2.df) <- c("proptime", "resource_dens", "growth_curve", "mort_curve", "fitness", "NCE_CE", "mu_g")

# Growth follows a Type II FR, where instead of speed we're manipulating resource density
refuge2.df$proptime <- proptime_vec 
refuge2.df$resource_dens <- seq(0, 1, by = 0.01) # This needs to be the same length as proptime_vec
refuge2.df$growth_curve <- growth.curve(s = 1, R = refuge2.df$resource_dens, h = 2)
refuge2.df$growth_curve <- refuge2.df$growth_curve * ( 1 / max( refuge2.df$growth_curve ) ) # Anchor at 1

# Mortality curve, which linearly declines from val_A to val_B
refuge2.df$mort_curve <- ( 1 - refuge2.df$proptime ) * mort_val_A + ( refuge2.df$proptime) * mort_val_B

# Calculate fitness and relative strength of NCEs and CEs
refuge2.df$fitness <- 1 - ( 1 - refuge2.df$growth_curve ) - refuge2.df$mort_curve
refuge2.df$NCE_CE <- ( 1 - refuge2.df$growth_curve ) / refuge2.df$mort_curve
refuge2.df$mu_g <- refuge2.df$mort_curve / refuge2.df$growth_curve

# Replace NaNs with zeroes
refuge.df[ 1, 7] <- 0
refuge2.df[ 1, 7] <- 0



## 3) Calculate summary stats for each data set, inc. optimum, fitness, CE and NCE strength, mu/g, etc
refuge_res_summary.df <- as.data.frame( mat.or.vec( 2, 9 ) )
colnames(refuge_res_summary.df) <- c("case", "h_val", "P_A", "P_B", "optim", "fitness", "CE", "NCE", "optim_mu_g")

refuge_res_summary.df$case <- c( "base", "fast_h" )
refuge_res_summary.df$h_val <- c( "10", "2" )
refuge_res_summary.df$P_A <- c( "0.4", "0.4" )
refuge_res_summary.df$P_B <- c( "0", "0" )

# For baseline scenario
refuge_res_summary.df[ 1, 5 ] <- 
  refuge.df$proptime[which(refuge.df$fitness == max(refuge.df$fitness))] # Optimum under 1 - NCE - CE
refuge_res_summary.df[ 1, 6 ] <- 
  refuge.df$fitness[which(refuge.df$fitness == max(refuge.df$fitness))] # Fitness at optimum
refuge_res_summary.df[ 1, 7 ] <- 
  refuge.df$mort_curve[which(refuge.df$fitness == max(refuge.df$fitness))] # CE strength at optimum
refuge_res_summary.df[ 1, 8 ] <- 
  1 - refuge_res_summary.df[ 1, 7 ] - refuge_res_summary.df[ 1, 6 ] # NCE strength at optimum
refuge_res_summary.df[ 1, 9 ] <- 
    refuge.df$proptime[which(refuge.df$mu_g == min(refuge.df$mu_g))] # Optimum using mu/g

# For fast h scenario
refuge_res_summary.df[ 2, 5 ] <- 
  refuge2.df$proptime[which(refuge2.df$fitness == max(refuge2.df$fitness))] # Optimum under 1 - NCE - CE
refuge_res_summary.df[ 2, 6 ] <- 
  refuge2.df$fitness[which(refuge2.df$fitness == max(refuge2.df$fitness))] # Fitness at optimum
refuge_res_summary.df[ 2, 7 ] <- 
  refuge2.df$mort_curve[which(refuge2.df$fitness == max(refuge2.df$fitness))] # CE strength at optimum
refuge_res_summary.df[ 2, 8 ] <- 
  1 - refuge_res_summary.df[ 2, 7 ] - refuge_res_summary.df[ 1, 6 ] # NCE strength at optimum
refuge_res_summary.df[ 2, 9 ] <- 
    refuge.df$proptime[which(refuge2.df$mu_g == min(refuge2.df$mu_g))] # Optimum using mu/g



## 4) Calculate % NCE of TPE and optimal RITR
# Calculate change in NCE % of TPE, and optimal RITR, with decreasing h:
h_vec <- seq( 15, 1.5, by = -0.1 )
prey_h_pctNCE.df <- as.data.frame(mat.or.vec( length(h_vec), 5 ) )
colnames(prey_h_pctNCE.df) <- c("h", "NCE_TPE", "OptimalRITR", "Fitness", "inverse_h")

for (i in 1:length( h_vec ) ){
prey_h_pctNCE.df[i, 1] <- h_vec[i]

# Growth follows a Type II FR, where instead of speed we're manipulating resource density
resource_vec <- seq(0, 1, by = 0.01) # This needs to be the same length as proptime_vec, here the refuge has ZERO resources
ph_growth_curve <- growth.curve(s = 1, R = resource_vec, h = h_vec[i])
ph_growth_curve <- ph_growth_curve * (1/max(ph_growth_curve)) # Anchor at 1

# Mortality curve, which linearly declines from val_A to val_B
mort_val_A <- 0.4 # Mortality outside the refuge during an hour
mort_val_B <- 0 # Mortality inside the refuge during an hour
ph_mort_curve <- ( 1 - refuge.df$proptime ) * mort_val_A + ( refuge.df$proptime) * mort_val_B


ph_fitness <- ph_growth_curve - ph_mort_curve # fitness
ph_NCE_TPE <- (1 - ph_growth_curve)/(ph_mort_curve + (1 - ph_growth_curve))# NCE/TPE

prey_h_pctNCE.df[i, 2] <- ph_NCE_TPE[ which.max(ph_fitness) ]

# Some other important metrics
prey_h_pctNCE.df[i, 3] <- proptime_vec[ which.max(ph_fitness) ]
prey_h_pctNCE.df[i, 4] <- max(ph_fitness)
}

prey_h_pctNCE.df$inverse_h <- 1/prey_h_pctNCE.df$h
```


```{r Refuge Use Resource Data Export, eval = FALSE}
# Export the four data sets
write.csv(refuge.df, "2-1_refuge_res_base.csv", row.names = FALSE)
write.csv(refuge2.df, "2-2_refuge_res_fast_h.csv", row.names = FALSE)
write.csv(prey_h_pctNCE.df, "2-3_refuge_res_TPE_Opt.csv", row.names = FALSE)
write.csv(refuge_res_summary.df, "2-4_refuge_res_summary.csv", row.names = FALSE)
```


```{r Refuge Use Case Studies, pt. 2}
## Derive the trait curves
therm_refuge.df <- as.data.frame( mat.or.vec( length(proptime_vec), 6 ) ) 
colnames(therm_refuge.df) <- c("proptime", "growth_curve", "mort_curve", "fitness", "NCE_CE", "mu_g")
therm_refuge.df$proptime <- rev(proptime_vec)


## The growth-trait curve
# a.growth determines the width of the performance curve and should be negative, so that the curve is concave down
# h.growth determines the temperature at which prey growth peaks
# c.growth is a scaling constant
# temp_C is temperature in degrees Celsius
growth.curve_warming <- function( c_g, a_g, h_g, temp_C ){
  c_g * exp( a_g * ( ( temp_C - h_g )^2 ) + 1)
}

a_g.val <- -0.05 # Set a.growth to an arbitrary value that makes the curve broader than the predator's distribution; smaller values yield broader curves
h_g.val <- 20 # Set h.growth to the same point that the predator's density peaks
c_g.val <- 1/exp(1) # c.growth needs to be 1/e for this parameterization to peak at 1


## 1) Base Scenario
# The growth curve will follow a temperature gradient along the prey's TPC
temp_C_vec <- seq(20, 23, by = 0.03) # this has to be the same length as proptime
therm_refuge.df$growth_curve <- (1 - therm_refuge.df$proptime) * growth.curve_warming( c_g = c_g.val, a_g = a_g.val, h_g = h_g.val, temp_C = 19 ) + (therm_refuge.df$proptime) * growth.curve_warming( c_g = c_g.val, a_g = a_g.val, h_g = h_g.val, temp_C = 22 )

# Mortality curve
therm_refuge.df$mort_curve <- ( 1 - therm_refuge.df$proptime ) * mort_val_A + ( therm_refuge.df$proptime) * mort_val_B

# Calculate fitness and relative strength of NCEs and CEs
therm_refuge.df$fitness <- 1 - ( 1 - therm_refuge.df$growth_curve ) - therm_refuge.df$mort_curve
therm_refuge.df$NCE_CE <- therm_refuge.df$growth_curve / therm_refuge.df$mort_curve
therm_refuge.df$mu_g <- therm_refuge.df$mort_curve / therm_refuge.df$growth_curve



## 2) Warmed Scenario
therm_refuge2.df <- as.data.frame( mat.or.vec( length(proptime_vec), 6 ) ) 
colnames(therm_refuge2.df) <- c("proptime", "growth_curve", "mort_curve", "fitness", "NCE_CE", "mu_g")
therm_refuge2.df$proptime <- rev( proptime_vec )

# The growth curve will follow a temperature gradient along the prey's TPC
temp_C_vec <- seq(21.5, 24.5, by = 0.03) # this has to be the same length as proptime
therm_refuge2.df$growth_curve <- (1 - therm_refuge.df$proptime) * growth.curve_warming( c_g = c_g.val, a_g = a_g.val, h_g = h_g.val, temp_C = 21.5 ) + (therm_refuge.df$proptime) * growth.curve_warming( c_g = c_g.val, a_g = a_g.val, h_g = h_g.val, temp_C = 24.5 )
therm_refuge2.df$growth_curve <- therm_refuge2.df$growth_curve * (1/ max(therm_refuge2.df$growth_curve) )

# Weighted mean of risk during time spent outside refuge + risk during time spent inside refuge
therm_refuge2.df$mort_curve <- ( 1 - therm_refuge2.df$proptime ) * mort_val_A + ( therm_refuge2.df$proptime ) * mort_val_B 

# Calculate fitness and relative strength of NCEs and CEs
therm_refuge2.df$fitness <- 1 - ( 1 - therm_refuge2.df$growth_curve ) - therm_refuge2.df$mort_curve
therm_refuge2.df$NCE_CE <- therm_refuge2.df$growth_curve / therm_refuge2.df$mort_curve
therm_refuge2.df$mu_g <- therm_refuge2.df$mort_curve / therm_refuge2.df$growth_curve



## 3) Calculate summary stats for each data set, inc. optimum, fitness, CE and NCE strength, mu/g, etc
refuge_therm_summary.df <- as.data.frame( mat.or.vec( 2, 10 ) )
colnames(refuge_therm_summary.df) <- c("case", "T_A", "T_B", "P_A", "P_B", "optim", "fitness", "CE", "NCE", "optim_mu_g")

refuge_therm_summary.df$case <- c( "base", "warmed" )
refuge_therm_summary.df$T_A <- c( "20", "21.5" )
refuge_therm_summary.df$T_B <- c( "23", "24.5" )
refuge_therm_summary.df$P_A <- c( "0.4", "0.4" )
refuge_therm_summary.df$P_B <- c( "0", "0" )

refuge_therm_summary.df[ 1, 6 ] <- 
  therm_refuge.df$proptime[which(therm_refuge.df$fitness == max(therm_refuge.df$fitness))] # Optimum under 1 - NCE - CE
refuge_therm_summary.df[ 1, 7 ] <- 
  max ( therm_refuge.df$fitness ) # Fitness at optimum
refuge_therm_summary.df[ 1, 8 ] <- 
  therm_refuge.df$mort_curve[which(therm_refuge.df$fitness == max(therm_refuge.df$fitness))] # CE strength at optimum
refuge_therm_summary.df[ 1, 9 ] <-   
  1 - refuge_therm_summary.df[ 1, 8 ] - refuge_therm_summary.df[ 1, 7 ] # NCE strength at optimum
refuge_therm_summary.df[ 1, 10 ] <- 
    therm_refuge.df$proptime[which(therm_refuge.df$mu_g == min(therm_refuge.df$mu_g))] # Optimum using mu/g

refuge_therm_summary.df[ 2, 6 ] <- 
  therm_refuge2.df$proptime[which(therm_refuge2.df$fitness == max(therm_refuge2.df$fitness))] # Optimum under 1 - NCE - CE
refuge_therm_summary.df[ 2, 7 ] <- 
  max ( therm_refuge2.df$fitness ) # Fitness at optimum
refuge_therm_summary.df[ 2, 8 ] <- 
  therm_refuge2.df$mort_curve[which(therm_refuge2.df$fitness == max(therm_refuge2.df$fitness))] # CE strength at optimum
refuge_therm_summary.df[ 2, 9 ] <-   
  1 - refuge_therm_summary.df[ 2, 8 ] - refuge_therm_summary.df[ 2, 7 ] # NCE strength at optimum
refuge_therm_summary.df[ 2, 10 ] <- 
    therm_refuge2.df$proptime[which(therm_refuge2.df$mu_g == min(therm_refuge2.df$mu_g))] # Optimum using mu/g
```

```{r Refuge Use Thermal Data Export, eval = FALSE}
# Export the three data sets
write.csv(therm_refuge.df, "3-1_refuge_therm_base.csv", row.names = FALSE)
write.csv(therm_refuge2.df, "3-2_refuge_therm_warmed.csv", row.names = FALSE)
write.csv(refuge_therm_summary.df, "3-3_refuge_therm_summary.csv", row.names = FALSE)
```