---
title: 'Trait: Refuge Use'
author: "Andy Davidson"
date: "2025-12-08"
output: pdf_document
---

```{r Setup Config}
library(ggbrace)
library(knitr)
knitr::opts_chunk$set(tidy.opts = list(width.cutoff = 60), tidy = TRUE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

# Trait: Refuge Use

Here, we use the example of prey that make use of a safe but resource-absent refuge habitat to avoid predation. We describe the amount of time prey spend in the refuge using the term $\delta$. We use it to illustrate how different modeling assumptions can dramatically change the conclusions we draw using IREF.

## 1. Predation Curve

As before, we assume that predators exhibit a Type I functional response, and thus the amount of predation that prey experience will be a linear function of the predator's functional response in both habitats. We simplify this here by assuming that this translates to a mortality probability in Habitat A (the main habitat) and a different, lower mortality probability in Habitat B (the refuge habitat).


## 2. Growth Curve

We present two ways in which the growth curve can be derived. In the first example, the curve is calculated as above: as the linear sum of the prey's functional response (assumed to be Type II) in Habitats A and B.

In the second example, we assume that prey may exhibit satiation, and thus they will experience diminishing returns on growth with further time spent foraging. We approximate this using a curvilinear function that is mathematically similar to a Type II functional response.

```{r Growth Curve}
## The Growth Curve
# R is always assumed to be 1, while s and h are manipulated values
growth.curve <- function(s, R, h){
  (R * s)/(1 + s*R*h)
}
```

# Case 1: Linear Growth

When both growth and predation are linearly related to $\delta$, the optimal RITR is either to spend all of the time in the refuge, or none of it.

```{r Linear Growth, No Refuge Use}
delta_vec <- seq(1, 0, by = -0.01) # Proportion of time spent in Habitat B
linear_growth_noRef.df <- as.data.frame( mat.or.vec( length(delta_vec), 5 ) ) 
colnames(linear_growth_noRef.df) <- c("delta", "growth_curve", "pred_curve", "fitness", "NCE_CE")

# Growth is the sum of the functional response in A and B
linear_growth_noRef.df$delta <- delta_vec
cost_constant_val <- 2
linear_growth_noRef.df$growth_curve <- (1 - linear_growth_noRef.df$delta) * cost_constant_val * growth.curve( s = 1, R = 1, h = 1 ) + ( linear_growth_noRef.df$delta ) * growth.curve( s = 1, R = 0.1, h = 5 )

# Mortality is the sum of risk in A and B
mort_val_A <- 0.5 # Mortality outside the refuge during an hour
mort_val_B <- 0 # Mortality inside the refuge during an hour
linear_growth_noRef.df$pred_curve <- ( 1 - linear_growth_noRef.df$delta ) * mort_val_A + ( linear_growth_noRef.df$delta ) * mort_val_B

# Calculate fitness and relative strength of NCEs and CEs
linear_growth_noRef.df$fitness <- 1 - ( 1 - linear_growth_noRef.df$growth_curve ) - linear_growth_noRef.df$pred_curve
linear_growth_noRef.df$NCE_CE <- (1 - linear_growth_noRef.df$growth_curve) / linear_growth_noRef.df$pred_curve
```

```{r Linear Growth, No Refuge Use}
linear_growth_allRef.df <- as.data.frame( mat.or.vec( length(delta_vec), 5 ) ) 
colnames(linear_growth_allRef.df) <- c("delta", "growth_curve", "pred_curve", "fitness", "NCE_CE")

# Growth is the sum of the functional response in A and B
linear_growth_allRef.df$delta <- delta_vec
linear_growth_allRef.df$growth_curve <- (1 - linear_growth_allRef.df$delta) * cost_constant_val * growth.curve( s = 1, R = 1, h = 1 ) + ( linear_growth_allRef.df$delta ) * growth.curve( s = 1, R = 0.1, h = 5 )

# Mortality is the sum of risk in A and B
mort_val_A <- 1 # Mortality outside the refuge during an hour
mort_val_B <- 0 # Mortality inside the refuge during an hour
linear_growth_allRef.df$pred_curve <- ( 1 - linear_growth_allRef.df$delta ) * mort_val_A + ( linear_growth_allRef.df$delta ) * mort_val_B

# Calculate fitness and relative strength of NCEs and CEs
linear_growth_allRef.df$fitness <- 1 - ( 1 - linear_growth_allRef.df$growth_curve ) - linear_growth_allRef.df$pred_curve
linear_growth_allRef.df$NCE_CE <- (1 - linear_growth_allRef.df$growth_curve) / linear_growth_allRef.df$pred_curve
```

# Case 2: Diminishing Returns Due to Satiation

If instead, growth is nonlinearly related to $\delta$, intermediate optimal RITRs are possible.

```{r With Satiation}
nonlinear.df <- as.data.frame( mat.or.vec( length(delta_vec), 5 ) ) 
colnames(nonlinear.df) <- c("delta", "growth_curve", "pred_curve", "fitness", "NCE_CE")

# Assume no resources, no predators in refuge
# Growth follows a Type II FR, where instead of speed we're manipulating resource density
nonlinear.df$delta <- delta_vec
nonlinear.df$growth_curve <- growth.curve( s = 1, R = 1 - nonlinear.df$delta, h = 5 )
nonlinear.df$growth_curve <- nonlinear.df$growth_curve * (0.9 / 0.16666667) + 0.1 # Anchor at 1

# Mortality curve, which linearly declines from val_A to val_B
mort_val_A <- 1 # Mortality outside the refuge during an hour
mort_val_B <- 0 # Mortality inside the refuge during an hour
nonlinear.df$pred_curve <- ( 1 - nonlinear.df$delta ) * mort_val_A + ( nonlinear.df$delta ) * mort_val_B

# Calculate fitness and relative strength of NCEs and CEs
nonlinear.df$fitness <- 1 - ( 1 - nonlinear.df$growth_curve ) - nonlinear.df$pred_curve
nonlinear.df$NCE_CE <- (1 - nonlinear.df$growth_curve) / nonlinear.df$pred_curve
```

```{r Summary Stats}
## Calculate summary stats for each data set, inc. optimum, fitness, CE and NCE strength,  etc
refuge_summary.df <- as.data.frame( mat.or.vec( 3, 5) )
colnames(refuge_summary.df) <- c("case", "optim", "fitness", "CE", "NCE")

refuge_summary.df$case <- c("NoUse", "AllUse", "Nonlinear")


#For No Use scenario:
refuge_summary.df[1, 2] <- 
  linear_growth_noRef.df$delta[which(linear_growth_noRef.df$fitness == max(linear_growth_noRef.df$fitness))] # Optimum under 1 - NCE - CE
refuge_summary.df[1, 3] <- 
  linear_growth_noRef.df$fitness[which(linear_growth_noRef.df$fitness == max(linear_growth_noRef.df$fitness))] # Fitness at optimum
refuge_summary.df[1, 4] <- 
  linear_growth_noRef.df$pred_curve[which(linear_growth_noRef.df$fitness == max(linear_growth_noRef.df$fitness))] # CE strength at optimum
refuge_summary.df[1, 5] <- 
  1 - refuge_summary.df[1, 4] - refuge_summary.df[1, 3] # NCE strength at optimum

#For All Use scenario:
refuge_summary.df[2, 2] <- 
  linear_growth_allRef.df$delta[which(linear_growth_allRef.df$fitness == max(linear_growth_allRef.df$fitness))] # Optimum under 1 - NCE - CE
refuge_summary.df[2, 3] <- 
  linear_growth_allRef.df$fitness[which(linear_growth_allRef.df$fitness == max(linear_growth_allRef.df$fitness))] # Fitness at optimum
refuge_summary.df[2, 4] <- 
  linear_growth_allRef.df$pred_curve[which(linear_growth_allRef.df$fitness == max(linear_growth_allRef.df$fitness))] # CE strength at optimum
refuge_summary.df[2, 5] <- 
  1 - refuge_summary.df[2, 4] - refuge_summary.df[2, 3] # NCE strength at optimum

#For Nonlinear Growth scenario:
refuge_summary.df[3, 2] <- 
  nonlinear.df$delta[which(nonlinear.df$fitness == max(nonlinear.df$fitness))] # Optimum under 1 - NCE - CE
refuge_summary.df[3, 3] <- 
  nonlinear.df$fitness[which(nonlinear.df$fitness == max(nonlinear.df$fitness))] # Fitness at optimum
refuge_summary.df[3, 4] <- 
  nonlinear.df$pred_curve[which(nonlinear.df$fitness == max(nonlinear.df$fitness))] # CE strength at optimum
refuge_summary.df[3, 5] <- 
  1 - refuge_summary.df[3, 4] - refuge_summary.df[3, 3] # NCE strength at optimum

```


```{r Refuge Use Resource Data Export, eval = FALSE}
# Export the four data sets
write.csv(linear_growth_noRef.df, "2-1_refuge_linear_NoUse.csv", row.names = FALSE)
write.csv(linear_growth_allRef.df, "2-2_refuge_linear_AllUse.csv", row.names = FALSE)
write.csv(nonlinear.df, "2-3_refuge_nonlinear.csv", row.names = FALSE)
write.csv(refuge_summary.df, "2-4_refuge_summary.csv", row.names = FALSE)
```