---
title: "IREF - NSF Grant Scenarios"
author: "Andy Davidson"
date: "2025-01-21"
output: pdf_document
---

```{r, include = FALSE}
library(ggplot2)
library(cowplot)
library(ggbrace)
library(knitr)
knitr::opts_chunk$set(tidy.opts = list(width.cutoff = 60), tidy = TRUE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

# Prey Displacement Rate

## Part 1: Predator Encounter Rate

For part I, the encounter rate equation is:

$$ Z = 2rN*sqrt(v^2 + s^2)$$

where $s$ is prey speed (i.e., displacement rate), $v$ is predator speed, $N$ is predator density, and $r$ is the radius within which the predator can perceive prey.

Scott simulates this equation for varying levels of $s$, and four constant levels of $v$ (0, 0.5, 1, and 2). Predator density and detection radius are kept constant such that $2rN = 1$.

$v = 0$ would simply mean a sit-and-wait predator.


## Part 2: Cost vs. Gain for Displacement Rate

Scott mentions incorporating a Type II functional response, such that higher body velocity has diminishing returns due to time spent handling food.

$$ f(R)= \frac{sR}{1 + shR}$$

where $s$ is prey displacement rate, $h$ is handling time, and $R$ is resource density.


Scott describes the fitness curve (without predation, so the growth curve) using the equation from Part 2, and predation using the equation from Part 1.

```{r}
# Generating the baseline curves
trait.vec <- seq(0, 1, by = 0.01)
speed.vec <- 1 - trait.vec


# These formulas are simplified by setting some of the values as constants
# r is always assumed to be 1, while k, s, and v are manipulated values
pred.curve <- function(k, v, s){
  k*sqrt(v^2 + s^2)
}

# R is always assumed to be 1, while s and h are manipulated values
growth.curve <- function(s, R, h){
  (R * s)/(1 + s*R*h)
}
```


```{r}
# Combine the two to generate IREF figures

# This is the baseline scenario
baseline.df <- as.data.frame( mat.or.vec( length( speed.vec ), 6) )
colnames(baseline.df) <- c("trait", "speed", "growth_curve", "mort_curve", "fitness", "NCE_CE")
baseline.df$trait <- trait.vec
baseline.df$speed <- speed.vec
h_val <- 8
R_val <- 1
cost_constant_val <- 9
v_val <- 0.2
k_val <- 0.45
pred_constant_val <- 0.98058

baseline.df$growth_curve <- cost_constant_val * growth.curve( s = baseline.df$speed, R = R_val, h = h_val ) # Growth curve
baseline.df$mort_curve <- pred_constant_val * pred.curve( k = k_val, v = v_val, s = baseline.df$speed ) # Mortality curve
baseline.df$fitness <- baseline.df$growth_curve - baseline.df$mort_curve # fitness
baseline.df$NCE_CE <- (1 - baseline.df$growth_curve)/baseline.df$mort_curve # NCE/CE

baseline_optim <- baseline.df$trait[which(baseline.df$fitness == max(baseline.df$fitness))]
baseline_optim_fitness <- baseline.df$fitness[which(baseline.df$fitness == max(baseline.df$fitness))]
baseline_optim_CE <- baseline.df$mort_curve[which(baseline.df$fitness == max(baseline.df$fitness))]
baseline_optim_NCE <- 1 - baseline_optim_CE - baseline_optim_fitness
```


```{r}
# Grant figures *appear* to be these params for predator density
# Predator density (k_val) is doubled
preddens.df <- as.data.frame( mat.or.vec( length( speed.vec ), 6) )
colnames(preddens.df) <- c("trait", "speed", "growth_curve", "mort_curve", "fitness", "NCE_CE")
preddens.df$trait <- trait.vec
preddens.df$speed <- speed.vec
h_val <- 8
R_val <- 1
cost_constant_val <- 9
v_val <- 0.2 
k_val <- 0.9 
pred_constant_val <- 0.98058

preddens.df$growth_curve <- cost_constant_val * growth.curve( s = preddens.df$speed, R = R_val, h = h_val ) # Growth curve
preddens.df$mort_curve <- pred_constant_val * pred.curve( k = k_val, v = v_val, s = preddens.df$speed ) # Mortality curve
preddens.df$fitness <- preddens.df$growth_curve - preddens.df$mort_curve # fitness
preddens.df$NCE_CE <- (1 - preddens.df$growth_curve)/preddens.df$mort_curve # NCE/CE

preddens_optim <- preddens.df$trait[which(preddens.df$fitness == max(preddens.df$fitness))]
preddens_optim_fitness <- preddens.df$fitness[which(preddens.df$fitness == max(preddens.df$fitness))]
preddens_optim_CE <- preddens.df$mort_curve[which(preddens.df$fitness == max(preddens.df$fitness))]
preddens_optim_NCE <- 1 - preddens_optim_CE - preddens_optim_fitness
# Maximum mortality is at 0.8999994
```


```{r}
# These *appear* to be the params for foraging mode
# v is five times higher, but k is also (more than) doubled? why?
predmode.df <- as.data.frame( mat.or.vec( length( speed.vec ), 6) )
colnames(predmode.df) <- c("trait", "speed", "growth_curve", "mort_curve", "fitness", "NCE_CE")
predmode.df$trait <- trait.vec
predmode.df$speed <- speed.vec
h_val <- 8
R_val <- 1
cost_constant_val <- 9
v_val <- 1 # this value changes between the baseline and this one, not sure why
k_val <- 1 # This changes from 0.45, which I think is where the change in predator density is coming from
pred_constant_val <- 0.7071068 * 0.8999994 # this changes to anchor the mortality curve in the same spot


predmode.df$growth_curve <- cost_constant_val * growth.curve( s = predmode.df$speed, R = R_val, h = h_val ) # Growth curve
predmode.df$mort_curve <- pred_constant_val * pred.curve( k = k_val, v = v_val, s = predmode.df$speed ) # Mortality curve
predmode.df$fitness <- predmode.df$growth_curve - predmode.df$mort_curve # fitness
predmode.df$NCE_CE <- (1 - predmode.df$growth_curve)/predmode.df$mort_curve # NCE/CE

predmode_optim <- predmode.df$trait[which(predmode.df$fitness == max(predmode.df$fitness))]
predmode_optim_fitness <- predmode.df$fitness[which(predmode.df$fitness == max(predmode.df$fitness))]
predmode_optim_CE <- predmode.df$mort_curve[which(predmode.df$fitness == max(predmode.df$fitness))]
predmode_optim_NCE <- 1 - predmode_optim_CE - predmode_optim_fitness
```

```{r}
# Fig. 1
fig1_a.plot <- ggplot( data = baseline.df, aes( x = trait ) ) +
  geom_line( aes( y = growth_curve, color = "Growth" ), size = 1 ) +
  geom_line( aes( y = mort_curve, color = "Mortality" ), size = 1, linetype = "dashed" ) +
  theme_classic( base_size = 10, base_line_size = 1 ) + 
  ylab("Growth and Mortality") +
  xlab("Displacement Rate") +
  scale_x_continuous(expand = c(0, 0), breaks = c(0, 0.5, 1), labels = c(1000, 500, 1) ) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1), breaks = c(0, 0.5, 1) ) +
  annotate( "text", x = 0.25, y = 0.85, label = "Growth", size = 9/.pt  ) +
  annotate( "text", x = 0.5, y = 0.35, label = "Mortality", size = 9/.pt  ) +
  scale_color_manual(name = "Curve", values = c( "blue", "red" )) + 
  theme( panel.border = element_rect( color = "black", fill = NA, linewidth = 0.75 ), aspect.ratio = 1, legend.position = "bottom" ) + 
  theme(legend.key = element_rect(fill = "white"), legend.text = element_text(color = "white"), legend.title = element_text(color = "white")) +
  guides(color = guide_legend(override.aes = list(color = NA)))

fig1_b.plot <- ggplot( data = baseline.df, aes( x = trait ) ) +
  geom_line( aes( y = growth_curve, color = "Growth" ), size = 1 ) +
  geom_line( aes( y = mort_curve, color = "Mortality" ), size = 1, linetype = "dashed" ) +
  geom_ribbon( aes(xmin = 0.01, xmax = 0.99, ymin = growth_curve, ymax = 0.998), 
               color = NA, fill = "lightblue", alpha = 0.5) +
  geom_ribbon( aes(xmin = 0.01, xmax = 0.99, ymin = 0, ymax = mort_curve), 
               color = NA, fill = "red", alpha = 0.5) +
  geom_ribbon( aes(xmin = 0.01, xmax = 0.99, ymin = mort_curve, ymax = growth_curve), 
               color = NA, fill = "lightgreen", alpha = 0.5) +
  theme_classic( base_size = 10, base_line_size = 1 ) + 
  ylab("Growth and Mortality") +
  xlab("Displacement Rate") +
  annotate( "text", x = 0.5, y = 0.5, label = "Fitness = 1 - NCE - CE", size = 9/.pt  ) +
  annotate( "text", x = 0.5, y = 0.95, label = "NCE", size = 9/.pt  ) +
  annotate( "text", x = 0.5, y = 0.15, label = "CE", size = 9/.pt  ) +
  scale_x_continuous(expand = c(0, 0), breaks = c(0, 0.5, 1), labels = c(1000, 500, 1) ) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1), breaks = c(0, 0.5, 1) ) +
  scale_color_manual(name = "Curve", values = c( "blue", "red" )) + 
  theme( panel.border = element_rect( color = "black", fill = NA, linewidth = 0.75 ), aspect.ratio = 1,
         axis.title.y = element_blank(), legend.position = "bottom" )

fig1_c.plot <- ggplot( data = baseline.df, aes( x = trait ) ) +
  geom_line( aes( y = growth_curve, color = "Growth" ), size = 1 ) +
  geom_line( aes( y = mort_curve, color = "Mortality" ), size = 1, linetype = "dashed" ) +
  theme_classic( base_size = 10, base_line_size = 1 ) + 
  ylab("Growth and Mortality") +
  xlab("Displacement Rate") +
  annotate(geom = "rect", xmin = baseline_optim - 0.025, xmax = baseline_optim + 0.025, 
           ymin = 1 - baseline_optim_NCE, ymax = 0.998, 
           color = "gray2", fill = "lightblue", alpha = 0.5) +
  annotate( "text", x = baseline_optim + 0.05, y = 1 - baseline_optim_NCE/2, 
            label = "NCE = 0.12", hjust = "left", size = 9/.pt  ) +
  annotate(geom = "rect", xmin = baseline_optim - 0.025, xmax = baseline_optim + 0.025,
           ymin = baseline_optim_CE, ymax = baseline_optim_CE + baseline_optim_fitness, 
           color = "gray2", fill = "lightgreen", alpha = 0.5) +
  annotate( "text", x = baseline_optim + 0.1, y = baseline_optim_CE + baseline_optim_fitness/2, 
            label = "0.66", hjust = "left", size = 9/.pt  ) +
  annotate(geom = "rect", xmin = baseline_optim - 0.025, xmax = baseline_optim + 0.025, 
           ymin = 0.001, ymax = baseline_optim_CE, 
           color = "gray2", fill = "red", alpha = 0.5) +
  annotate( "text", x = baseline_optim + 0.05, y = baseline_optim_CE/2 - 0.05, 
            label = "CE = 0.22", hjust = "left", size = 9/.pt  ) +
  scale_x_continuous(expand = c(0, 0), breaks = c(0, 0.5, 1), labels = c(1000, 500, 1) ) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1), breaks = c(0, 0.5, 1) ) +
  scale_color_manual(name = "Curve", values = c( "blue", "red" )) + 
  theme( panel.border = element_rect( color = "black", fill = NA, linewidth = 0.75 ), aspect.ratio = 1,
         axis.title.y = element_blank(), legend.position = "bottom" ) + 
  theme(legend.key = element_rect(fill = "white"), legend.text = element_text(color = "white"), legend.title = element_text(color = "white")) +
  guides(color = guide_legend(override.aes = list(color = NA)))


fig1.plot <- plot_grid( fig1_a.plot, fig1_b.plot, fig1_c.plot, ncol = 3, labels = "auto", label_y = 0.75, align = c("hv") )

# ggsave("fig1.png", plot = fig1.plot, dpi = 300)

```

```{r}
# Fig. 2 - predator density
preddens.df <- cbind(preddens.df, baseline.df$mort_curve)
colnames(preddens.df) <- c("trait", "speed", "growth_curve", "mort_curve", "fitness", "NCE_CE", "mort_baseline")

preddens_1.plot <- ggplot( data = preddens.df, aes( x = trait ) ) +
  geom_line( aes( y = growth_curve, color = "Growth" ) ) +
  geom_line( aes( y = mort_baseline, color = "Mortality" ), linetype = "dashed" ) +
  annotate(geom = "rect", xmin = baseline_optim - 0.025, xmax = baseline_optim + 0.025, 
           ymin = 1 - baseline_optim_NCE, ymax = 0.998, 
           color = "gray2", fill = "lightblue", alpha = 0.5) +
  annotate( "text", x = baseline_optim + 0.05, y = 1 - baseline_optim_NCE/2, 
            label = "NCE = 0.12", hjust = "left", size = 8/.pt  ) +
  annotate(geom = "rect", xmin = baseline_optim - 0.025, xmax = baseline_optim + 0.025,
           ymin = baseline_optim_CE, ymax = baseline_optim_CE + baseline_optim_fitness, 
           color = "gray2", fill = "lightgreen", alpha = 0.5) +
  annotate( "text", x = baseline_optim + 0.1, y = baseline_optim_CE + baseline_optim_fitness/2, 
            label = "0.66", hjust = "left", size = 8/.pt  ) +
  annotate(geom = "rect", xmin = baseline_optim - 0.025, xmax = baseline_optim + 0.025, 
           ymin = 0.001, ymax = baseline_optim_CE, 
           color = "gray2", fill = "red", alpha = 0.5) +
  annotate( "text", x = baseline_optim + 0.05, y = baseline_optim_CE/2 - 0.05, 
            label = "CE = 0.22", hjust = "left", size = 8/.pt  ) +
  annotate("text", x = 0.125, y = 0.05, label = "P = 0.45", size = 9/.pt ) +
  theme_classic( base_size = 10, base_line_size = 1 ) + 
  ylab("Growth and Mortality") +
  xlab("Displacement Rate") +
  scale_x_continuous(expand = c(0, 0), breaks = c(0, 0.5, 1), labels = c(1000, 500, 1) ) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1), breaks = c(0, 0.5, 1) ) +
  scale_color_manual(name = "Curve", values = c( "blue", "red" )) + 
  theme( panel.border = element_rect( color = "black", fill = NA, linewidth = 0.75 ), aspect.ratio = 1,
         axis.title.x = element_blank(),
         axis.text.x = element_blank(), legend.position = "bottom" ) + 
  theme(legend.key = element_rect(fill = "white"), legend.text = element_text(color = "white"), legend.title = element_text(color = "white")) +
  guides(color = guide_legend(override.aes = list(color = NA)))

preddens_2.plot <- ggplot( data = preddens.df, aes( x = trait ) ) +
  geom_line( aes( y = growth_curve, colour = "Growth" ), ) +
  geom_line( aes( y = mort_curve, colour = "Mortality" ), linetype = "dashed" ) +
  theme_classic( base_size = 10, base_line_size = 1 ) + 
  ylab("Growth and Mortality") +
  xlab("Displacement Rate") +
  annotate(geom = "rect", xmin = preddens_optim - 0.025, xmax = preddens_optim + 0.025, 
           ymin = 1 - preddens_optim_NCE, ymax = 0.998, 
           color = "gray2", fill = "lightblue", alpha = 0.5) +
  annotate( "text", x = preddens_optim + 0.05, y = 1 - preddens_optim_NCE/2, 
            label = "NCE = 0.20", hjust = "left", size = 7/.pt  ) +
  annotate(geom = "rect", xmin = preddens_optim - 0.025, xmax = preddens_optim + 0.025,
           ymin = preddens_optim_CE, ymax = preddens_optim_CE + preddens_optim_fitness, 
           color = "gray2", fill = "lightgreen", alpha = 0.5) +
  annotate( "text", x = preddens_optim + 0.1, y = preddens_optim_CE + preddens_optim_fitness/2 - 0.1, 
            label = "0.48", hjust = "left", size = 7/.pt  ) +
  annotate(geom = "rect", xmin = preddens_optim - 0.025, xmax = preddens_optim + 0.025, 
           ymin = 0.001, ymax = preddens_optim_CE, 
           color = "gray2", fill = "red", alpha = 0.5) +
  annotate( "text", x = preddens_optim + 0.05, y = preddens_optim_CE/2 - 0.05, 
            label = "CE = 0.32", hjust = "left", size = 7/.pt  ) +
  annotate("text", x = 0.125, y = 0.05, label = "P = 0.90", size = 9/.pt ) +
  scale_x_continuous(expand = c(0, 0), breaks = c(0, 0.5, 1), labels = c(1000, 500, 1) ) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1), breaks = c(0, 0.5, 1) ) +
  scale_colour_manual(name = "Curve", values = c("blue", "red")) +
  theme( panel.border = element_rect( color = "black", fill = NA, linewidth = 0.75 ), aspect.ratio = 1 ) +
  theme( legend.position = "bottom" )


# Calculate change in NCE % of total predator effect, and optimal RITR, with decreasing h:
k_vec <- seq( 0, 2.5, by = 0.05 )
preddens_pctNCE.df <- as.data.frame(mat.or.vec( length(k_vec), 4 ) ) 
colnames(preddens_pctNCE.df) <- c("k", "NCE_TPE", "OptimalRITR", "Fitness")

for (i in 1:length( k_vec ) ){
preddens_pctNCE.df[i, 1] <- k_vec[i]

pd_growth_curve <- cost_constant_val * growth.curve( s = speed.vec, R = 1, h = 8 ) # Growth curve
pd_mort_curve <- pred_constant_val * pred.curve( k = k_vec[i], v = 0.2, s = speed.vec ) # Mortality curve
pd_fitness <- pd_growth_curve - pd_mort_curve # fitness
pd_NCE_TPE <- (1 - pd_growth_curve)/(pd_mort_curve + (1 - pd_growth_curve))# NCE/TPE

preddens_pctNCE.df[i, 2] <- pd_NCE_TPE[ which.max(pd_fitness) ]

# Some other important metrics
preddens_pctNCE.df[i, 3] <- speed.vec[ which.max(pd_fitness) ]
preddens_pctNCE.df[i, 4] <- max(pd_fitness)
}

# Drop NaN from first row
preddens_pctNCE.df[1, 2] <- 0

# Plot both:
preddens_3.plot <- ggplot( data = preddens_pctNCE.df, aes( x = k ) ) +
  geom_point( aes( y = NCE_TPE, fill = Fitness ), color = "black", shape = 21, size = 2.5 ) +
  theme_classic( base_size = 10, base_line_size = 1 ) + 
  ylab("% NCE Contribution to TPE") +
  xlab("Predator Density") +
  annotate('text', x = 0.45, y = 0.37, label = 'a') +
  annotate('text', x = 0.9, y = 0.42, label = 'b') +
  scale_x_continuous(expand = c(0, 0) ) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1), breaks = c(0, 0.25, 0.50, 0.75, 1) ) +
  scale_fill_continuous( type = "viridis" ) +
  theme( panel.border = element_rect( color = "black", fill = NA, linewidth = 0.75 ), aspect.ratio = 1, legend.position = "bottom")

preddens_4.plot <- ggplot( data = preddens_pctNCE.df, aes( x = k ) ) +
  geom_point( aes( y = OptimalRITR, fill = Fitness ), shape = 21, size = 2.5 ) +
  theme_classic( base_size = 10, base_line_size = 1 ) + 
  ylab("Optimal RITR") +
  xlab("Predator Density") +
  annotate('text', x = 0.45, y = 0.5, label = 'a') +
  annotate('text', x = 0.9, y = 0.35, label = 'b') +
  scale_x_continuous(expand = c(0, 0) ) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1), breaks = c(0, 0.50, 1), labels = c(0, 500, 1000) ) +
  scale_fill_continuous( type = "viridis" ) +
  theme( panel.border = element_rect( color = "black", fill = NA, linewidth = 0.75 ), aspect.ratio = 1,
         axis.title.x = element_blank(),
         axis.text.x = element_blank(), legend.position = "bottom" ) +
  guides(fill = "none")

preddens.fig <- plot_grid( preddens_1.plot, preddens_4.plot, preddens_2.plot, preddens_3.plot, ncol = 2, labels = c("a", "c", "b", "d"), hjust = -1.5, align = c("hv") )

 ggsave("fig2.png", plot = preddens.fig, dpi = 300)
```

```{r}
# Fig. 3 - predator hunting mode
predmode.df <- cbind(predmode.df, preddens.df$mort_curve)
colnames(predmode.df) <- c("trait", "speed", "growth_curve", "mort_curve", "fitness", "NCE_CE", "mort_baseline")

predmode_1.plot <- ggplot( data = predmode.df, aes( x = trait ) ) +
  geom_line( aes( y = growth_curve, color = "Growth" ) ) +
  geom_line( aes( y = mort_baseline, color = "Mortality" ), linetype = "dashed" ) +
  annotate(geom = "rect", xmin = preddens_optim - 0.025, xmax = preddens_optim + 0.025, 
           ymin = 1 - preddens_optim_NCE, ymax = 0.998, 
           color = "gray2", fill = "lightblue", alpha = 0.5) +
  annotate( "text", x = preddens_optim + 0.05, y = 1 - preddens_optim_NCE/2, 
            label = "NCE = 0.20", hjust = "left", size = 7/.pt  ) +
  annotate(geom = "rect", xmin = preddens_optim - 0.025, xmax = preddens_optim + 0.025,
           ymin = preddens_optim_CE, ymax = preddens_optim_CE + preddens_optim_fitness, 
           color = "gray2", fill = "lightgreen", alpha = 0.5) +
  annotate( "text", x = preddens_optim + 0.1, y = preddens_optim_CE + preddens_optim_fitness/2 - 0.05, 
            label = "0.48", hjust = "left", size = 7/.pt  ) +
  annotate(geom = "rect", xmin = preddens_optim - 0.025, xmax = preddens_optim + 0.025, 
           ymin = 0.001, ymax = preddens_optim_CE, 
           color = "gray2", fill = "red", alpha = 0.5) +
  annotate( "text", x = preddens_optim + 0.05, y = preddens_optim_CE/2, 
            label = "CE = 0.32", hjust = "left", size = 7/.pt  ) +
  annotate("text", x = 0.1, y = 0.05, label = "v = 0.2", size = 9/.pt ) +
  theme_classic( base_size = 10, base_line_size = 1 ) + 
  ylab("Growth and Mortality") +
  xlab("Displacement Rate") +
  scale_x_continuous(expand = c(0, 0), breaks = c(0, 0.5, 1), labels = c(1000, 500, 1) ) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1), breaks = c(0, 0.5, 1) ) +
  scale_color_manual(name = "Curve", values = c( "blue", "red" )) + 
  theme( panel.border = element_rect( color = "black", fill = NA, linewidth = 0.75 ), aspect.ratio = 1,
         axis.title.x = element_blank(),
         axis.text.x = element_blank(), legend.position = "bottom" ) + 
  guides(color = "none")


predmode_2.plot <- ggplot( data = predmode.df, aes( x = trait ) ) +
  geom_line( aes( y = growth_curve, colour = "Growth" ), ) +
  geom_line( aes( y = mort_curve, colour = "Mortality" ), linetype = "dashed" ) +
  theme_classic( base_size = 10, base_line_size = 1 ) + 
  ylab("Growth and Mortality") +
  xlab("Displacement Rate") +
  annotate(geom = "rect", xmin = predmode_optim - 0.025, xmax = predmode_optim + 0.025, 
           ymin = 1 - predmode_optim_NCE, ymax = 0.998, 
           color = "gray2", fill = "lightblue", alpha = 0.5) +
  annotate( "text", x = predmode_optim + 0.05, y = 1 - predmode_optim_NCE/2, 
            label = "NCE = 0.08", hjust = "left", size = 7/.pt  ) +
  annotate(geom = "rect", xmin = predmode_optim - 0.025, xmax = predmode_optim + 0.025,
           ymin = predmode_optim_CE, ymax = predmode_optim_CE + predmode_optim_fitness, 
           color = "gray2", fill = "lightgreen", alpha = 0.5) +
  annotate( "text", x = predmode_optim + 0.1, y = predmode_optim_CE + predmode_optim_fitness/2 - 0.05, 
            label = "0.19", hjust = "left", size = 7/.pt  ) +
  annotate(geom = "rect", xmin = predmode_optim - 0.025, xmax = predmode_optim + 0.025, 
           ymin = 0.001, ymax = predmode_optim_CE, 
           color = "gray2", fill = "red", alpha = 0.5) +
  annotate( "text", x = predmode_optim + 0.05, y = predmode_optim_CE/2, 
            label = "CE = 0.73", hjust = "left", size = 7/.pt  ) +
  annotate("text", x = 0.1, y = 0.05, label = "v = 1.0", size = 9/.pt ) +
  scale_x_continuous(expand = c(0, 0), breaks = c(0, 0.5, 1), labels = c(1000, 500, 1) ) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1), breaks = c(0, 0.5, 1) ) +
  scale_colour_manual(name = "Curve", values = c("blue", "red")) +
  theme( panel.border = element_rect( color = "black", fill = NA, linewidth = 0.75 ), aspect.ratio = 1 ) +
  theme( legend.position = "bottom" )


# Calculate change in NCE % of total predator effect, and optimal RITR, with decreasing h:
v_vec <- seq( 0, 2.5, by = 0.05 )
predmode_pctNCE.df <- as.data.frame(mat.or.vec( length(v_vec), 4 ) )
pred_constant_val <- 0.7071068 * 0.8999994 # this changes to anchor the mortality curve in the same spot
colnames(predmode_pctNCE.df) <- c("v", "NCE_TPE", "OptimalRITR", "Fitness")

for (i in 1:length( v_vec ) ){
predmode_pctNCE.df[i, 1] <- v_vec[i]

pm_growth_curve <- cost_constant_val * growth.curve( s = speed.vec, R = 1, h = 8 ) # Growth curve
pm_mort_curve <- pred_constant_val * pred.curve( k = 1, v = v_vec[i], s = speed.vec ) # Mortality curve
pm_fitness <- pm_growth_curve - pm_mort_curve # fitness
pm_NCE_TPE <- (1 - pm_growth_curve)/(pm_mort_curve + (1 - pm_growth_curve))# NCE/TPE

predmode_pctNCE.df[i, 2] <- pm_NCE_TPE[ which.max(pm_fitness) ]

# Some other important metrics
predmode_pctNCE.df[i, 3] <- speed.vec[ which.max(pm_fitness) ]
predmode_pctNCE.df[i, 4] <- max(pm_fitness)
}

# Plot both:
predmode_3.plot <- ggplot( data = predmode_pctNCE.df, aes( x = v ) ) +
  geom_point( aes( y = NCE_TPE, fill = Fitness ), shape = 21, color = "black", size = 2.5) +
  theme_classic( base_size = 10, base_line_size = 1 ) + 
  ylab("% NCE Contribution to TPE") +
  xlab("Predator Velocity") +
  annotate('text', x = 0.2, y = 0.5, label = 'a') +
  annotate('text', x = 1, y = 0.17, label = 'b') +
  scale_x_continuous(expand = c(0, 0) ) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1), breaks = c(0, 0.25, 0.50, 0.75, 1) ) +
  scale_fill_continuous( type = "viridis" ) +
  theme( panel.border = element_rect( color = "black", fill = NA, linewidth = 0.75 ), aspect.ratio = 1, legend.position = "bottom")

predmode_4.plot <- ggplot( data = predmode_pctNCE.df, aes( x = v ) ) +
  geom_point( aes( y = OptimalRITR, fill = Fitness ), shape = 21, color = "black", size = 2.5 ) +
  theme_classic( base_size = 10, base_line_size = 1 ) + 
  ylab("Optimal RITR") +
  xlab("Predator Velocity") +
  annotate('text', x = 0.2, y = 0.45, label = 'a') +
  annotate('text', x = 1, y = 0.62, label = 'b') +
  scale_x_continuous(expand = c(0, 0) ) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1), breaks = c(0, 0.50, 1), labels = c(0, 500, 1000) ) +
  scale_fill_continuous( type = "viridis" ) +
  theme( panel.border = element_rect( color = "black", fill = NA, linewidth = 0.75 ), 
         axis.title.x = element_blank(),
         axis.text.x = element_blank(),
         aspect.ratio = 1, 
         legend.position = "bottom" ) +
  guides( fill = "none" )

predmode.fig <- plot_grid( predmode_1.plot, predmode_4.plot, predmode_2.plot, predmode_3.plot, ncol = 2, labels = c("a", "c", "b", "d"), hjust = -1.5, align = c("hv") )

ggsave("fig3.png", plot = predmode.fig, dpi = 300)
```


# A different trait response for resource intake rate:

```{r}
# Fig. 4

## Derive the trait curves
proptime_vec <- seq(1, 0, by = -0.01)

refuge.df <- as.data.frame( mat.or.vec( length(proptime_vec), 5 ) ) 
colnames(refuge.df) <- c("proptime", "growth_curve", "mort_curve", "fitness", "NCE_CE")
refuge.df$proptime <- proptime_vec 

# Growth follows a Type II FR, where instead of speed we're manipulating resource density
resource_vec <- seq(0, 1, by = 0.01) # This needs to be the same length as proptime_vec, here the refuge has ZERO resources
refuge.df$growth_curve <- growth.curve(s = 1, R = resource_vec, h = 10)
refuge.df$growth_curve <- refuge.df$growth_curve * (1/max(refuge.df$growth_curve)) # Anchor at 1

# Mortality curve, which linearly declines from val_A to val_B
mort_val_A <- 0.4 # Mortality outside the refuge during an hour
mort_val_B <- 0 # Mortality inside the refuge during an hour
refuge.df$mort_curve <- ( 1 - refuge.df$proptime ) * mort_val_A + ( refuge.df$proptime) * mort_val_B

# Calculate fitness and relative strength of NCEs and CEs
refuge.df$fitness <- 1 - ( 1 - refuge.df$growth_curve ) - refuge.df$mort_curve
refuge.df$NCE_CE <- refuge.df$growth_curve / refuge.df$mort_curve

# Some other important metrics
refuge_optim <- refuge.df$proptime[which(refuge.df$fitness == max(refuge.df$fitness))] # The level of refuge use (along x-axis) that fitness is optimized at
refuge_optim_fitness <- refuge.df$fitness[which(refuge.df$fitness == max(refuge.df$fitness))] # The fitness value at the optimum
refuge_optim_CE <- refuge.df$mort_curve[which(refuge.df$fitness == max(refuge.df$fitness))] # The strength of the CE at the optimum
refuge_optim_NCE <- 1 - refuge_optim_CE - refuge_optim_fitness # The strength of the NCE at the optimum





### Second round, lower h to 1
## Derive the trait curves
refuge2.df <- as.data.frame( mat.or.vec( length(proptime_vec), 5 ) ) 
colnames(refuge2.df) <- c("proptime", "growth_curve", "mort_curve", "fitness", "NCE_CE")
refuge2.df$proptime <- proptime_vec 

# Growth follows a Type II FR, where instead of speed we're manipulating resource density
refuge2.df$growth_curve <- growth.curve(s = 1, R = resource_vec, h = 2)
refuge2.df$growth_curve <- refuge2.df$growth_curve * (1/max(refuge2.df$growth_curve)) # Anchor at 1


# Mortality curve, which linearly declines from val_A to val_B
refuge2.df$mort_curve <- ( 1 - refuge2.df$proptime ) * mort_val_A + ( refuge2.df$proptime) * mort_val_B

# Calculate fitness and relative strength of NCEs and CEs
refuge2.df$fitness <- 1 - ( 1 - refuge2.df$growth_curve ) - refuge2.df$mort_curve
refuge2.df$NCE_CE <- refuge2.df$growth_curve / refuge2.df$mort_curve

# Some other important metrics
refuge_optim2 <- refuge2.df$proptime[which(refuge2.df$fitness == max(refuge2.df$fitness))] # The level of refuge use (along x-axis) that fitness is optimized at
refuge_optim_fitness2 <- refuge2.df$fitness[which(refuge2.df$fitness == max(refuge2.df$fitness))] # The fitness value at the optimum
refuge_optim_CE2 <- refuge2.df$mort_curve[which(refuge2.df$fitness == max(refuge2.df$fitness))] # The strength of the CE at the optimum
refuge_optim_NCE2 <- 1 - refuge_optim_CE2 - refuge_optim_fitness2 # The strength of the NCE at the optimum



# Plot them all
refuge_1.plot <- ggplot( data = refuge.df, aes( x = proptime ) ) +
  geom_line( aes( y = mort_curve, color = "Mortality" ), linetype = "dashed" ) +
  geom_line( aes( y = growth_curve, color = "Growth" ) ) +
  theme_classic( base_size = 10, base_line_size = 1 ) + 
  annotate(geom = "rect", xmin = refuge_optim - 0.025, xmax = refuge_optim + 0.025, 
           ymin = 1 - refuge_optim_NCE, ymax = 0.998, 
           color = "gray2", fill = "lightblue", alpha = 0.5) +
  annotate( "text", x = refuge_optim - 0.25, y = 1 - refuge_optim_NCE/2 + 0.035, 
            label = "NCE = 0.11", hjust = "left", size = 7/.pt  ) +
  annotate(geom = "rect", xmin = refuge_optim - 0.025, xmax = refuge_optim + 0.025,
           ymin = refuge_optim_CE, ymax = refuge_optim_CE + refuge_optim_fitness, 
           color = "gray2", fill = "lightgreen", alpha = 0.5) +
  annotate( "text", x = refuge_optim - 0.2, y = refuge_optim_CE + refuge_optim_fitness/2, 
            label = "0.72", hjust = "left", size = 7/.pt  ) +
  annotate(geom = "rect", xmin = refuge_optim - 0.025, xmax = refuge_optim + 0.025, 
           ymin = 0.001, ymax = refuge_optim_CE, 
           color = "gray2", fill = "red", alpha = 0.5) +
  annotate( "text", x = refuge_optim - 0.25, y = refuge_optim_CE/2, 
            label = "CE = 0.17", hjust = "left", size = 7/.pt  ) +
  annotate("text", x = 0.85, y = 0.95, label = "1/h = 0.1", size = 9/.pt ) +
  ylab("Growth and Mortality") +
  xlab("Proportion of Time Spent in Refuge") +
  scale_x_continuous(expand = c(0, 0), breaks = c(0, 0.5, 1), labels = c(0, 0.5, 1) ) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1), breaks = c(0, 0.5, 1) ) +
  scale_color_manual(name = "Curve", values = c( "blue", "red" )) + 
  theme( panel.border = element_rect( color = "black", fill = NA, linewidth = 0.75 ), aspect.ratio = 1,
         axis.title.x = element_blank(),
         axis.text.x = element_blank(), legend.position = "bottom" ) + 
  guides(color = "none")


refuge_2.plot <- ggplot( data = refuge2.df, aes( x = proptime ) ) +
  geom_line( aes( y = growth_curve, color = "Growth" ) ) +
  geom_line( aes( y = mort_curve, color = "Mortality" ), linetype = "dashed" ) +
  annotate(geom = "rect", xmin = refuge_optim2 - 0.025, xmax = refuge_optim2 + 0.025, 
           ymin = 1 - refuge_optim_NCE2, ymax = 0.998, 
           color = "gray2", fill = "lightblue", alpha = 0.5) +
  annotate( "text", x = refuge_optim2 + 0.05, y = 1 - refuge_optim_NCE2/2, 
            label = "NCE = 0.05", hjust = "left", size = 7/.pt  ) +
  annotate(geom = "rect", xmin = refuge_optim2 - 0.025, xmax = refuge_optim2 + 0.025,
           ymin = refuge_optim_CE2, ymax = refuge_optim_CE2 + refuge_optim_fitness2, 
           color = "gray2", fill = "lightgreen", alpha = 0.5) +
  annotate( "text", x = refuge_optim2 + 0.1, y = refuge_optim_CE2 + refuge_optim_fitness2/2, 
            label = "0.60", hjust = "left", size = 7/.pt  ) +
  annotate(geom = "rect", xmin = refuge_optim2 - 0.025, xmax = refuge_optim2 + 0.025, 
           ymin = 0.001, ymax = refuge_optim_CE2, 
           color = "gray2", fill = "red", alpha = 0.5) +
  annotate( "text", x = refuge_optim2 + 0.05, y = refuge_optim_CE2/2, 
            label = "CE = 0.35", hjust = "left", size = 7/.pt  ) +
  annotate("text", x = 0.85, y = 0.95, label = "1/h = 0.5", size = 9/.pt ) +
  theme_classic( base_size = 10, base_line_size = 1 ) + 
  ylab("Growth and Mortality") +
  xlab("Proportion of Time Spent in Refuge") +
  scale_x_continuous(expand = c(0, 0), breaks = c(0, 0.5, 1), labels = c(0, 0.5, 1) ) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1), breaks = c(0, 0.5, 1) ) +
  scale_color_manual(name = "Curve", values = c( "blue", "red" )) + 
  theme( panel.border = element_rect( color = "black", fill = NA, linewidth = 0.75 ), aspect.ratio = 1 ) +
  theme( legend.position = "bottom" )


# Plot % NPE of TPE and optimal RITR
# Calculate change in NCE % of total predator effect, and optimal RITR, with decreasing h:
h_vec <- seq( 15, 1.5, by = -0.1 )
prey_h_pctNCE.df <- as.data.frame(mat.or.vec( length(h_vec), 4 ) )
colnames(prey_h_pctNCE.df) <- c("h", "NCE_TPE", "OptimalRITR", "Fitness")

for (i in 1:length( h_vec ) ){
prey_h_pctNCE.df[i, 1] <- h_vec[i]

# Growth follows a Type II FR, where instead of speed we're manipulating resource density
resource_vec <- seq(0, 1, by = 0.01) # This needs to be the same length as proptime_vec, here the refuge has ZERO resources
ph_growth_curve <- growth.curve(s = 1, R = resource_vec, h = h_vec[i])
ph_growth_curve <- ph_growth_curve * (1/max(ph_growth_curve)) # Anchor at 1

# Mortality curve, which linearly declines from val_A to val_B
mort_val_A <- 0.4 # Mortality outside the refuge during an hour
mort_val_B <- 0 # Mortality inside the refuge during an hour
ph_mort_curve <- ( 1 - refuge.df$proptime ) * mort_val_A + ( refuge.df$proptime) * mort_val_B


ph_fitness <- ph_growth_curve - ph_mort_curve # fitness
ph_NCE_TPE <- (1 - ph_growth_curve)/(ph_mort_curve + (1 - ph_growth_curve))# NCE/TPE

prey_h_pctNCE.df[i, 2] <- ph_NCE_TPE[ which.max(ph_fitness) ]

# Some other important metrics
prey_h_pctNCE.df[i, 3] <- proptime_vec[ which.max(ph_fitness) ]
prey_h_pctNCE.df[i, 4] <- max(ph_fitness)
}

prey_h_pctNCE.df$intake <- 1/prey_h_pctNCE.df$h




# Plot both:
refuge_3.plot <- ggplot( data = prey_h_pctNCE.df, aes( x = intake ) ) +
  geom_point( aes( y = NCE_TPE, fill = Fitness ), shape = 21, size = 2.5, color = "black" ) +
  theme_classic( base_size = 10, base_line_size = 1 ) + 
  ylab("% NCE Contribution to TPE") +
  xlab("Prey Resource Processing Rate (1/h)") +
  annotate('text', x = 0.1, y = 0.475, label = 'a') +
  annotate('text', x = 0.5, y = 0.2, label = 'b') +
  scale_x_continuous(expand = c(0, 0) ) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1), breaks = c(0, 0.25, 0.50, 0.75, 1) ) +
  scale_fill_continuous(type = "viridis" ) +
  theme( panel.border = element_rect( color = "black", fill = NA, linewidth = 0.75 ), aspect.ratio = 1, 
         legend.position = "bottom" )


refuge_4.plot <- ggplot( data = prey_h_pctNCE.df, aes( x = intake ) ) +
  geom_point( aes( y = OptimalRITR, fill = Fitness ), shape = 21, size = 2.5, color = "black" ) +
  theme_classic( base_size = 10, base_line_size = 1 ) + 
  ylab("Optimal RITR") +
  xlab("Prey Resource Processing Rate (1/h)") +
  annotate('text', x = 0.1, y = 0.675, label = 'a') +
  annotate('text', x = 0.5, y = 0.19, label = 'b') +
  scale_x_continuous(expand = c(0, 0) ) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1), breaks = c(0, 0.25, 0.5, 0.75, 1) ) +
  scale_fill_continuous(type = "viridis") +
  theme( panel.border = element_rect( color = "black", fill = NA, linewidth = 0.75 ), axis.title.x = element_blank(),
         axis.text.x = element_blank(), aspect.ratio = 1, legend.position = "bottom" ) +
  guides(fill = "none" )


predmode.fig <- plot_grid( refuge_1.plot, refuge_4.plot, refuge_2.plot, refuge_3.plot, ncol = 2, labels = c("a", "c", "b", "d"), hjust = -1.5, align = c("hv") )

ggsave("fig4.png", plot = predmode.fig, dpi = 300)

```

# Warming vs. prop time

```{r}
## Derive the trait curves
therm_refuge.df <- as.data.frame( mat.or.vec( length(proptime_vec), 5 ) ) 
colnames(therm_refuge.df) <- c("proptime", "growth_curve", "mort_curve", "fitness", "NCE_CE")
therm_refuge.df$proptime <- rev(proptime_vec)


## The growth-trait curve
# a.growth determines the width of the performance curve and should be negative, so that the curve is concave down
# h.growth determines the temperature at which prey growth peaks
# c.growth is a scaling constant
# temp_C is temperature in degrees Celsius
growth.curve_warming <- function( c_g, a_g, h_g, temp_C ){
  c_g * exp( a_g * ( ( temp_C - h_g )^2 ) + 1)
}

# Set a.growth to an arbitrary value that makes the curve broader than the predator's distribution; smaller values yield broader curves
a_g.val <- -0.05

# Set h.growth to the same point that the predator's density peaks
h_g.val <- 20

# c.growth needs to be 1/e for this parameterization to peak at 1
c_g.val <- 1/exp(1)

# The growth curve will follow a temperature gradient along the prey's TPC
temp_C_vec <- seq(20, 23, by = 0.03) # this has to be the same length as proptime
therm_refuge.df$growth_curve <- growth.curve_warming( c_g = c_g.val, a_g = a_g.val, h_g = h_g.val, temp_C = temp_C_vec )

# Mortality curve
mort_val_A <- 0.45 # Mortality outside the refuge during an hour
mort_val_B <- 0.2 # Mortality inside the refuge during an hour
therm_refuge.df$mort_curve <- ( 1 - therm_refuge.df$proptime ) * mort_val_A + ( therm_refuge.df$proptime) * mort_val_B


# Calculate fitness and relative strength of NCEs and CEs
therm_refuge.df$fitness <- 1 - ( 1 - therm_refuge.df$growth_curve ) - therm_refuge.df$mort_curve
therm_refuge.df$NCE_CE <- therm_refuge.df$growth_curve / therm_refuge.df$mort_curve

# Some other important metrics
therm_refuge_optim <- therm_refuge.df$proptime[which(therm_refuge.df$fitness == max(therm_refuge.df$fitness))] # The level of refuge use (along x-axis) that fitness is optimized at
therm_refuge_optim_fitness <- therm_refuge.df$fitness[which(therm_refuge.df$fitness == max(therm_refuge.df$fitness))] # The fitness value at the optimum
therm_refuge_optim_CE <- therm_refuge.df$mort_curve[which(therm_refuge.df$fitness == max(therm_refuge.df$fitness))] # The strength of the CE at the optimum
therm_refuge_optim_NCE <- 1 - therm_refuge_optim_CE - therm_refuge_optim_fitness # The strength of the NCE at the optimum

#Make a plot of the initial TPC
temp_vec <- seq( 10, 26, by = 0.1 )
tpc.df <- as.data.frame( mat.or.vec (length(temp_vec), 2) )
colnames(tpc.df) <- c("temperature", "performance")
tpc.df$temperature <- temp_vec
tpc.df$performance <- growth.curve_warming( c_g = c_g.val, a_g = a_g.val, h_g = h_g.val, temp_C = temp_vec)

tpc_labels.df <- as.data.frame( mat.or.vec ( 2, 2 ) )
colnames(tpc_labels.df) <- c("temperature", "performance")
tpc_labels.df$temperature <- c(20, 23)
tpc_labels.df$performance <- c( 1, 0.637628152)

tpc_1.plot <- ggplot(data = tpc.df, aes( x = temperature, y = performance ) ) +
  geom_line() + 
  geom_point( data = tpc_labels.df ) +
  geom_text( data = tpc_labels.df, label = c("A", "B"), nudge_x = 1 ) + 
  theme_classic( base_size = 10, base_line_size = 1 ) + 
  ylab("Prey Growth") +
  xlab("Temperature (°C)") +
  scale_x_continuous(expand = c(0, 0), breaks = c(10, 17.5, 25), labels = c("Cool", "Warm", "Hot") ) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1.1), breaks = c(0, 0.5, 1) ) +
  theme( panel.border = element_rect( color = "black", fill = NA, linewidth = 0.75 ), aspect.ratio = 1,
         axis.title.x = element_blank(),
         axis.text.x = element_blank() ) 




# Plot of Unwarmed Scenario
therm_refuge_1.plot <- ggplot( data = therm_refuge.df, aes( x = proptime ) ) +
  geom_line( aes( y = growth_curve, color = "Growth" ) ) +
  geom_line( aes( y = mort_curve, color = "red" ), linetype = "dashed" ) +
  annotate(geom = "rect", xmin = therm_refuge_optim - 0.025, xmax = therm_refuge_optim + 0.025, 
           ymin = 1 - therm_refuge_optim_NCE, ymax = 0.998, 
           color = "gray2", fill = "lightblue", alpha = 0.5) +
  annotate( "text", x = therm_refuge_optim + 0.05, y = 0.975, 
            label = "NCE = 0.04", hjust = "left", size = 8/.pt  ) +
  annotate(geom = "rect", xmin = therm_refuge_optim - 0.025, xmax = therm_refuge_optim + 0.025,
           ymin = therm_refuge_optim_CE, ymax = therm_refuge_optim_CE + therm_refuge_optim_fitness, 
           color = "gray2", fill = "lightgreen", alpha = 0.5) +
  annotate( "text", x = therm_refuge_optim + 0.1, y = therm_refuge_optim_CE + therm_refuge_optim_fitness/2, 
            label = "0.59", hjust = "left", size = 8/.pt  ) +
  annotate(geom = "rect", xmin = therm_refuge_optim - 0.025, xmax = therm_refuge_optim + 0.025, 
           ymin = 0.001, ymax = therm_refuge_optim_CE, 
           color = "gray2", fill = "red", alpha = 0.5) +
  annotate( "text", x = therm_refuge_optim + 0.05, y = 0.05, 
            label = "CE = 0.37", hjust = "left", size = 8/.pt  ) +
  theme_classic( base_size = 10, base_line_size = 1 ) + 
  ylab("Growth and Mortality") +
  xlab("Proportion of Time Spent in Refuge") +
  scale_x_continuous(expand = c(0, 0), breaks = c(0, 0.5, 1), labels = c(1000, 500, 1) ) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1), breaks = c(0, 0.5, 1) ) +
  scale_color_manual(name = "Curve", values = c( "blue", "red" )) + 
  theme( panel.border = element_rect( color = "black", fill = NA, linewidth = 0.75 ), aspect.ratio = 1,
         axis.title.x = element_blank(),
         axis.text.x = element_blank(), legend.position = "bottom" ) + 
  theme(legend.key = element_rect(fill = "white"), legend.text = element_text(color = "white"), legend.title = element_text(color = "white")) +
  guides(color = guide_legend(override.aes = list(color = NA)))


# Vs. warmed scenario

## Derive the trait curves
therm_refuge2.df <- as.data.frame( mat.or.vec( length(proptime_vec), 5 ) ) 
colnames(therm_refuge2.df) <- c("proptime", "growth_curve", "mort_curve", "fitness", "NCE_CE")
therm_refuge2.df$proptime <- rev( proptime_vec )

# The growth curve will follow a temperature gradient along the prey's TPC
temp_C_vec <- seq(22.5, 25.5, by = 0.03) # this has to be the same length as proptime
therm_refuge2.df$growth_curve <- growth.curve_warming( c_g = c_g.val, a_g = a_g.val, h_g = h_g.val, temp_C = temp_C_vec )
therm_refuge2.df$growth_curve <- therm_refuge2.df$growth_curve * (1/ max(therm_refuge2.df$growth_curve) )


# Mortality curve
mort_val_A <- 0.45 # Mortality outside the refuge during an hour
mort_val_B <- 0.2 # Mortality inside the refuge during an hour

# Weighted mean of risk during time spent outside refuge + risk during time spent inside refuge
therm_refuge2.df$mort_curve <- ( 1 - therm_refuge2.df$proptime ) * mort_val_A + ( therm_refuge2.df$proptime ) * mort_val_B 

# Calculate fitness and relative strength of NCEs and CEs
therm_refuge2.df$fitness <- 1 - ( 1 - therm_refuge2.df$growth_curve ) - therm_refuge2.df$mort_curve
therm_refuge2.df$NCE_CE <- therm_refuge2.df$growth_curve / therm_refuge2.df$mort_curve

# Some other important metrics
therm_refuge_optim2 <- therm_refuge2.df$proptime[which(therm_refuge2.df$fitness == max(therm_refuge2.df$fitness))] # The level of refuge use (along x-axis) that fitness is optimized at
therm_refuge_optim_fitness2 <- therm_refuge2.df$fitness[which(therm_refuge2.df$fitness == max(therm_refuge2.df$fitness))] # The fitness value at the optimum
therm_refuge_optim_CE2 <- therm_refuge2.df$mort_curve[which(therm_refuge2.df$fitness == max(therm_refuge2.df$fitness))] # The strength of the CE at the optimum
therm_refuge_optim_NCE2 <- 1 - therm_refuge_optim_CE2 - therm_refuge_optim_fitness2 # The strength of the NCE at the optimum


# Repeat TPC plot with new temperatures
tpc_labels2.df <- as.data.frame( mat.or.vec ( 2, 2 ) )
colnames(tpc_labels2.df) <- c("temperature", "performance")
tpc_labels2.df$temperature <- c(22.5, 25.5)
tpc_labels2.df$performance <- c( 0.731615629, 0.220358393)

tpc_2.plot <- ggplot(data = tpc.df, aes( x = temperature, y = performance ) ) +
  geom_line() + 
  geom_point( data = tpc_labels2.df ) +
  geom_text( data = tpc_labels2.df, label = c("A", "B"), nudge_x = -1 ) + 
  annotate("segment", x = 17.5, y = 0.25, xend = 23, yend = 0.25,
    arrow = arrow(),
    color = "red",
    size = 1) +
  annotate( "text", x = 20, y = 0.4, color = "red", label = "Warming") +
  theme_classic( base_size = 10, base_line_size = 1 ) + 
  ylab("Prey Growth") +
  xlab("Temperature (°C)") +
  scale_x_continuous(expand = c(0, 0), breaks = c(10, 17.5, 25), labels = c("Cool", "Warm", "Hot") ) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1.1), breaks = c(0, 0.5, 1) ) +
  theme( panel.border = element_rect( color = "black", fill = NA, linewidth = 0.75 ), aspect.ratio = 1 )


therm_refuge_2.plot <- ggplot( data = therm_refuge2.df, aes( x = proptime ) ) +
  geom_line( aes( y = growth_curve, color = "Growth" ) ) +
  geom_line( aes( y = mort_curve, color = "Mortality" ), linetype = "dashed" ) +
  annotate(geom = "rect", xmin = therm_refuge_optim2, xmax = therm_refuge_optim2 + 0.05, 
           ymin = 1 - therm_refuge_optim_NCE2, ymax = 0.998, 
           color = "gray2", fill = "lightblue", alpha = 0.5) +
  annotate( "text", x = therm_refuge_optim2 + 0.1, y = 0.975, 
            label = "NCE = 0", hjust = "left", size = 8/.pt  ) +
  annotate(geom = "rect", xmin = therm_refuge_optim2, xmax = therm_refuge_optim2 + 0.05,
           ymin = therm_refuge_optim_CE2, ymax = therm_refuge_optim_CE2 + therm_refuge_optim_fitness2, 
           color = "gray2", fill = "lightgreen", alpha = 0.5) +
  annotate( "text", x = therm_refuge_optim2 + 0.15, y = therm_refuge_optim_CE2 + therm_refuge_optim_fitness2/2, 
            label = "0.55", hjust = "left", size = 8/.pt  ) +
  annotate(geom = "rect", xmin = therm_refuge_optim2, xmax = therm_refuge_optim2 + 0.05, 
           ymin = 0.001, ymax = therm_refuge_optim_CE2, 
           color = "gray2", fill = "red", alpha = 0.5) +
  annotate( "text", x = therm_refuge_optim2 + 0.1, y = therm_refuge_optim_CE2/2, 
            label = "CE = 0.45", hjust = "left", size = 8/.pt  ) +
  theme_classic( base_size = 10, base_line_size = 1 ) + 
  ylab("Growth and Mortality") +
  xlab("Proportion of Time Spent in Refuge") +
  scale_x_continuous(expand = c(0, 0), breaks = c(0, 0.5, 1), labels = c(0, 0.5, 1) ) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1), breaks = c(0, 0.5, 1) ) +
  scale_color_manual(name = "Curve", values = c( "blue", "red" )) + 
  theme( panel.border = element_rect( color = "black", fill = NA, linewidth = 0.75 ), aspect.ratio = 1) +
  theme( legend.position = "bottom" )

warming.fig <- plot_grid( tpc_1.plot, therm_refuge_1.plot, tpc_2.plot, therm_refuge_2.plot, ncol = 2, labels = c("a", "c", "b", "d"), hjust = -1.5, align = c("hv") )


ggsave("fig5.png", plot = warming.fig, dpi = 300)
```