---
title: 'Trait: Displacement Rate'
author: "Andy Davidson"
date: "2025-12-08"
output: pdf_document
---

```{r Setup Config}
library(ggbrace)
library(knitr)
knitr::opts_chunk$set(tidy.opts = list(width.cutoff = 60), tidy = TRUE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

# Trait: Prey Displacement Rate

We describe a prey species that modifies its displacement rate (as a proxy for foraging rate) in response to predation risk. We explore, using IREF, how manipulating different parameters describing the predator and prey's biology will influence the optimal RITR and strength of NCEs and CEs.

```{r Displacement Trait Curve Functions}
## Establishing trait ranges and formulae for generating the curves
trait.vec <- seq(0, 1, by = 0.01) # Provide a gradient of prey trait expression
speed.vec <- 1 - trait.vec # Prey speed, s, is 1 - that vector
```

## 1. Predation Curve

We assume that the predator exhibits a Type I functional response. Thus, the encounter rate equation is:

$$ Z = 2rN*sqrt(v^2 + s^2)$$

where $s$ is prey speed (i.e., displacement rate), $v$ is predator speed, $N$ is predator density, and $r$ is the radius within which the predator can perceive prey.

Scott simulates this equation for varying levels of $s$, and four constant levels of $v$ (0, 0.5, 1, and 2). Predator density and detection radius are kept constant such that $2rN = 1$.

$v = 0$ would simply mean a sit-and-wait predator.

```{r Predation Curve}
## The Predation curve
# r is always assumed to be 1, while P, s, and v are manipulated values
pred.curve <- function(P, v, s){
  P*sqrt(v^2 + s^2)
}
```


## 2. Growth Curve

For the growth curve, we assume the prey exhibits a Type II functional response, such that higher body velocity has diminishing returns due to time spent handling food:

$$ f(R)= \frac{sR}{1 + shR}$$

where $s$ is prey displacement rate, $h$ is handling time, and $R$ is resource density.

```{r Growth Curve}
## The Growth Curve
# R is always assumed to be 1, while s and h are manipulated values
growth.curve <- function(s, R, h){
  (R * s)/(1 + s*R*h)
}
```


## 3. Fitness Curve

The fitness curve is simply the difference between the growth and predation curves (i.e., b - d as in Abrams 1984).


# IREF Case Studies

## "Baseline" Parameters

We first establish a baseline to compare against future runs, where we will manipulate various parameters.

```{r Baseline Parameters}
## This is the baseline scenario
baseline.df <- as.data.frame( mat.or.vec( length( speed.vec ), 6) )
colnames(baseline.df) <- c("trait", "speed", "growth_curve", "pred_curve", "fitness", "NCE_CE")
baseline.df$trait <- trait.vec
baseline.df$speed <- speed.vec
h_val <- 8 # Prey handling time
R_val <- 1 # Resource density, held constant across all examples at 1
cost_constant_val <- 9 # This scaling constant is used to anchor the growth curve at 1 where s = 1000, its maximum
v_val <- 0.2 # Predator velocity
P_val <- 0.45 # Predator density
pred_constant_val <- 0.98058 # This scaling constant is used to hold the predation curve's intercept constant, i.e., we control for the absolute lethality of the predator between examples

baseline.df$growth_curve <- cost_constant_val * growth.curve( s = baseline.df$speed, R = R_val, h = h_val ) # Growth curve
baseline.df$pred_curve <- pred_constant_val * pred.curve( P = P_val, v = v_val, s = baseline.df$speed ) # Predation curve
baseline.df$fitness <- baseline.df$growth_curve - baseline.df$pred_curve # fitness, b - d
baseline.df$NCE_CE <- (1 - baseline.df$growth_curve)/baseline.df$pred_curve # NCE/CE
```


## Manipulation 1: Predator Density, $P$

Here, we increase predator density, $P$, from 0.45 to 0.9.

```{r Predator Density}
## 1) Change predator density by increasing P
preddens.df <- as.data.frame( mat.or.vec( length( speed.vec ), 6) )
colnames(preddens.df) <- c("trait", "speed", "growth_curve", "pred_curve", "fitness", "NCE_CE")
preddens.df$trait <- trait.vec
preddens.df$speed <- speed.vec
P_val <- 0.9 # The parameters remain as above, but predator density is now doubled

preddens.df$growth_curve <- cost_constant_val * growth.curve( s = preddens.df$speed, R = R_val, h = h_val ) # Growth curve
preddens.df$pred_curve <- pred_constant_val * pred.curve( P = P_val, v = v_val, s = preddens.df$speed ) # Mortality curve
preddens.df$fitness <- preddens.df$growth_curve - preddens.df$pred_curve # fitness
preddens.df$NCE_CE <- (1 - preddens.df$growth_curve)/preddens.df$pred_curve # NCE/CE
```


## Manipulation 2: Predator Body Velocity, $v$

Next, we will consider what happens if prey are faced with a more active predator. We accomplish this by increasing $v$.

```{r Predator Body Velocity}
## 2) Change predator foraging mode from a sit-and-wait to active predator by increasing v
predmode.df <- as.data.frame( mat.or.vec( length( speed.vec ), 6) )
colnames(predmode.df) <- c("trait", "speed", "growth_curve", "pred_curve", "fitness", "NCE_CE")
predmode.df$trait <- trait.vec
predmode.df$speed <- speed.vec
v_val <- 1 # We increase v from 0.2 to 1. We leave predator density as it was before, so that we can compare these two scenarios directly
pred_constant_val <- 0.7071063 # We adjust this constant here to anchor absolute predator lethality in the same place


predmode.df$growth_curve <- cost_constant_val * growth.curve( s = predmode.df$speed, R = R_val, h = h_val ) # Growth curve
predmode.df$pred_curve <- pred_constant_val * pred.curve( P = P_val, v = v_val, s = predmode.df$speed ) # Mortality curve
predmode.df$fitness <- predmode.df$growth_curve - predmode.df$pred_curve # fitness
predmode.df$NCE_CE <- (1 - predmode.df$growth_curve)/predmode.df$pred_curve # NCE/CE
```


## Manipulation 3: Prey Handling time, $h$

Last, we manipulate prey handling time, $h$, to consider what would happen if prey had a faster resource intake rate ($1/h$).

```{r Displacement Case Studies}
## 3) Change prey handling time by increasing h
prey_h_1.df <- as.data.frame( mat.or.vec( length( speed.vec ), 6) )
colnames(prey_h_1.df) <- c("trait", "speed", "growth_curve", "pred_curve", "fitness", "NCE_CE")
prey_h_1.df$trait <- trait.vec
prey_h_1.df$speed <- speed.vec
h_val <- 1 
cost_constant_val <- 2 
v_val <- 0.2 # Predator params return to those from the second example 
pred_constant_val <- 0.98058 # Controlling for absolute lethality


prey_h_1.df$growth_curve <- cost_constant_val * growth.curve( s = prey_h_1.df$speed, R = R_val, h = h_val ) # Growth curve
prey_h_1.df$pred_curve <- pred_constant_val * pred.curve( P = P_val, v = v_val, s = prey_h_1.df$speed ) # Mortality curve
prey_h_1.df$fitness <- prey_h_1.df$growth_curve - prey_h_1.df$pred_curve # fitness
prey_h_1.df$NCE_CE <- (1 - prey_h_1.df$growth_curve)/prey_h_1.df$pred_curve # NCE/CE

# Increase h to 2
prey_h_2.df <- as.data.frame( mat.or.vec( length( speed.vec ), 6) )
colnames(prey_h_2.df) <- c("trait", "speed", "growth_curve", "pred_curve", "fitness", "NCE_CE")
prey_h_2.df$trait <- trait.vec
prey_h_2.df$speed <- speed.vec
h_val <- 2 
cost_constant_val <- 2 # Adjusted so that the growth curve anchors at 1 where s = 1000
v_val <- 0.2 # Predator params return to those from the second example 
pred_constant_val <- 0.98058 # Controlling for absolute lethality


prey_h_2.df$growth_curve <- cost_constant_val * growth.curve( s = prey_h_2.df$speed, R = R_val, h = h_val ) # Growth curve
prey_h_2.df$pred_curve <- pred_constant_val * pred.curve( P = P_val, v = v_val, s = prey_h_2.df$speed ) # Mortality curve
prey_h_2.df$fitness <- prey_h_2.df$growth_curve - prey_h_2.df$pred_curve # fitness
prey_h_2.df$NCE_CE <- (1 - prey_h_2.df$growth_curve)/prey_h_2.df$pred_curve # NCE/CE
```


# Summarization and Data Export

Last, we will generate some useful summary stats for plotting.

```{r Summary Stats}
## Calculate summary stats for each data set, inc. optimum, fitness, CE and NCE strength,  etc
displacement_summary.df <- as.data.frame( mat.or.vec( 5, 8) )
colnames(displacement_summary.df) <- c("case", "v", "P", "h", "optim", "fitness", "CE", "NCE")

displacement_summary.df$case <- c("base", "preddens", "predmode", "preyh_1", "preyh_2")
displacement_summary.df$v <- c("0.2", "0.2", "1", "0.2", "0.2")
displacement_summary.df$P <- c("0.45", "0.9", "0.9", "0.9", "0.9")
displacement_summary.df$h <- c("8", "8", "8", "1", "2")

#For base scenario:
displacement_summary.df[1, 5] <- 
  baseline.df$trait[which(baseline.df$fitness == max(baseline.df$fitness))] # Optimum under 1 - NCE - CE
displacement_summary.df[1, 6] <- 
  baseline.df$fitness[which(baseline.df$fitness == max(baseline.df$fitness))] # Fitness at optimum
displacement_summary.df[1, 7] <- 
  baseline.df$pred_curve[which(baseline.df$fitness == max(baseline.df$fitness))] # CE strength at optimum
displacement_summary.df[1, 8] <- 
  1 - displacement_summary.df[1, 7] - displacement_summary.df[1, 6] # NCE strength at optimum

#For predator density manipulation:
displacement_summary.df[2, 5] <- 
  preddens.df$trait[which(preddens.df$fitness == max(preddens.df$fitness))] # Optimum under 1 - NCE - CE
displacement_summary.df[2, 6] <- 
  preddens.df$fitness[which(preddens.df$fitness == max(preddens.df$fitness))] # Fitness at optimum
displacement_summary.df[2, 7] <- 
  preddens.df$pred_curve[which(preddens.df$fitness == max(preddens.df$fitness))] # CE strength at optimum
displacement_summary.df[2, 8] <- 
  1 - displacement_summary.df[2, 7] - displacement_summary.df[2, 6] # NCE strength at optimum

# For predator mode manipulation:
displacement_summary.df[3, 5] <- 
  predmode.df$trait[which(predmode.df$fitness == max(predmode.df$fitness))] # Optimum under 1 - NCE - CE
displacement_summary.df[3, 6] <- 
  predmode.df$fitness[which(predmode.df$fitness == max(predmode.df$fitness))] # Fitness at optimum
displacement_summary.df[3, 7] <- 
  predmode.df$pred_curve[which(predmode.df$fitness == max(predmode.df$fitness))] # CE strength at optimum
displacement_summary.df[3, 8] <- 
  1 - displacement_summary.df[3, 7] - displacement_summary.df[3, 6] # NCE strength at optimum

# For prey h manipulation:
displacement_summary.df[4, 5] <- 
  prey_h_1.df$trait[which(prey_h_1.df$fitness == max(prey_h_1.df$fitness))] # Optimum under 1 - NCE - CE
displacement_summary.df[4, 6] <- 
  prey_h_1.df$fitness[which(prey_h_1.df$fitness == max(prey_h_1.df$fitness))] # Fitness at optimum
displacement_summary.df[4, 7] <- 
  prey_h_1.df$pred_curve[which(prey_h_1.df$fitness == max(prey_h_1.df$fitness))] # CE strength at optimum
displacement_summary.df[4, 8] <- 
  1 - displacement_summary.df[4, 7] - displacement_summary.df[4, 6] # NCE strength at optimum

displacement_summary.df[5, 5] <- 
  prey_h_2.df$trait[which(prey_h_2.df$fitness == max(prey_h_2.df$fitness))] # Optimum under 1 - NCE - CE
displacement_summary.df[5, 6] <- 
  prey_h_2.df$fitness[which(prey_h_2.df$fitness == max(prey_h_2.df$fitness))] # Fitness at optimum
displacement_summary.df[5, 7] <- 
  prey_h_2.df$pred_curve[which(prey_h_2.df$fitness == max(prey_h_2.df$fitness))] # CE strength at optimum
displacement_summary.df[5, 8] <- 
  1 - displacement_summary.df[5, 7] - displacement_summary.df[5, 6] # NCE strength at optimum
```

```{r NCE % of TPE}
## Calculate change in NCE % of total predator effect, and optimal RITR, with increasing k, v, h:

# For predator density:
P_vec <- seq( 0, 2.5, by = 0.05 )
preddens_pctNCE.df <- as.data.frame(mat.or.vec( length(P_vec), 6 ) ) 
cost_constant_val <- 9
pred_constant_val <- 0.98058
colnames(preddens_pctNCE.df) <- c("P", "NCE", "CE", "NCE_TPE", "OptimalRITR", "Fitness")

for (i in 1:length( P_vec ) ){
preddens_pctNCE.df[i, 1] <- P_vec[i]

pd_growth_curve <- cost_constant_val * growth.curve( s = speed.vec, R = 1, h = 8 ) # Growth curve
pd_pred_curve <- pred_constant_val * pred.curve( P = P_vec[i], v = 0.2, s = speed.vec ) # Mortality curve
pd_fitness <- pd_growth_curve - pd_pred_curve # fitness
pd_NCE_TPE <- (1 - pd_growth_curve)/(pd_pred_curve + (1 - pd_growth_curve))# NCE/TPE

preddens_pctNCE.df[i, 2] <- 1 - pd_growth_curve[ which.max( pd_fitness ) ]
preddens_pctNCE.df[i, 3] <- pd_pred_curve[ which.max( pd_fitness ) ]
preddens_pctNCE.df[i, 4] <- pd_NCE_TPE[ which.max(pd_fitness) ]

# Some other important metrics
preddens_pctNCE.df[i, 5] <- speed.vec[ which.max(pd_fitness) ]
preddens_pctNCE.df[i, 6] <- max(pd_fitness)
}


# For predator foraging mode:
v_vec <- seq( 0, 2.5, by = 0.05 )
predmode_pctNCE.df <- as.data.frame(mat.or.vec( length(v_vec), 6 ) )
pred_constant_val <- 0.7071063 # this changes to anchor the mortality curve in the same spot
colnames(predmode_pctNCE.df) <- c("v", "NCE", "CE", "NCE_TPE", "OptimalRITR", "Fitness")

for (i in 1:length( v_vec ) ){
predmode_pctNCE.df[i, 1] <- v_vec[i]

pm_growth_curve <- cost_constant_val * growth.curve( s = speed.vec, R = 1, h = 8 ) # Growth curve
pm_pred_curve <- pred_constant_val * pred.curve( P = 0.9, v = v_vec[i], s = speed.vec ) # Mortality curve
pm_fitness <- pm_growth_curve - pm_pred_curve # fitness
pm_NCE_TPE <- (1 - pm_growth_curve)/(pm_pred_curve + (1 - pm_growth_curve))# NCE/TPE

predmode_pctNCE.df[i, 2] <- 1 - pm_growth_curve[ which.max( pm_fitness ) ]
predmode_pctNCE.df[i, 3] <- pm_pred_curve[ which.max( pm_fitness ) ]
predmode_pctNCE.df[i, 4] <- pm_NCE_TPE[ which.max(pm_fitness) ]

# Some other important metrics
predmode_pctNCE.df[i, 5] <- speed.vec[ which.max(pm_fitness) ]
predmode_pctNCE.df[i, 6] <- max(pm_fitness)
}


# For prey handling time:
h_vec <- seq( 0.75, 10, by = 0.25 )
preyh_pctNCE.df <- as.data.frame(mat.or.vec( length(h_vec), 6 ) ) 
cost_constant_val <- 2
pred_constant_val <- 0.98058
colnames(preyh_pctNCE.df) <- c("h", "NCE", "CE", "NCE_TPE", "OptimalRITR", "Fitness")

for (i in 1:length( h_vec ) ){
preyh_pctNCE.df[i, 1] <- h_vec[i]

ph_growth_curve <- growth.curve( s = speed.vec, R = 1, h = h_vec[i] ) * cost_constant_val # Growth curve
ph_pred_curve <- pred_constant_val * pred.curve( P = 0.9, v = 0.2, s = speed.vec ) # Mortality curve
ph_fitness <- ph_growth_curve - ph_pred_curve # fitness

preyh_pctNCE.df[i, 2] <- 1 - ph_growth_curve[ which.max( ph_fitness ) ]
preyh_pctNCE.df[i, 3] <- ph_pred_curve[ which.max( ph_fitness ) ]
preyh_pctNCE.df[i, 4] <- preyh_pctNCE.df[i, 2] / ( preyh_pctNCE.df[i, 2] + preyh_pctNCE.df[i, 3] )

# Some other important metrics
preyh_pctNCE.df[i, 5] <- speed.vec[ which.max(ph_fitness) ]
preyh_pctNCE.df[i, 6] <- max(ph_fitness)
}
```

```{r Displacement Data Export, eval = FALSE}
# Export the six data sets
write.csv(baseline.df, "1-1_displacement_baseline.csv", row.names = FALSE)
write.csv(preddens.df, "1-2-1_displacement_preddens.csv", row.names = FALSE)
write.csv(preddens_pctNCE.df, "1-2-2_preddens_TPE_Opt.csv", row.names = FALSE)
write.csv(predmode.df, "1-3-1_displacement_predmode.csv", row.names = FALSE)
write.csv(predmode_pctNCE.df, "1-3-2_predmode_TPE_Opt.csv", row.names = FALSE)
write.csv(prey_h_1.df, "1-4-1_displacement_preyh_1.csv", row.names = FALSE)
write.csv(prey_h_2.df, "1-4-1_displacement_preyh_2.csv", row.names = FALSE)
write.csv(preyh_pctNCE.df, "1-4-2_preyh_TPE_Opt.csv", row.names = FALSE)
write.csv(displacement_summary.df, "1-5_displacement_summary.csv", row.names = FALSE)
```