---
title: "Warming Scenarios"
author: "Andy Davidson"
date: "2025-04-10"
output: pdf_document
---

```{r, echo = FALSE, warning = FALSE}
library(ggplot2)
library(formatR)
library(knitr)
library(cowplot)
knitr::opts_chunk$set(tidy.opts = list(width.cutoff = 60), tidy = TRUE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

# Model Description

For the following scenarios, we will consider a predator-prey system where the prey reduces its foraging activity to avoid predation. We will simulate this for a baseline scenario and several "warmed" scenarios, which we will accomplish by manipulating specific parameters relating to the foraging process of the predator and prey.

Temperature governs the rate of metabolic and physiological processes in ectotherms in a way that is generally nonlinear and hump-shaped, i.e., warming temperatures will first increase traits such as foraging rates, resource handling times, etc., but eventually thermal stress causes those rates to reach a maximum and then decline. For now, we will explicitly ignore scenarios where warming leads to supra-optimal temperatures for the predator or prey. Thus, warmer temperatures will always increase predator and prey metabolism under our scenarios.

We will describe the process for deriving the mortality- and growth-trait curves below, then use those curves to predict the optimal antipredator strategy and the strength of CEs and NCEs under warming conditions.

## The Mortality-Trait Curve

The mortality-trait curve is determined by the predator's encounter rate ($Z$) with its prey, modeled as:

$$ Z = 2rN*sqrt(v^2 + s^2)$$

where $s$ is prey velocity (i.e., displacement rate), $v$ is predator velocity, $N$ is predator density, and $r$ is the radius within which the predator can perceive prey. For simplicity, we will keep detection radius constant such that $2r = 1$, and we will hold predator density constant at $N = 0.45$, both of which match the baseline scenario in the grant (Fig. 2).

In the warming scenarios, we will assume that the predator is an ectotherm and that temperature affects its foraging rates (which are captured by proxy in predator body velocity, $v$), because at warmer temperatures their metabolic needs will generally be greater. In the baseline IREF simulation (Fig. 2 in the grant), $v = 0.2$, so we will use this value and manipulate it to reflect warming by increasing it by 50%.

Thus, like in the grant, this yields a mortality-trait curve that is slightly concave up with respect to prey body velocity, because at high prey displacement rate, encounters are driven largely by prey velocity, whereas at low prey displacement rate, encounters are driven largely by predator velocity. As temperatures warm, this will be most similar to our previous scenarios involving predator foraging mode, i.e., the predator will be more active and thus the mortality-trait curve will be shallower (as in Fig. 3b in the grant).

Below is a figure demonstrating what increasing predator velocity (by 50%, 100%, and 200%) does to the mortality-trait curve.


```{r, echo=FALSE}
## The mortality-trait curve
# This forumula is simplified by setting some of the values as constants
# r is assumed to be 0.5 (which cancels out the 2), while k, v, and s are manipulated values
# k is predator density
# v is predator velocity
# s is prey velocity

pred.curve <- function(k, v, s){
 k * sqrt( v^2 + s^2 )
}

# Include plots of all four mortality-trait curves here, on one figure
s_vec <- seq( 0, 1, by = 0.05 )
v_vec <- c(0.2, 0.3, 0.4, 0.6) # predator velocity
k_val <- 0.45 # predator density outside refuge
pred_constant_val <- 0.98058 # a constant Scott applied in the calcs to the mortality curve to make the numbers cleaner

mortcurve.df <- as.data.frame(mat.or.vec(length(s_vec) * length(v_vec), 3))
colnames(mortcurve.df) <- c("s", "v", "mort")

row <- 0

for (i in 1:length(v_vec)){
  for (j in 1:length(s_vec)){
  row <- row + 1
  mortcurve.df$s[row] <- s_vec[j]
  mortcurve.df$v[row] <- v_vec[i]
  mortcurve.df$mort[row] <- pred.curve( k = 0.45, v = v_vec[i], s = s_vec[j] )
  }
}

mortcurve.df$v <- as.factor(mortcurve.df$v)

for (i in 1:length(v_vec)){ # scale each mortality curve so that the y-intercepts are kept constant
  maxmort <- max(mortcurve.df$mort[which(mortcurve.df$v == v_vec[i])])
  mortcurve.df$mort[which(mortcurve.df$v == v_vec[i])] <- ( mortcurve.df$mort[which(mortcurve.df$v == v_vec[i])] / maxmort ) * 0.9
}

ggplot( data = mortcurve.df, aes( x = s, y = mort, colour = v ) ) +
  geom_smooth(se = FALSE) +
  theme_classic( base_size = 10, base_line_size = 1 ) + 
  ylab("Mortality") +
  xlab("Displacement Rate") +
  scale_x_reverse(expand = c(0, 0), breaks = c(0, 0.5, 1), labels = c(0, 500, 1000) ) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1), breaks = c(0, 0.5, 1) ) +
  theme( panel.border = element_rect( color = "black", fill = NA, linewidth = 0.75 ), aspect.ratio = 1 )

```
\pagebreak

## The Growth-Trait Curve

Manipulating the growth-trait curve to reflect warming is a bit more challenging. Thermal ecology theory predicts that prey will generally be more active with warming (like their predators), but we will assume in these scenarios that prey will prioritize responding to predators when determining their activity (and thus displacement rate) across the landscape. 

However, other traits like their resource processing rate ($1/h$) are metabolically governed in a way that the prey may have less control over (beyond seeking thermal refugia, which we will assume they cannot do here), so this is the parameter we will manipulate. As above, we will decrease handling time (thus increasing resource processing rate) by 50%.

As in the grant, we will assume that the growth-trait curve is determined by the rate at which the prey consumes its resource, $R$. This takes the form of a Type II functional response:

$$ f(R)= \frac{sR}{1 + shR}$$

where $s$ is prey displacement rate, $h$ is handling time, and $R$ is resource density. This results in a growth curve that is strongly concave down, because higher body velocity has diminishing returns due to time spent handling food. As conditions warm, we expect that the growth curve will become more linear (as in Fig. 3c in the grant, where we manipulated prey resource processing rate).

Below is a figure demonstrating what decreasing prey handling time (by 50%, 100%, and 200%) does to the growth-trait curve.


```{r, echo = FALSE}
## The growth-trait curve
# R is prey resource density
# s is prey body velocity
# h is handling time

growth.curve <- function(s, h, R){
  ( s * R )/( 1 + s * h * R )
}

# Include plots of all four growth-trait curves here, on one figure
h_vec <- c(8, 4, 2, 1) # prey handling time
R_val <- 1 # resource density
cost_constant_val <- 9 # a constant Scott applied to the growth curve in the calcs to make the numbers cleaner

growthcurve.df <- as.data.frame(mat.or.vec(length(s_vec) * length(v_vec), 3))
colnames(growthcurve.df) <- c("s", "h", "growth")

row <- 0

for (i in 1:length(h_vec)){
  for (j in 1:length(s_vec)){
  row <- row + 1
  growthcurve.df$s[row] <- s_vec[j]
  growthcurve.df$h[row] <- h_vec[i]
  growthcurve.df$growth[row] <- growth.curve( s = s_vec[j], h = h_vec[i], R = R_val )
  }
}

growthcurve.df$h <- as.factor(growthcurve.df$h)

for (i in 1:length(h_vec)){ # scale each growth curve so that y = 1 at x = 1000
  maxgrowth <- max(growthcurve.df$growth[which(growthcurve.df$h == h_vec[i])])
  growthcurve.df$growth[which(growthcurve.df$h == h_vec[i])] <- growthcurve.df$growth[which(growthcurve.df$h == h_vec[i])] / maxgrowth 
}

ggplot( data = growthcurve.df, aes( x = s, y = growth, colour = h ) ) +
  geom_smooth(se = FALSE) +
  theme_classic( base_size = 10, base_line_size = 1 ) + 
  ylab("Growth") +
  xlab("Displacement Rate") +
  scale_x_reverse(expand = c(0, 0), breaks = c(0, 0.5, 1), labels = c(0, 500, 1000) ) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1), breaks = c(0, 0.5, 1) ) +
  theme( panel.border = element_rect( color = "black", fill = NA, linewidth = 0.75 ), aspect.ratio = 1 )

```
\pagebreak

# Warming Scenarios 

We will now explore how, taken together, changes in the growth and mortality curve due to warming influence this predator-prey interaction. To do so, we will examine different combinations of our two variables of interest ($v$ and $1/h$) at two different levels (baseline and a 50% increase; i.e., a 2x2 design). We can use this manipulation to ask the question: how does warming impact this interaction --- specifically, the relative strength of NCEs and CEs, and the optimal antipredator strategy --- when the predator and prey respond to warming in different ways? This is analogous to exploring scenarios with predator and prey species (or populations) with different levels of sensitivity to warming.

This produces the following trait curves and predictions under the IREF.


```{r, echo = FALSE}
## Generate four plots of the trait curves
## Need to separate them out into dataframes
growthcurve_base.df <- growthcurve.df[which(growthcurve.df$h == 8),] 
mortcurve_base.df <- mortcurve.df[which(mortcurve.df$v == 0.2),]
growthcurve_manip.df <- growthcurve.df[which(growthcurve.df$h == 4),]
mortcurve_manip.df <- mortcurve.df[which(mortcurve.df$v == 0.3),]

basescen.df <- merge(growthcurve_base.df, mortcurve_base.df, by = 's')
predmanip.df <- merge(growthcurve_base.df, mortcurve_manip.df, by = 's')
preymanip.df <- merge(growthcurve_manip.df, mortcurve_base.df, by = 's')
bothmanip.df <- merge(growthcurve_manip.df, mortcurve_manip.df, by = 's')


## Plot 1: no change in v, no change in 1/h
# Calculate optimum, NCE and CE strength, fitness
basescen.df$fitness <- 1 - ( 1 - basescen.df$growth ) - basescen.df$mort
basescen_optim <- round(basescen.df$s[which(basescen.df$fitness == max(basescen.df$fitness))], digits = 2) 
basescen_optim_fitness <- round(basescen.df$fitness[which(basescen.df$fitness == max(basescen.df$fitness))], digits = 2)
basescen_optim_CE <- round(basescen.df$mort[which(basescen.df$fitness == max(basescen.df$fitness))], digits = 2) 
basescen_optim_NCE <- round(1 - basescen_optim_CE - basescen_optim_fitness, digits = 2) 

base.plot <- ggplot( data = basescen.df, aes( x = s ) ) +
  geom_smooth( aes( y = growth, colour = "Growth" ), se = FALSE ) +
  geom_smooth( aes( y = mort, colour = "Mortality" ), se = FALSE ) +
  theme_classic( base_size = 10, base_line_size = 1 ) + 
  ylab("Growth and Mortality") +
  xlab("Displacement Rate") +
  geom_vline( xintercept = basescen_optim,  size = 0.5, linetype = "dashed", colour = "green" ) +
  annotate( "text", x = basescen_optim, y = 0.9, label = paste0("NCE = ", basescen_optim_NCE), size = 8/.pt  ) +
  annotate( "text", x = basescen_optim, y = 0.5, label = paste0("Fitness = ", basescen_optim_fitness), size = 8/.pt  ) +
  annotate( "text", x = basescen_optim, y = 0.2, label = paste0("CE = ", basescen_optim_CE), size = 8/.pt  ) +
  annotate( "text", x = basescen_optim, y = 0.1, label = paste0("Optim = ", basescen_optim * 1000), size = 8/.pt  ) +
  scale_x_reverse(expand = c(0, 0), breaks = c(0, 0.5, 1), labels = c(0, 500, 1000) ) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1), breaks = c(0, 0.5, 1) ) +
  scale_colour_manual(name = "Curve", values = c("blue", "red")) +
  theme( panel.border = element_rect( color = "black", fill = NA, linewidth = 0.75 ), aspect.ratio = 1,
         axis.text.x  = element_blank(),
         axis.title.x = element_blank(),
         legend.position = "none" )


## Plot 2: 50% change in v, no change in 1/h
# Calculate optimum, NCE and CE strength, fitness
predmanip.df$fitness <- 1 - ( 1 - predmanip.df$growth ) - predmanip.df$mort
predmanip_optim <- round(predmanip.df$s[which(predmanip.df$fitness == max(predmanip.df$fitness))], digits = 2) 
predmanip_optim_fitness <- round(predmanip.df$fitness[which(predmanip.df$fitness == max(predmanip.df$fitness))], digits = 2)
predmanip_optim_CE <- round(predmanip.df$mort[which(predmanip.df$fitness == max(predmanip.df$fitness))], digits = 2) 
predmanip_optim_NCE <- round(1 - predmanip_optim_CE - predmanip_optim_fitness, digits = 2) 

predmanip.plot <- ggplot( data = predmanip.df, aes( x = s ) ) +
  geom_smooth( aes( y = growth, colour = "Growth" ), se = FALSE ) +
  geom_smooth( aes( y = mort, colour = "Mortality" ), se = FALSE ) +
  theme_classic( base_size = 10, base_line_size = 1 ) + 
  ylab("Growth and Mortality") +
  xlab("Displacement Rate") +
  geom_vline( xintercept = predmanip_optim,  size = 0.5, linetype = "dashed", colour = "green" ) +
  annotate( "text", x = predmanip_optim, y = 0.9, label = paste0("NCE = ", predmanip_optim_NCE), size = 8/.pt  ) +
  annotate( "text", x = predmanip_optim, y = 0.5, label = paste0("Fitness = ", predmanip_optim_fitness), size = 8/.pt  ) +
  annotate( "text", x = predmanip_optim, y = 0.2, label = paste0("CE = ", predmanip_optim_CE), size = 8/.pt  ) +
  annotate( "text", x = predmanip_optim, y = 0.1, label = paste0("Optim = ", predmanip_optim * 1000), size = 8/.pt  ) +
  scale_x_reverse(expand = c(0, 0), breaks = c(0, 0.5, 1), labels = c(0, 500, 1000) ) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1), breaks = c(0, 0.5, 1) ) +
  scale_colour_manual(name = "Curve", values = c("blue", "red")) +
  theme( panel.border = element_rect( color = "black", fill = NA, linewidth = 0.75 ), aspect.ratio = 1,
         axis.text.y = element_blank(),
         axis.title.y = element_blank(),
         axis.text.x  = element_blank(),
         axis.title.x = element_blank(),
         legend.position = "none" )


## Plot 3: no change in v, 50% change in 1/h
# Calculate optimum, NCE and CE strength, fitness
preymanip.df$fitness <- 1 - ( 1 - preymanip.df$growth ) - preymanip.df$mort
preymanip_optim <- round(preymanip.df$s[which(preymanip.df$fitness == max(preymanip.df$fitness))], digits = 2) 
preymanip_optim_fitness <- round(preymanip.df$fitness[which(preymanip.df$fitness == max(preymanip.df$fitness))], digits = 2)
preymanip_optim_CE <- round(preymanip.df$mort[which(preymanip.df$fitness == max(preymanip.df$fitness))], digits = 2) 
preymanip_optim_NCE <- round(1 - preymanip_optim_CE - preymanip_optim_fitness, digits = 2) 

preymanip.plot <- ggplot( data = preymanip.df, aes( x = s ) ) +
  geom_smooth( aes( y = growth, colour = "Growth" ), se = FALSE ) +
  geom_smooth( aes( y = mort, colour = "Mortality" ), se = FALSE ) +
  theme_classic( base_size = 10, base_line_size = 1 ) + 
  ylab("Growth and Mortality") +
  xlab("Displacement Rate") +
  geom_vline( xintercept = preymanip_optim,  size = 0.5, linetype = "dashed", colour = "green" ) +
  annotate( "text", x = preymanip_optim, y = 0.9, label = paste0("NCE = ", preymanip_optim_NCE), size = 8/.pt  ) +
  annotate( "text", x = preymanip_optim, y = 0.5, label = paste0("Fitness = ", preymanip_optim_fitness), size = 8/.pt  ) +
  annotate( "text", x = preymanip_optim, y = 0.2, label = paste0("CE = ", preymanip_optim_CE), size = 8/.pt  ) +
  annotate( "text", x = preymanip_optim, y = 0.1, label = paste0("Optim = ", preymanip_optim * 1000), size = 8/.pt  ) +
  scale_x_reverse(expand = c(0, 0), breaks = c(0, 0.5, 1), labels = c(0, 500, 1000) ) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1), breaks = c(0, 0.5, 1) ) +
  scale_colour_manual(name = "Curve", values = c("blue", "red")) +
  theme( panel.border = element_rect( color = "black", fill = NA, linewidth = 0.75 ), aspect.ratio = 1,
         legend.position = "none" )


## Plot 4: 50% change in v, 50% change in 1/h
# Calculate optimum, NCE and CE strength, fitness
bothmanip.df$fitness <- 1 - ( 1 - bothmanip.df$growth ) - bothmanip.df$mort
bothmanip_optim <- round(bothmanip.df$s[which(bothmanip.df$fitness == max(bothmanip.df$fitness))], digits = 2) 
bothmanip_optim_fitness <- round(bothmanip.df$fitness[which(bothmanip.df$fitness == max(bothmanip.df$fitness))], digits = 2)
bothmanip_optim_CE <- round(bothmanip.df$mort[which(bothmanip.df$fitness == max(bothmanip.df$fitness))], digits = 2) 
bothmanip_optim_NCE <- round(1 - bothmanip_optim_CE - bothmanip_optim_fitness, digits = 2) 

bothmanip.plot <- ggplot( data = bothmanip.df, aes( x = s ) ) +
  geom_smooth( aes( y = growth, colour = "Growth" ), se = FALSE ) +
  geom_smooth( aes( y = mort, colour = "Mortality" ), se = FALSE ) +
  theme_classic( base_size = 10, base_line_size = 1 ) + 
  ylab("Growth and Mortality") +
  xlab("Displacement Rate") +
  geom_vline( xintercept = bothmanip_optim,  size = 0.5, linetype = "dashed", colour = "green" ) +
  annotate( "text", x = bothmanip_optim, y = 0.9, label = paste0("NCE = ", bothmanip_optim_NCE), size = 8/.pt  ) +
  annotate( "text", x = bothmanip_optim, y = 0.5, label = paste0("Fitness = ", bothmanip_optim_fitness), size = 8/.pt  ) +
  annotate( "text", x = bothmanip_optim, y = 0.2, label = paste0("CE = ", bothmanip_optim_CE), size = 8/.pt  ) +
  annotate( "text", x = bothmanip_optim, y = 0.1, label = paste0("Optim = ", bothmanip_optim * 1000), size = 8/.pt  ) +
  scale_x_reverse(expand = c(0, 0), breaks = c(0, 0.5, 1), labels = c(0, 500, 1000) ) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1), breaks = c(0, 0.5, 1) ) +
  scale_colour_manual(name = "Curve", values = c("blue", "red")) +
  theme( panel.border = element_rect( color = "black", fill = NA, linewidth = 0.75 ), aspect.ratio = 1,
         axis.text.y = element_blank(),
         axis.title.y = element_blank(),
         legend.position = "bottom" )

## Stitch them all together
get_legend_35 <- function(plot) {
  # return all legend candidates
  legends <- get_plot_component(plot, "guide-box", return_all = TRUE)
  # find non-zero legends
  nonzero <- vapply(legends, \(x) !inherits(x, "zeroGrob"), TRUE)
  idx <- which(nonzero)
  # return first non-zero legend if exists, and otherwise first element (which will be a zeroGrob) 
  if (length(idx) > 0) {
    return(legends[[idx[1]]])
  } else {
    return(legends[[1]])
  }
}

fig.legend <- get_legend_35(bothmanip.plot)
bothmanip.plot <- bothmanip.plot + theme(legend.position = "none")

fig.plot <- plot_grid( base.plot, predmanip.plot, preymanip.plot, bothmanip.plot, ncol = 2, labels = c("a) Neither", "b) Pred", "c) Prey", "d) Both"), label_size = 8, label_y = 1, label_x = 0, align = c("hv") )
plot_grid( fig.plot, fig.legend, rel_heights = c(1, 0.1), ncol = 1)
```

In the above figure, the growth and mortality rates are shown as a function of the proportion of prey displacement rate (i.e., foraging rate). The blue line is the growth-trait curve, while the red dashed line is the mortality-trait curve. The optimal level of trait expression is noted with the vertical green dashed line, which also provides estimates for NCE and CE strength as well as overall fitness ($1 - NCE - CE$). 

The panels illustrate the impacts of warming when it leads to either a) no response by predator or prey, b) an increase in only predator body velocity (an increase in $v$ of 50%), c) an increase in only prey resource processing rate (an increase in $1/h$ of 50%), or d) an increase in both predator body velocity and prey resource processing rate (50% each). These can be thought of as asking the question of: how will warming impact the predator-prey interaction and optimal antipredator strategy of the prey if neither species is temperature-sensitive, only the predator is temperature-sensitive, only the prey is temperature-sensitive, or both are equally temperature-sensitive?

Temperature-driven increases in b) predator body velocity by itself lead to a shift towards 15% weaker NCEs and 25% stronger CEs than the baseline scenario, and 10.5% lower overall prey fitness. In this case, the optimal strategy by prey is to reduce foraging by 16.7% relative to the baseline scenario. 

Temperature-driven increases in c) prey resource processing rate by itself, on the other hand, lead to an increase in both NCE and CE strength (by 20% and 21.9%, respectively) relative to the baseline scenario, and a stronger reduction in prey fitness (22.9%). Here, the optimal strategy is a larger reduction in foraging rates, by 33.3% relative to the baseline scenario.

Together, d) they cause stronger NCEs and CEs than the baseline scenario, but the increase is not additive: NCEs are 15% stronger than the baseline scenario, CEs are 34.4% stronger, and fitness is reduced by 29.1%. Further, the optimal antipredator strategy (i.e., the optimal displacement rate) is the same as it is when only prey resource processing rates are affected by temperature.